### 2022-08-19 Ecov_study/Ecov_plots_GLB_recruitment.Rmd
simple ecov recruit sims

---
title: "A preliminary look at some ecov simulations"
author:
  - Gregory Britten, ...
output:
  pdf_document: default
---

```{r}
em_input <- readRDS(file.path(write.dir, "em_input_GLB_recruitment.RDS"))

yrs <- em_fits[[15]][[1]]$years
```

PLOT SIMULATED AND ESTIMATION ECOV REs
```{r}
par(mfrow=c(2,2))
for(i in 1:4){
  sim <- sim_input[[15]][[i]]$data$Ecov_re
  est <- em_fits[[15]][[i]]$parList$Ecov_re

  plot(yrs,  sim,xlim=c(1980,2021),ylim=range(c(sim,est)))
  points(yrs,est,col='red')
}
```

PLOT ESTIMATED BETAS WITH SIMULATED VALUE
```{r}
ms <- c(1,15)
par(mfrow=c(1,2)) 
for(i in 1:length(ms)){
    hist(unlist(lapply(1:nsim, function(x){
    betas <- em_fits[[ms[i]]][[x]]$parList$Ecov_beta[1,1,1,1]
    return(betas)
})), main='',xlab='')
abline(v=df.mods[ms[i],]$beta)
}
```