### 2022-11-16 Ecov_study/catchability/Catchability_ecov_sims.Rmd
Add functions and code to run env-linked q simulations, still needs OM/EM specifics.

---
title: "Catchability_ecov_sims"
output: html_document
date: "2022-11-14"
---
# Function definitions
Functions based on simulation testing vignette: https://github.com/timjmiller/wham/blob/devel/vignettes/ex10_simulation.Rmd

## Generate simulated OM data
```{r}
#' @param OM A fitted WHAM model to use as the operating model, no default.
#' 
#' @return A simulated OM dataset, may be used to overwrite input$data either entirely or in part in simulations

simOM <- function(OM=NULL){
        input <- OM$input
        dataOM = OM$simulate(complete=TRUE) # Overwrite input with simulated data & return
        return(dataOM)
}
```

## Fit specified EMs to simulated OM data and store results
```{r}
#' @param nsim A number specifying how many simulations to run, default = 1.
#' @param OM A fitted WHAM model to use as an operating model, no default.
#' @param inputEMlist A list of WHAM input data objects for all estimation models, no default.
#' @param outdir A filename specifying the directory where time-stamped results should be stored, if NULL (default) then the results are not saved as RData,
#' 
#' @return A nested list with names and results corresponding to the operating models and all estimation models. For each model results contains a sub-list with items for each simulation's results stored as sub-sub lists (because WHAM results are stored as a list). If outdir != NULL then results are saved as RData and returned as an object from the function.

simTestWHAM <- function(nsim = 1,
                        OM = NULL,
                        inputEMlist = NULL,
                        outdir = NULL){
        # Simulation set-up
        obs_names = c("agg_indices","agg_catch","catch_paa","index_paa", "Ecov_obs", "obsvec") # Data overwritten by OM
        namesEM <- lapply(inputEMlist, FUN = function(inputEMlist){inputEMlist$model_name}) %>% unlist() # EM model names

        # Storage set-up
        results <- NULL #!!!!! Still need to work on setting this up
        storeEM <- lapply(namesEM, FUN = assign, value = list())
        names(storeEM) <- namesEM

        # Generate OM data for all simulations
        sim_inputs <- replicate(nsim, simOM(OM), simplify = FALSE)

        for(isim in 1:nsim){ # Loop over simulations
                for(iEM in 1:length(inputEMlist)){ # Loop over estimation models
                        # Pull EM initial input from list
                        inputEM <- inputEMlist[[iEM]]

                        # Update EM input with OM simulated data
                        inputEM$data[obs_names] = sim_inputs[[isim]][obs_names]

                        # Fit WHAM model to updated EM input and save results
                        resultEM <- fit_wham(inputEM, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = TRUE) #!!!! Figure out how to store

                        # Store model-specific results for given isim here (append EM results to storage list for that EM)
                        storeEM[which(namesEM == namesEM[iEM])][[1]] <- append(storeEM[which(namesEM == namesEM[iEM])], list(resultEM))

                } # End loop over EM
        } # End loop over nsims

        # Store all results
        results$OM <- sim_inputs
        names(results) <- OM$model_name # Name OM column based on model_name
        results <- append(results, storeEM) # Should add named EM results to the final resuts object

        # Save time-stamped results as RData if output directory provided as argument
        timeStamp <- Sys.time() %>% gsub(" ", "_",.) %>% gsub(":", "-", .) # Change millisecond half of Sys.time() output to avoid having spaces/weird characters in filenames
        if(is.null(outdir)==FALSE){
                save(results, file = paste(outdir, paste0("simWHAM_", nsim, "_nsim_", OM$model_name, "_OM_", timeStamp, ".RData"), sep="/"))
        }

        return(results) # Always return results as an object
}
```

### Visual describing results structure
WHAM results objects only displayed for OM simulation 1, similar structure for other simulations and for estimation models.
```{r}
library(DiagrammeR)

grViz("

digraph boxes_and_circles{
## Add node statements
# Define shape
node[shape = box]
OM[label = 'OM \n results$OM']
sim1[label = 'Simulation 1 results \n results$OM[[1]]']
sim2[label = 'Simulation 1 results \n results$OM[[2]]']
sim3[label = 'Simulation 1 results \n results$OM[[3]]']

EM1[label = 'EM1 \n results$EM1']
sim1_1[label = 'Simulation 1 results \n results$EM1[[1]]']
sim1_2[label = 'Simulation 1 results \n results$EM1[[1]]']
sim1_3[label = 'Simulation 1 results \n results$EM1[[1]]']
EM2[label = 'EM2 \n results$EM2']
sim2_1[label = 'Simulation 1 results \n results$EM2[[1]]']
sim2_2[label = 'Simulation 1 results \n results$EM2[[1]]']
sim2_3[label = 'Simulation 1 results \n results$EM2[[1]]']
ageslab[label = 'ages.lab']

# Add edge statements to connect nodes
edge[] # optional arrowhead options
# Connect nodes (may be labeled - appears on line)
results -> OM; OM -> sim1; OM -> sim2; OM -> sim3;
results -> EM1; EM1 -> sim1_1; EM1 -> sim1_2; EM1 -> sim1_3;
results -> EM2; EM2 -> sim2_1; EM2 -> sim2_2; EM2 -> sim2_3;
sim1 -> par;  
sim1 -> fn;
sim1 -> gr;
sim1 -> he;
sim1 -> hessian;
sim1 -> method;
sim1 -> retape;
sim1 -> env;
sim1 -> report;
sim1 -> simulate;
sim1 -> years;
sim1 -> years_full;
sim1 -> ageslab;
sim1 -> model_name;
sim1 -> input;
sim1 -> wham_version;
sim1 -> opt;
sim1 -> date;
sim1 -> dir;
sim1 -> rep;
sim1 -> TMB_version;
sim1 -> parList;
sim1 -> final_gradient;
sim1 -> is_sdrep;
sim1 -> na_sdrep;
sim1 -> runtime;
sim1 -> sdrep;

# Add graph statement
graph[nodesep = 0.1, rankdir = LR] # set distance between nodes # for left-right orientation use , rankdir = LR

}   
")

```

## Post-process results
```{r}
#' @param filenames A vector of strings (including .Rda extensions) indicating what Rdata results files to read in and post process
#' 
#' @return

postprocess_simTestWHAM <- function(filenames = NULL){
        # Set up storage for processed results

        # Loop over files
        for(ifile in 1:length(filenames)){
                # Read in Rdata object
                fileDat <- readRDS(filenames[ifile])

                nsim <- # Based on length of results in given filename

                # Loop over simulations in each file
                for(isim in 1:nsim){
                        # Calculate performance metrics here & add to storage
                }

        } # End loop over filenames
}
```

# Specifics for environmentally-linked catchability WHAM runs
## Define Operating Models (OMs)
All OMs should have catchability linked to an environmental covariate but may differ in 
- Uncertainty around this relationship #!!!!!!!!
- Sign of this relationship (+/-) #!!!!!!!!

Set up input and fit OM1
```{r}
inputOM1 <- prepare_wham_input(model_name = "OM1") #!!!!!!!!!!!!
OM1 <- fit_wham(inputOM1, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = TRUE)
```

## Define Estimation Models (EMs)
Correctly specified link between catchability and the environment
```{r}
EMq_env <- prepare_wham_input(model_name = "EMq_env")
```

Catchability random effect and a link between catchability and the environment
```{r}
EMq_rand_env <- prepare_wham_input(model_name = "EMq_rand_env")
```

Only a catchability random effect, fit to environmental model without link
```{r}
EMq_rand <- prepare_wham_input(model_name ="EMq_rand")
```

No catchability considerations (i.e. status quo), fit to environmental model without link !!!! how does this work if you feed in the ecov_beta?
```{r}
EMq <- prepare_wham_input(model_name = "EMq")
```

## Simulation testing 
```{r}
simTestWHAM(nsim = 5, OM = OM1, inputEMlist = list(EMq_env, EMq_rand_env, EMq_rand, EMq)) # Update once OM/EM specifics finalized!!!!

```

## Post-process simulation results
```{r}
postprocess_simTestWHAM(filenames = c("filename1.Rds", "filename2.Rds", "ect.Rds")) # !!! Update once completed runs
```

!!!! post-process based on storage & how sims run
- could run sims based varying isim
- could also loop over OM but this could be computationally expensive
- I recommend labeling runs with OM and EM names (and setting these up with informative) so it is easy to pull out results & compare based on model setup
  - pull from mod$model_name to help aggregate data

proposed function with:
- fitted OM model object
- list of EM inputs

### Result formatted as list of lists, save as rdata object in specified directory with timestamp in name
run1_11_14_22.Rda<-result
 - EM1 
 - EM2 
 - EM3
 - EM4
 - OM result

# Other things to consider
- Other EMs can be developed (e.g. with alternative ecov process assumptions) but should be given DIFFERENT model_names to help with post-processing

# Short script for function testing purposes only!!!
```{r}
library(wham)
# For testing function
testdata <- read_asap3_dat(paste("/Users/amandahart/Research/Plaice_Methods_Development", "data", "PlaiceWHAM-2019_revised_NEFSC-LW-WAA_splitNEFSC-BigUnit_fixSeason.DAT", sep="/")) # this can be swapped out for any ASAP file you have to set up a WHAM model for debugging
testinput <- prepare_wham_input(testdata)
testWHAM <- fit_wham(testinput, MakeADFun.silent = TRUE, do.osa = FALSE, do.retro=FALSE)

OM <- testWHAM
testinput1 <- prepare_wham_input(testdata, model_name = "test1")
testinput2 <- prepare_wham_input(testdata, model_name="test2")
testinput3 <- prepare_wham_input(testdata, model_name="test3")
inputEMlist <- list(testinput1, testinput2, testinput3)

test <- simTestWHAM(nsim = 2, OM = OM, inputEMlist = inputEMlist)
```