### 2022-12-07 Project_0/paper/paper.Rmd
saving progress on Project_0

---
title: State-space age-structured assessment models provide reliable inferences(?)
author: 
  - Timothy J. Miller^1^, Britten, Brooks, Fay, Legault, Stock, 
date: ''
fontsize: 12pt
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    includes:
      in_header:
      - options.sty
      - journalshortcuts.tex
    keep_tex: yes
    number_sections: yes
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{xspace}
- \usepackage{bm}
- \usepackage{caption,graphics}
- \usepackage{graphicx}
- \usepackage{makecell}
- \renewcommand\figurename{Fig.}
- \captionsetup{labelsep=period, singlelinecheck=false}
- \newcommand{\changesize}[1]{\fontsize{#1pt}{#1pt}\selectfont}
- \renewcommand{\arraystretch}{1.5}
- \renewcommand\theadfont{}
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
biblio-style: cjfas2.bst
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE, fig.align = "center")
knitr::opts_knit$set(root.dir = '../' )
library(knitr)
library(tidyverse)
library(pander)
library(kableExtra)
library(png)
library(forcats)
```

$^1$corresponding author: timothy.j.miller@noaa.gov, Northeast Fisheries Science Center, Woods Hole Laboratory, 166 Water Street, Woods Hole, MA 02543 USA\

\pagebreak

## Abstract {-}

### Keywords {-}

\pagebreak

# Introduction

Much is known about the reliabiltiy of state-space models that are linear or Gaussian, but applications in fisheris management are
nonlinear and typically include multiple types of observations with varying distributional assumptions. We know relatively little about
the statistical reliability of such models. Also, there is a wide range of potential random effects structures in assessment models
and we  know little about the ability of information criteria to distinguish among such alternative structures.

But those studies focus primarily on Gaussain 
Reliability of hidden process models

# Methods

Used the WHAM package commit = XXXXXXX (Stock and Miller 2021, Miller and Stock 2020). This packages has also been used
to configure operating and estimating models for closed loop simulations evaluating index-based asssessment methods (Legault et al. In press).

We completed a simulation study with 24 NAA re operating models, 16 M re operating models, 16 Sel re operating models and 16 q re operating models. 
We simulated N data sets for each operating model.
For each simulated data from each operating model we fit a set of estimating models.

Y estimating models fit to each 
## Operating models

common to all:   
  ages, 
  M
  maturity
  waa
  observation error assumptions
  2.5 Fmsy -> Fmsy after 20 years
  Fmsy througout

marginal standard deviations for random effects are defined in tables of operating models.

NAA_oms
  process error assumtions

Population - 
  Initial population, 
  ages, 
  M
  maturity
  waa

Fishing history
All operating models assumed one of two different fishing histories. 
One : Fishing mortality is equal to Fmsy for the whole 40 year period.
Two : Fishing mortality is 2.5 times Fmsy for the first 20 years then changes to Fmsy for the last 20 years.

Overfishing assumptions is based on average estimates of overfishing for NE groundfish stocks from Wiedenman et al. (20XX). Legault et al. (2023) also used similar approaches to defining fishing mortality histories for operating models.

We assumed a Beverton-Holt stock recruit function with constant pre-recruit mortality parameters for all operating models. All post-recruit productivity components are constant in the NAA and survey catchability process error operating models. Therefore steepness and unfished recruitment are also constant over the time period for those operating models (Miller and Brooks 2021). We specified unfished recruitment = $e^10$ and $F_{\text{MSY}} = F_{40}$ equated to a steepness of 0.68 and a and b parameters X and X?

For operating models with time-varying random effects for M, steepness is not constant, but we used the same alpha and beta parameters as other operating models this equates to a steepness and R0 at the mean of the time series process for M.

For operating models with time-varying random effects for fishery selectivity, Fmsy is also not constant however we use the same Fhistory as other operating models which corresposds to Fmsy at the mean selectivity parameters.

Initial population starts at equilibium population structure associated with F= x*Fmsy where x = 1 or 2.5.

surveys, selectivity

fleets, selectivity

M_oms
NOTE: inv_trans_rho function in set_M.R is mis-defined. Will affect correlation parameters assigned in operating models?

## Estimating models

32 estimating models
1-20 fit to each NAA RE operating model
5-24 fit to each M RE operating model
5-20,25-28 to each sel RE operating model
5-20, 29-32 to each q RE operating model

SR estimation or not

Make plot of S-R curve, Fmsy = F40
Initial values for BH parameters are the true values.
Initial values for mean R model = true R0.

M estimation or not

NAA_re
Random effects on Recruitment only or random effects on recruitment and transitions among older numbers at age.

M_re
Random effects on Recruitment only, M constant across age .

sel_re
Random effects on Recruitment only, fleet logistic selectivity RE model?

q_re
Random effects on Recruitment only, one survey catchability RE model?

Simulations were all carried out on the University of Massachussetts Green High-Performance Computing Cluster. Code for completing the simulations and summarization of results can be found at github.com/timjmiller/SSRTWG/Project_0. We used the wham package version 1.X.X (commit XXXXX).

# Results

Do each of these by type of operating model (Naa, M, sel, q)
Convergence performance

AIC performance

SR estimation? M estimation?

Bias, Mean Square error

  Certain basic parameters (stock-recruit pars, M, variance parameters)
  SSB, F, R

\clearpage 
## Numbers at age operating models

### Estimating models include alternative random effects options: NAA, M, sel, q
\clearpage

\begin{table}
\caption{NAA operating models, estimating models all assume a B-H stock recruit relationship and M is fixed at the true value.}
{\input{naa_om_em_SR_MF_aic_table}}
\end{table}

\begin{table}
\caption{NAA operating models, estimating models all assume a B-H stock recruit relationship and M is estimated.}
{\input{naa_om_em_SR_ME_aic_table}}
\end{table}

\begin{table}
\caption{NAA operating models, estimating models all estimate a mean recruitment and M is fixed at the true value.}
{\input{naa_om_em_R_MF_aic_table}}
\end{table}

\begin{table}
\caption{NAA operating models, estimating models all estimate a mean recruitment and M estimated.}
{\input{naa_om_em_R_MF_aic_table}}
\end{table}
\clearpage

\begin{landscape}
\begin{figure}
\caption{Median relative bias of for SSB for estimating models that estimate mean recruitment and M is fixed at the true value.}\label{naa_om_em_R_MF_relbias_ssb}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_R_MF_relbias_ssb.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative bias of for SSB for estimating models that estimate mean recruitment and M is estimated.}\label{naa_om_em_R_ME_relbias_ssb}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_R_ME_relbias_ssb.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative bias of for SSB for estimating models that estimates a BH stock-recruitment function and M is fixed at the true value.}\label{naa_om_em_SR_MF_relbias_ssb}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_SR_MF_relbias_ssb.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative bias of for SSB for estimating models that estimates a BH stock-recruitment function and M is estimated.}\label{naa_om_em_SR_ME_relbias_ssb}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_SR_ME_relbias_ssb.png}
\end{center}
\end{figure}
\end{landscape}

### Estimating models include NAA random effects and estimation assumes mean R or BH SR
\clearpage

\begin{table}
\caption{Operating models and estimation models all assume RE on recruitment only, estimating models assume mean recruitment or a B-H stock recruit relationship and M is fixed at the true value.}
{\input{rec_om_em_R_SR_MF_aic_table}}
\end{table}

\begin{table}
\caption{Operating models and estimation models all assume RE on recruitment only, estimating models assume mean recruitment or a B-H stock recruit relationship and M is estimated.}
{\input{rec_om_em_R_SR_ME_aic_table}}
\end{table}

\begin{table}
\caption{Operating models and estimation models all assume RE on all abundances at age, estimating models assume mean recruitment or a B-H stock recruit relationship and M is fixed at the true value.}
{\input{naa_om_em_R_SR_MF_aic_table}}
\end{table}

\begin{table}
\caption{Operating models and estimation models all assume RE on all abundances at age, estimating models assume mean recruitment or a B-H stock recruit relationship and M is estimated.}
{\input{naa_om_em_R_SR_ME_aic_table}}
\end{table}
\clearpage

# Discussion

The estimating models assumed variances of aggregate catch and index observations was known. This approximation may be appropriate for indices
where we have a reliable estimate of uncertainty based on the survey design (), but there may be better approaches for the aggregate
catch such as an informed prior on the standard errors with realistic bounds.

# Acknowledgements {-}

This work was funded by NOAA Fisheries Northeast Fisheries Science Center.

<!--
# CRediT authorship contribution statement {-}

**Timothy J. Miller:** Conceptualization, Methodology, Writing - original draft, Formal analysis, Visualization.
 -->

\pagebreak

\bibliography{sweep-paper}
<!-- \bibliography{sweep-paper}
\bibliographystyle{/home/tmiller2/work/bibliographies/cjfas2}  -->

<div id="refs"></div>

\pagebreak
<!-- 
\begin{figure}
\caption{Diagram of twin-trawl gear configuration. One of the two nets is rigged with a rockhopper sweep (8) and the other is rigged with a chain sweep (7) and for both a restrictor rope (5) is used to obtain consistent net spread. The other importand components are the side wires (1), middle wire (2), doors (3), the clump weight (4), and the acoustic mensuration system (6).  The side where the rockhopper and chainsweep gears gears were deployed varied throughout the experimental tows of each.}\label{twin_trawl_diagram}
\begin{center}
\includegraphics[width = \textwidth]{twin_trawl_diagram.pdf}
\end{center}
\end{figure}

\begin{figure}
\caption{The \textsl{F/V Karen Elizabeth} twin-trawl vessel rigged with rockhopper sweep gear on the right and chain sweep gear on the left.}\label{twin_trawl_photo}
\begin{center}
\includegraphics[width = \textwidth]{twin_trawl_photo_small.png}
\end{center}
\end{figure} -->

<!--
\begin{figure}
\caption{Diagram of the standard Northeast Fisheries Science Center rockhopper sweep center and wing sections.}\label{rockhopper_schematic}
\begin{center}
\includegraphics[width = 0.7\textwidth]{rockhopper_schematic_1.pdf}
\includegraphics[width = 0.7\textwidth]{rockhopper_schematic_2.pdf}
\end{center}
\end{figure}

\begin{figure}
\caption{Diagram of the chain sweep designed maximize bottom contact and flatfish capture.}\label{chainsweep_schematic}
\begin{center}
\includegraphics[width = \textwidth]{chainsweep_schematic.pdf}
\end{center}
\end{figure}
\clearpage
  -->
<!-- \begin{landscape}
\begin{figure}
\caption{Annual locations of stations where the F/V Karen Elizabeth conducted twin-trawl sets with the standard bottom trawl gear and the gear with a chain sweep instead of the rockhopper sweep.}\label{tow_locations}
\begin{center}
\includegraphics[width = \textwidth]{TwinTrawlLocationsAllYears.jpg}
\end{center}
\end{figure}
\end{landscape}

\clearpage

\begin{figure}
\caption{Relative efficiency  of gears using chain and rockhopper sweeps from the best performing model for each species (Table \ref{best_model_compare}). Blue and red denote results for day and night data, respectively, and thick and thin lines represent overall and paired-tow specific estimates of relative catch efficiency, respectively. There was no diel effect in the best model for American plaice. Points represent empirical estimates of relative efficiency for paired observation by length and paired tow. Polygons and dashed lines represent hessian-based and bootstrap-based 95\% confidence intervals, respectively.}\label{sp_rho_plot}
\begin{center}
\includegraphics[height = 0.8\textheight]{sp_rho_plot.pdf}
\end{center}
\end{figure}

\clearpage

\begin{figure}
\caption{Annual spring (blue) and fall (red) biomass estimates for each managed stock assuming 100\% efficiency for chain sweep gear with shaded polygons representing bootstrap-based 95\% confidence intervals. Relative catch efficiency at size estimates and bootstraps are from the best performing model for each species (Table \ref{best_model_compare}).}\label{stock_biomass_plot}
\begin{center}
\includegraphics[height = 0.8\textheight]{stock_biomass_plot.pdf}
\end{center}
\end{figure}

\clearpage

\begin{figure}
\caption{Implied catch efficiency of annual spring (blue) and fall (red) bottom trawl survey biomass estimates for each managed stock assuming 100\% efficiency for chain sweep gear with shaded polygons representing bootstrap-based 95\% confidence intervals. Relative catch efficiency at size estimates and bootstraps are from the best performing model for each species (Table \ref{best_model_compare}).}\label{stock_biomass_efficiency_plot}
\begin{center}
\includegraphics[height = 0.8\textheight]{stock_biomass_efficiency_plot.pdf}
\end{center}
\end{figure} -->

\clearpage

<!-- \begin{table}
\caption{Managed stocks associated with the species for which relative catch efficiency was estimated.}\label{stock_definition_table}
{\input{stock_definition_table}}
\end{table}

\begin{landscape}
\begin{table}
\caption{Description of relative catch efficiency ($\rho$) and beta-binomial dispersion ($\phi$) parameterizations for binomial and beta-binomial models and number of marginal likelihood parameters ($n_p$) for the 13 base models from \citet{miller13} and fit to paired chain sweep and rockhoppersweep tow data for each species.}\label{base_model_description}
{\input{base_model_description}}
\end{table}
\end{landscape}

\begin{landscape}
\begin{table}
\caption{Number of paired tows where fish were captured and the number of fish captured and measured for lengths for each species in total and by day or night.}\label{data_table}
{\changesize{10}\input{data_table}}
\end{table}
\end{landscape}

\begin{landscape}
\begin{table}\caption{Difference in AIC for each of the 13 models described in Table \ref{base_model_description} from the best model ($\bf 0$) by species.}\label{base_model_compare}
{\input{base_model_compare}}
\end{table}
\end{landscape}

\begin{landscape}
\begin{table}\caption{Best performing models from Table \ref{base_model_compare} and extended models that include diel effects on relative catch efficiency for each species with the number of parameters for each model ($n_p$) and the differences in AIC ($\Delta$AIC) from the best of the three models ($\bf 0$) by species.}\label{best_model_compare}
{\changesize{7}\input{best_model_compare}}
\end{table}
\end{landscape}

\begin{table}
\caption{Average of annual (2009-2019) ratios of coefficients of variation for calibrated and uncalibrated biomass indices for each stock by seasonal survey. Coefficients of variation are based on bootstrap resampling of paired tow observations, survey station data and associated length and weight observations. Annual indices for fall 2017 were not available for summer flounder, SNE-MA windowpane, and SNE-MA yellowtail flounder.}\label{stock_cv_ratios}
{\input{stock_cv_ratios}}
\end{table}

\begin{table}
\caption{Average correlation of annual (2009-2019) calibrated biomass indices for each stock by seasonal survey. Annual indices for fall 2017 were not available for SNE-MA windowpane and SNE-MA yellowtail flounder.}\label{stock_mean_cor}
{\input{stock_mean_cor}}
\end{table} -->