### 2022-12-16 Ecov_study/catchability/Catchability_ecov_sims.Rmd
Update results overview figure, add code to postprocess multiple simulation results files.

---
title: "Catchability_ecov_sims"
output: html_document
date: "2022-11-14"
---
# Function definitions
Functions based on simulation testing vignette: https://github.com/timjmiller/wham/blob/devel/vignettes/ex10_simulation.Rmd

## Generate simulated OM data
```{r}
#' @param OM A fitted WHAM model to use as the operating model, no default.
#' 
#' @return A simulated OM dataset, may be used to overwrite input$data either entirely or in part in simulations

simOM <- function(OM=NULL){
        input <- OM$input
        dataOM = OM$simulate(complete=TRUE) # Overwrite input with simulated data & return
        return(dataOM)
}
```

## Fit specified EMs to simulated OM data and store results
```{r}
#' @param nsim A number specifying how many simulations to run, default = 1.
#' @param OM A fitted WHAM model to use as an operating model, no default.
#' @param inputEMlist A list of WHAM input data objects for all estimation models, no default.
#' @param outdir A filename specifying the directory where time-stamped results should be stored, if NULL (default) then the results are not saved as RData,
#' 
#' @return A nested list with names and results corresponding to the operating models and all estimation models. For each model results contains a sub-list with items for each simulation's results stored as sub-sub lists (because WHAM results are stored as a list). If outdir != NULL then results are saved as RData and returned as an object from the function.

simTestWHAM <- function(nsim = 1,
                        OM = NULL,
                        inputEMlist = NULL,
                        outdir = NULL){
        # Simulation set-up
        obs_names = c("agg_indices","agg_catch","catch_paa","index_paa", "Ecov_obs", "obsvec") # Data overwritten by OM
        namesEM <- lapply(inputEMlist, FUN = function(inputEMlist){inputEMlist$model_name}) %>% unlist() # EM model names

        # Storage set-up
        results <- NULL #!!!!! Still need to work on setting this up
        storeEM <- lapply(namesEM, FUN = assign, value = list())
        names(storeEM) <- namesEM

        # Generate OM data for all simulations
        sim_inputs <- replicate(nsim, simOM(OM), simplify = FALSE)

        for(isim in 1:nsim){ # Loop over simulations
          print(paste0("Simulation_", isim))
                for(iEM in 1:length(inputEMlist)){ # Loop over estimation models
                        # Pull EM initial input from list
                        inputEM <- inputEMlist[[iEM]]

                        # Update EM input with OM simulated data
                        inputEM$data[obs_names] = sim_inputs[[isim]][obs_names]

                        # Fit WHAM model to updated EM input and save results
                        resultEM <- fit_wham(inputEM, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = TRUE) #!!!! Figure out how to store

                        # Store model-specific results for given isim here (append EM results to storage list for that EM)
                        storeEM[which(namesEM == namesEM[iEM])][[1]] <- append(storeEM[which(namesEM == namesEM[iEM])][[1]], list(resultEM))

                } # End loop over EM
        } # End loop over nsims

        # Store all results
        results$OM <- sim_inputs
        names(results) <- OM$model_name # Name OM column based on model_name
        results <- append(results, storeEM) # Should add named EM results to the final resuts object

        # Save time-stamped results as RData if output directory provided as argument
        timeStamp <- Sys.time() %>% gsub(" ", "_",.) %>% gsub(":", "-", .) # Change millisecond half of Sys.time() output to avoid having spaces/weird characters in filenames
        if(is.null(outdir)==FALSE){
                save(results, file = paste(outdir, paste0("simWHAM_", nsim, "_nsim_", OM$model_name, "_OM_", timeStamp, ".RData"), sep="/"))
        }

        return(results) # Always return results as an object
}
```

### Visual describing results structure
WHAM results objects only displayed for OM simulation 1, similar structure for other simulations and for estimation models.
```{r}
library(DiagrammeR)

grViz("

digraph boxes_and_circles{
## Add node statements
# Define shape
node[shape = box]
OM[label = 'OM \n results$OM']
sim1[label = 'Simulation 1 results \n results$OM[1][[1]]']
sim2[label = 'Simulation 2 results \n results$OM[2][[1]]']
sim3[label = 'Simulation 3 results \n results$OM[3][[1]]']

EM1[label = 'EM1 \n results$EM1']
sim1_1[label = 'Simulation 1 results \n results[EMs[1]][[1]][1][[1]]']
sim1_2[label = 'Simulation 2 results \n results[EMs[1]][[1]][2][[1]]']
sim1_3[label = 'Simulation 3 results \n results[EMs[1]][[1]][2][[1]]']
EM2[label = 'EM2 \n results$EM2']
sim2_1[label = 'Simulation 1 results \n results[EMs[2]][[1]][1][[1]]']
sim2_2[label = 'Simulation 2 results \n results[EMs[2]][[1]][2][[1]]']
sim2_3[label = 'Simulation 3 results \n results[EMs[2]][[1]][3][[1]]']
ageslab[label = 'ages.lab']

# Add edge statements to connect nodes
edge[] # optional arrowhead options
# Connect nodes (may be labeled - appears on line)
results -> OM; OM -> sim1; OM -> sim2; OM -> sim3;
results -> EM1; EM1 -> sim1_1; EM1 -> sim1_2; EM1 -> sim1_3;
results -> EM2; EM2 -> sim2_1; EM2 -> sim2_2; EM2 -> sim2_3;
sim1 -> par;  
sim1 -> fn;
sim1 -> gr;
sim1 -> he;
sim1 -> hessian;
sim1 -> method;
sim1 -> retape;
sim1 -> env;
sim1 -> report;
sim1 -> simulate;
sim1 -> years;
sim1 -> years_full;
sim1 -> ageslab;
sim1 -> model_name;
sim1 -> input;
sim1 -> wham_version;
sim1 -> opt;
sim1 -> date;
sim1 -> dir;
sim1 -> rep;
sim1 -> TMB_version;
sim1 -> parList;
sim1 -> final_gradient;
sim1 -> is_sdrep;
sim1 -> na_sdrep;
sim1 -> runtime;
sim1 -> sdrep;

# Add graph statement
graph[nodesep = 0.1, rankdir = LR] # set distance between nodes # for left-right orientation use , rankdir = LR

}   
")

```

## Post-process results
```{r}
#' @param filenames A vector of strings (including .RData extensions) indicating what results files to read in and post process
#' 
#' @return A dataframe containing the following columns describing model performance
#' \itemize{
#'   \item{Year - Year associated with performance metric}
#'   \item{sim - Simulation number (based on total number of simulations in all files processed)}
#'   \item{EM - Name of estimation model used to calculate performance metric}
#'   \item{converged - Boolean indicating whether model converged with invertible hessian ("Yes") or not ("No")}
#'   \item{relSSB - Ratio of SSB from EM:OM}
#'   \item{relF - Ratio of F from EM:OM}
#'   \item{relR - Ratio of R from EM:OM}
#'   \item{relEcov_x - Ratio of Ecov_x from EM:OM}
#' }

postprocess_simTestWHAM <- function(filenames = NULL){
  # Set up storage for processed results
  perfMet <- NULL

  # Establish total number of simulations across all files (based on nsim in filename)
  nsim <- rep(NA, length(filenames))
  for(ifile in 1:length(filenames)){
    fileSpecs <- strsplit(filenames[ifile], "_")
    nsim[ifile] <- fileSpecs[[1]][which(grepl("simWHAM", fileSpecs[[1]]))+1] %>% as.numeric()
  }

  nsim <- c(0, nsim) # Add 0 so isim indexing in file processed output goes from 1:total nsim

  # Loop over files to calculate performance metrics
  for(ifile in 1:length(filenames)){
    # Read in RData object
    load(file=filenames[ifile]) # loads object named "results"

    # Loop over simulations in each file
    for(isim in 1:nsim[ifile+1]){ # start at ifile+1 since nsim[ifile] = 0 for sim numbering reasons

      # Pull out EM names 
      EMs <- names(results)[2:length(names(results))]

      # Loop over EM in each ifile
      for(iEM in 1:length(EMs)){

        # Store Years, isim, iEM, convergence (mod$opt$convergence == 0 and hessian invertible $na_sdrep == FALSE)
        Year <- results[EMs[iEM]][[1]][isim][[1]]$years
        sim <- rep((isim+nsim[ifile]), length(Year)) # + 0 for first file, starts sim numbering at nsim[ifile=1]+1 for second file
        EM <- rep(EMs[iEM], length(Year))
        if(check_convergence(results[EMs[iEM]][[1]][isim][[1]], ret=TRUE)$na_sdrep == TRUE){
          converged <- rep("No", length(Year))
        } else{
          converged <- rep("Yes", length(Year))
        }

        # Proportion of EM:OM SSB
          # EM SSB calculated based on plotting script https://github.com/timjmiller/wham/blob/6c30301d5b68c2bb0c0330cc4f112432c7950dfd/R/wham_plots_tables.R#L725-L738
        tempSSB_EM <- results[EMs[iEM]][[1]][isim][[1]]$sdrep %>% summary() #EM SSB must be calculated, this line triggers an NaNs produced warning for EMs that didn't converge with invertible hessians
        tempSSB_EM <- tempSSB_EM[rownames(tempSSB_EM) == "log_SSB",]
        tempSSB_EM <- exp(cbind(tempSSB_EM, tempSSB_EM[,1]+qnorm(0.975)*cbind(-tempSSB_EM[,2], tempSSB_EM[,2])))#/1000 Don't divide by 1000 so units in mt rather than kmt
        tempSSB_EM <- tempSSB_EM[,1] 
          # OM SSB
        tempSSB_OM <- results$OM[isim][[1]]$SSB # Pull SSB directly from simulated data
          # relative SSB
        relSSB <- tempSSB_EM/tempSSB_OM

        # Proportion of EM:OM F
          # EM calculated F based on plotting scripts
        mod <- results[EMs[iEM]][[1]][isim][[1]]
        std = results[EMs[iEM]][[1]][isim][[1]]$sdrep %>% summary()
        years_full <- mod$years_full
        n_ages = mod$env$data$n_ages
        faa.ind <- which(rownames(std) == "log_FAA_tot")
        log.faa <- matrix(std[faa.ind,1], length(years_full), n_ages)
        faa.cv <- matrix(std[faa.ind,2], length(years_full), n_ages)
        age.full.f <- apply(log.faa,1, function(x) max(which(x == max(x))))
        full.f.ind = cbind(1:length(years_full), age.full.f)
        log.full.f <- log.faa[full.f.ind]
        full.f.cv <- faa.cv[full.f.ind]
        full.f <- exp(log.full.f)
          # OM F
        tempF_OM <- c(results$OM[isim][[1]]$F)
          # relative F
        relF <- full.f/tempF_OM

        # Proportion of EM:OM Recruitment 
          # EM R calculated based on plotting script: https://github.com/timjmiller/wham/blob/6c30301d5b68c2bb0c0330cc4f112432c7950dfd/R/wham_plots_tables.R#L765-L780
        tempR_EM <-  results[EMs[iEM]][[1]][isim][[1]]$sdrep %>% summary()
        ind <- rownames(tempR_EM) == "log_NAA_rep"
        tempR_EM <-  exp(array(tempR_EM[ind,1], dim = c(length(results[EMs[iEM]][[1]][isim][[1]]$years), 8)))
        tempR_EM <- tempR_EM[,1]
          # OM recruitment
        tempR_OM <- results$OM[isim][[1]]$NAA[,1] 
          # relative R
        relR <- tempR_EM/tempR_OM

        # Relative Ecov_x values (EM:OM ratio)
        if("TRUE" %in% (results[EMs[iEM]][[1]][isim][[1]]$input$data$Ecov_where > 0)){ # If link to ecov specified in EM calculate EM:OM ratio for Ecov_x values
            # EM
        tempEcovx_EM <- results[EMs[iEM]][[1]][isim][[1]]$sdrep %>% summary() 
        tempEcovx_EM <- tempEcovx_EM[rownames(tempEcovx_EM)=="Ecov_x",]
          # OM
        tempEcovx_OM <- results$OM[isim][[1]]$Ecov_x
          # Relative Ecov_x
        relEcov_x <- tempEcovx_EM[,"Estimate"]/tempEcovx_OM
        } else{ # If no ecov link specified then save an NA for that model in all years 
          relEcov_x <- rep(NA, length(Year))
        }

        # # Relative Ecov_beta (EM:OM ratio) for all four indices
        #   # EM
        # tempEcovbeta_EM <- results[EMs[iEM]][[1]][isim][[1]]$sdrep %>% summary()
        # tempEcovbeta_EM <- tempEcovbeta_EM[rownames(tempEcovbeta_EM)=="Ecov_beta"]
        #   # OM
        # tempEcovbeta_OM <- results$OM[isim][[1]]$Ecov_beta # by age rather than by index?
        #

        # Store results
          # EM perfMets
          EMperfMets <- cbind(Year, sim, EM, converged, relSSB, relF, relR, relEcov_x)

          # Append perfMets to storage
          perfMet <- rbind(perfMet, EMperfMets)

      } # End loop over EM

    } # End loop over simulations in ifile

  } # End loop over filenames

  # Clean table rownames and ensure that numeric results are not treated as strings
  rownames(perfMet) <- NULL
  colnames(perfMet) <- c("Year", "sim", "EM", "converged", "relSSB", "relF", "relR", "relEcov_x")
  perfMet <- perfMet %>% as.data.frame() %>% 
    dplyr::summarise(Year = as.numeric(Year),
                     sim = as.numeric(sim),
                     EM = EM,
                     converged = converged,
                     relSSB = as.numeric(relSSB),
                     relF = as.numeric(relF),
                     relR = as.numeric(relR),
                     relEcov_x = as.numeric(relEcov_x))

  return(perfMet)
}
```

# Specifics for environmentally-linked catchability WHAM runs
## Define Operating Models (OMs)
All OMs should have catchability linked to an environmental covariate but may differ in 
- Uncertainty around this relationship #!!!!!!!!
- Sign of this relationship (+/-) #!!!!!!!!

Set up input and fit OM1
```{r}
inputOM1 <- prepare_wham_input(model_name = "OM1") #!!!!!!!!!!!!
OM1 <- fit_wham(inputOM1, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = TRUE)
```

## Define Estimation Models (EMs)
Correctly specified link between catchability and the environment
```{r}
EMq_env <- prepare_wham_input(model_name = "EMq_env")
```

Catchability random effect and a link between catchability and the environment
```{r}
EMq_rand_env <- prepare_wham_input(model_name = "EMq_rand_env")
```

Only a catchability random effect, fit to environmental model without link
```{r}
EMq_rand <- prepare_wham_input(model_name ="EMq_rand")
```

No catchability considerations (i.e. status quo), fit to environmental model without link !!!! how does this work if you feed in the ecov_beta?
```{r}
EMq <- prepare_wham_input(model_name = "EMq")
```

## Simulation testing 
```{r}
simTestWHAM(nsim = 5, OM = OM1, inputEMlist = list(EMq_env, EMq_rand_env, EMq_rand, EMq)) # Update once OM/EM specifics finalized!!!!

```

## Post-process simulation results
```{r}
postprocess_simTestWHAM(filenames = c("filename1.Rds", "filename2.Rds", "ect.Rds")) # !!! Update once completed runs
```

!!!! post-process based on storage & how sims run
- could run sims based varying isim
- could also loop over OM but this could be computationally expensive
- I recommend labeling runs with OM and EM names (and setting these up with informative) so it is easy to pull out results & compare based on model setup
  - pull from mod$model_name to help aggregate data

proposed function with:
- fitted OM model object
- list of EM inputs

### Result formatted as list of lists, save as rdata object in specified directory with timestamp in name
run1_11_14_22.Rda<-result
 - EM1 
 - EM2 
 - EM3
 - EM4
 - OM result

# Other things to consider
- Other EMs can be developed (e.g. with alternative ecov process assumptions) but should be given DIFFERENT model_names to help with post-processing

# Short script for function testing purposes only!!!
```{r}
library(wham)
# For testing function
testdata <- read_asap3_dat(paste("/Users/amandahart/Research/Plaice_Methods_Development", "data", "PlaiceWHAM-2019_revised_NEFSC-LW-WAA_splitNEFSC-BigUnit_fixSeason.DAT", sep="/")) # this can be swapped out for any ASAP file you have to set up a WHAM model for debugging
testinput <- prepare_wham_input(testdata)
testWHAM <- fit_wham(testinput, MakeADFun.silent = TRUE, do.osa = FALSE, do.retro=FALSE)

OM <- testWHAM
testinput1 <- prepare_wham_input(testdata, model_name = "test1")
testinput2 <- prepare_wham_input(testdata, model_name="test2")
testinput3 <- prepare_wham_input(testdata, model_name="test3")
inputEMlist <- list(testinput1, testinput2, testinput3)

test <- simTestWHAM(nsim = 2, OM = OM, inputEMlist = inputEMlist)
```