### 2023-09-10 Ecov_study/catchability/Setup_OMs.Rmd

---
title: "Setup_OMs"
output: html_document
date: "2023-02-14"
editor_options: 
  markdown: 
    wrap: 72
---

Generic OM & EM settings that are the same as project 0:

-   Cod-like life history

-   B-H stock-recruit (for now)

-   1 fleet

-   2 indices (spring and fall)

-   Logistic selectivity for the fleet and all indices

-   Initial M estimate = 0.2

-   Logistic normal age composition for fleet and all indices

Project-specific OM settings that are not variable:

-   Ecov ar1 process on catchability with mean = 0

OM settings that are variable:

-   Ecov standard deviation = 0.1 or 0.5

-   Ecov process correlation = 0 or 0.5

-   Ecov process observation standard deviation = 0.1 or 0.5

-   Ecov effect on q = 0, 0.25, or 0.5 !!! I think this is the beta
    parameter but should confirm

-   Fishing history = F_MSY for 40 year history OR fish 2.5\*F_MSY for
    20 years than fish at F_MSY for remaining 20 - NOT CURRENTLY IMPLEMENTED, IN PROGRESS

-   Seasonality of Ecov effect on catchability (neither, one, or both seasons impacted) - NOT CURRENTLY IMPLEMENTED, IN PROGRESS

EM settings that are variable:

-   Ecov season correctly specified = BOTH (spring & fall correct), ONE
    (only one season correct), NONE (effect on q not included when it
    should be)

-   Ecov link/random effect setting = NoEnv, NoEnvRand, WithEnvRand,
    WithEnv

-   Seasonality of Ecov effect on catchability

Run analysis in 3 pieces:
1) Factorial runs with different ecov process magnitude, fishing history and observation error
-   maybe don't vary obs/process NAA since we are testing ability to
    recover catchability effect, could do sensitivity run with survival
    random effects and pick one of these settings
```{r}
##### OM
# Ecov process & effect magnitude set up
Ecov_process_sig <- c(0.1, 0.5)
Ecov_process_cor <- c(0, 0.5)
Ecov_process_obs_sig <- c(0.1, 0.5)
Ecov_effect <- c(0, 0.25, 0.5) # beta

# Fishing history
F_hist <- c("H-MSY", "MSY") # High-then MSY vs. MSY for entire history

# Observation error
ageComp_sig <- c(0.3, 1.5) # Fleet and index age comp treated with same error
log_index_sig <- c(0.1, 0.4) # agg_index_sigma
log_catch_sig <- 0.1 # agg_catch_sigma

# Factorial combinations
OMsetup <- expand.grid(Eco_process_sig = Ecov_process_sig, Ecov_process_cor = Ecov_process_cor, Ecov_process_obs_sig = Ecov_process_obs_sig, Ecov_effect = Ecov_effect, F_hist = F_hist, ageComp_sig = ageComp_sig, logIndex_sig = logIndex_sig)
OMsetup
```

2) Pick subset of OM settings and vary seasonality of q effect in EM
```{r}
##### EM
# EM factorial settings
miss_season <- c("BOTH", "ONE", "NONE") # Number of seasons with correctly specified effect, assumes 2 survey indices (spring & fall)
miss_q <- c("NoEnv", "NoEnvRand", "WithEnvRand", "WithEnv") # Catchability setup for EM
expand.grid(miss_season = miss_season, miss_q = miss_q)
```

3) May run sensitivity run with no ecov effect in OM but compare same EMs
```{r}
# Catchability set up
q_Ecov <- c("includeEcov", "excludeEcov") # May want to include these as a sensitivity run to compare what happens if OM has no ecov effect but EM assumes there is one (ecov process setup won't matter since fitting to data without effect specified in model)

##### OM sensitivity runs
# Factorial combinations when no ecov effect in OM
expand.grid(q_Ecov = q_Ecov, F_hist = F_hist, ageComp_sig = ageComp_sig, logIndex_sig = logIndex_sig) # probably do a subset, we have to decide what subset
```

# set up generic OM

```{r}
library(wham)
# Get generic groundfish life history info
source(file.path(here::here(),"common_code", "make_basic_info.R"))
groundfish_info <- make_basic_info()

# Change timing for spring (0.25) and fall (0.75) indices
groundfish_info$fracyr_indices[,1] <- 0.25
groundfish_info$fracyr_indices[,2] <- 0.75

# Assume logistic selectivity for the fleet and all indices
sel_list <- list(model = c(rep("logistic", groundfish_info$n_fleets),rep("logistic", groundfish_info$n_indices)),
                   initial_pars = lapply(1:(groundfish_info$n_fleets + groundfish_info$n_indices), function(x) c(5,1)),
                 n_selblocks = (groundfish_info$n_fleets + groundfish_info$n_indices))

# Starting mean M estimate at 0.2, estimate constant M for all ages/years
M <-  list(initial_means = rep(0.2, length(groundfish_info$ages)))

genericInput <- prepare_wham_input(basic_info = groundfish_info, 
                                   selectivity = sel_list, 
                                   M = M)

# genericFit <- fit_wham(genericInput, MakeADFun.silent = TRUE, do.retro = FALSE, do.osa = FALSE)

```

# Use common_code set_ecov function to adjust based on OM settings above

This chunk of code uses common code functions & internal WHAM functions
to manipulate an OM input object to initialize the OM rather than doing
all setup when creating the input as is typically the procedure when
fitting operational WHAM models.

```{r}
source(file.path(here::here(),"common_code", "set_ecov.R"))

# Loop over rows in OMsetup to initialize OM and EMs
for(iom in 1:nrow(OMsetup)){
  # Pull generic input settings
  input <- genericInput

  # Ecov setup
    Ecov <- list(label = "Ecov",
                 mean = matrix(rep(0,length(groundfish_info$years)), ncol =1), # Mean = 0
                 logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), length(groundfish_info$years), ncol=1),
                 years = groundfish_info$years, 
                 use_obs = matrix(rep(TRUE, length(groundfish_info$years)), ncol = 1),
                 lag = 0, 
                 where = "q", # Where/how/indices settings need to change if we do sensitivity runs
                 how = 1,
                 indices = list(1,2),
                 process_model = "ar1", # End generic inputs for ecov
                 process_mean_vals = rep(0,length(groundfish_info$years)), # Mean = 0
                 process_sig_vals = rep(OMsetup$Eco_process_sig[iom]),
                 process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                 beta_vals = list(rep(list(matrix(rep(OMsetup$Ecov_effect[iom], length(groundfish_info$ages)), nrow = 1)), 4))) #!!! Why 4 rows?

  input <- set_ecov(input = input, ecov = Ecov)

#   # Fishing history
#   # input =
#           set_F_scenario(input, Fhist, Fmsy = Fmsy, max_mult = max_mult_Fmsy, min_mult= min_mult_Fmsy,
#     change_time = F_change_time) # Function pulled from get_FMSY.R
#   
# F_hist <- c("H-MSY", "MSY") # High-then MSY vs. MSY for entire history Fmsy is an option

# Observation error - set L-N SD parameters for catch and index age comps
input$par$catch_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
input$par$index_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
input$data$agg_catch_sigma[] <- OMsetup[iom,"log_catch_sig"]
input$data$agg_index_sigma[] <- OMsetup[iom,"log_index_sig"]

}

```