### 2023-09-21 Ecov_study/catchability/Setup_OMs.Rmd
Update how F history initialized, leverage common code functions to set up OM/EMs.

---
title: "Setup_OMs"
output: html_document
date: "2023-02-14"
editor_options: 
  markdown: 
    wrap: 72
---

Generic OM & EM settings that are the same as project 0:

-   Cod-like life history

-   B-H stock-recruit (for now)

-   1 fleet

-   2 indices (spring and fall)

-   Logistic selectivity for the fleet and all indices

-   Initial M estimate = 0.2

-   Logistic normal age composition for fleet and all indices

Project-specific OM settings that are not variable:

-   Ecov ar1 process on catchability with mean = 0

OM settings that are variable:

-   Ecov standard deviation = 0.1 or 0.5

-   Ecov process correlation = 0 or 0.5

-   Ecov process observation standard deviation = 0.1 or 0.5

-   Ecov effect on q = 0, 0.25, or 0.5 !!! I think this is the beta
    parameter but should confirm

-   Fishing history = F_MSY for 40 year history OR fish 2.5\*F_MSY for
    20 years than fish at F_MSY for remaining 20 - NOT CURRENTLY IMPLEMENTED, IN PROGRESS

-   Seasonality of Ecov effect on catchability (neither, one, or both seasons impacted) - NOT CURRENTLY IMPLEMENTED, IN PROGRESS

EM settings that are variable:

-   Ecov season correctly specified = BOTH (spring & fall correct), ONE
    (only one season correct), NONE (effect on q not included when it
    should be)

-   Ecov link/random effect setting = NoEnv, NoEnvRand, WithEnvRand,
    WithEnv

-   Seasonality of Ecov effect on catchability

Run analysis in 3 pieces:
# 1) Factorial runs with different ecov process magnitude, fishing history and observation error
-   maybe don't vary obs/process NAA since we are testing ability to
    recover catchability effect, could do sensitivity run with survival
    random effects and pick one of these settings
```{r}
##### OM
# Ecov process & effect magnitude set up
Ecov_process_sig <- c(0.1, 0.5)
Ecov_process_cor <- c(0, 0.5)
Ecov_process_obs_sig <- c(0.1, 0.5)
Ecov_effect <- c(0, 0.25, 0.5) # beta

# Fishing history
F_hist <- c("H-L", "Fmsy") # High-then FMSY vs. FMSY for entire history

# Observation error
ageComp_sig <- c(0.3, 1.5) # Fleet and index age comp treated with same error
log_index_sig <- c(0.1, 0.4) # agg_index_sigma
log_catch_sig <- 0.1 # agg_catch_sigma

# Factorial combinations
OMsetup <- expand.grid(Ecov_process_sig = Ecov_process_sig, Ecov_process_cor = Ecov_process_cor, Ecov_process_obs_sig = Ecov_process_obs_sig, Ecov_effect = Ecov_effect, F_hist = F_hist, ageComp_sig = ageComp_sig, log_index_sig = log_index_sig, log_catch_sig = log_catch_sig)
OMsetup
```

# 2) Vary seasonality of q effect in EM
```{r}
##### EM
# EM factorial settings
miss_season <- c("BOTH", "ONE", "NONE") # Number of seasons with correctly specified effect, assumes 2 survey indices (spring & fall)
miss_q <- c("NoEcov", "Ecov", "qRand", "qRandEcov") # Catchability setup for EM
EMsetup <- expand.grid(miss_season = miss_season, miss_q = miss_q)
EMsetup 
```

# 3) May run sensitivity run with no ecov effect in OM but compare same EMs
Check that ecov effect = 0 is essentially turning off ecov
```{r}
# Catchability set up
q_Ecov <- c("includeEcov", "excludeEcov") # May want to include these as a sensitivity run to compare what happens if OM has no ecov effect but EM assumes there is one (ecov process setup won't matter since fitting to data without effect specified in model)

##### OM sensitivity runs
# Factorial combinations when no ecov effect in OM
expand.grid(q_Ecov = q_Ecov, F_hist = F_hist, ageComp_sig = ageComp_sig, log_index_sig = log_index_sig) # probably do a subset, we have to decide what subset
```

## Set up OM and corresponding EMs
This chunk of code uses common code functions & internal WHAM functions
to manipulate an OM input object to initialize the OM rather than doing
all setup when creating the input as is typically the procedure when
fitting operational WHAM models.

Use project 0 make_om() to set up generic OM input then add additional ecov settings afterwards before fitting model
- By default catchability impacted by environment only for fall (2nd) index
- Each OM stored in a corresponding folder/directory

EMs are misspecified for each OM and stored a sub directory of the OM
```{r}
library(TAF)
library(tidyverse)
library(wham)
# source(here::here("Project_0", "code", "make_om.R"))
source(here::here("make_om.R")) # Use q copy which requires that Beverton-Holt S-R used (make_om.R line #11)
source(here::here("make_basic_info.R")) # from common_code
source(here::here("set_ecov.R")) # from common_code
source(here::here("set_NAA.R")) # from common code
source(here::here("set_q.R")) # from common code
source(here::here("set_selectivity.R")) # from common code
source(here::here("set_M.R")) # from common code
source(here::here("get_FMSY.R"))

n_fleets <- 1
n_indices <- 2
n_ages <- 10
n_years <- 40
# Assume logistic selectivity for the fleet and all indices
sel_list <- list(model = c(rep("logistic", n_fleets),rep("logistic", n_indices)),
                 initial_pars = lapply(1:(n_fleets + n_indices), function(x) c(5,1)),
                 n_selblocks = (n_fleets + n_indices))

# Starting mean M estimate at 0.2, estimate constant M for all ages/years
M_list <-  list(initial_means = rep(0.2, n_ages))

# NAA_re = list(recruit_model = 3) # This required for Beverton-Holt S-R

for(iom in 1:nrow(OMsetup)){
  print(iom)
  ##### Set up generic input #####
  # Set up input without ecov
  input <- make_om(Fhist = OMsetup[iom,"F_hist"],
                   N1_state = "Fmsy", # Default, could also pick "overfished" or "unfished"
                   selectivity = sel_list,
                   M = M_list, 
                   catchability = NULL,
                   NAA_re = NULL) 
  # age_comp, # Default = "logistic-normal-miss0"
  # brp_year, # Default = 1
  # eq_F_init, # Default = 0.3
  # om_input = TRUE, # Don't fit model, only structure generic input
  # max_mult_Fmsy, # Default = 2.5
  # min_mult_Fmsy, # Default = 1
  # F_change_time # Default = 0.5, sets when F changes (in H-L or L-H F_hist scenarios)

  # Observation error - set L-N SD parameters for catch and index age comps # pulled from project 0 q_om_setup.R
  input$par$catch_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
  input$par$index_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
  input$data$agg_catch_sigma[] <- OMsetup[iom,"log_catch_sig"]
  input$data$agg_index_sigma[] <- OMsetup[iom,"log_index_sig"]
  input$data$catch_Neff[] <- 1 # Change Neff so scalar doesn't affect L-N SD
  input$data$index_Neff[] <- 1

  ##### Set up OM ecov #####
  # Ecov setup
  Ecov <- list(label = "Ecov",
               mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
               logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
               years = input$years, 
               use_obs = matrix(rep(TRUE, n_years), ncol = 1),
               lag = 0, 
               where = "q", # Where/how/indices settings need to change if we do sensitivity runs
               how = 1,
               indices = list(2), # Ecov only impact fall index catchability
               process_model = "ar1", # End generic inputs for ecov
               process_mean_vals = rep(0,n_years), # Mean = 0
               process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
               process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
               beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                       rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

  inputOM <- set_ecov(input = input, ecov = Ecov) #!!! something wrong with ecov$beta_vals[[j]][[n]] indexing

  # Finish setup & save OM
  OM <- NULL # Clear prior OM object just in case
  OM <- fit_wham(input = inputOM, do.fit = FALSE, MakeADFun.silent = TRUE)
  mkdir(here::here("Results", paste0("OM_", iom))) # set up storage
  saveRDS(OM, file = here::here("Results", paste0("OM_", iom), paste0("OM_", iom, ".Rds"))) # save OM
  saveRDS(OMsetup[iom,], file = here::here("Results", paste0("OM_", iom), "OMsettings.Rds")) # save OM settings in a separate file since naming is based on arbitraty order in OMsetup table

  ##### Iterate over EM misspecifications for this OM
  for(iem in 1:nrow(EMsetup)){
    inputEM <- NULL # Reset at start of loop just in case

    # Seasonal misspecification (q random effect implemented with same misspecification as ecov impact)
    if(EMsetup[iem, "miss_season"] == "BOTH"){
      index_list <- list(1) # Both seasons misspecified
      qRand <- c("ar1", "none")
    } else if(EMsetup[iem, "miss_season"] == "ONE"){
      index_list <- list(1,2) # Spring misspecified, fall correct
      qRand <- c("ar1", "ar1")
    } else{
      index_list <- list(2) # Both seasons correctly specified
      qRand <- c("none", "ar1")
    }

    # Catchability model assumptions
    if(EMsetup[iem, "miss_q"] == "NoEcov"){ # No ecov impact on q or q random effect
      print("NoEcov")
      Ecov <- list(label = "NoEcov",
                   mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
                   logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
                   years = input$years, 
                   use_obs = matrix(rep(TRUE, n_years), ncol = 1),
                   lag = 0, 
                   where = "none", # Where/how/indices settings need to change if we do sensitivity runs
                   how = 1,
                   indices = index_list, # Ecov impact based on sesonal misspecification setting
                   process_model = "ar1", # End generic inputs for ecov
                   process_mean_vals = rep(0,n_years), # Mean = 0
                   process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
                   process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                   beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                           rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

      inputEM <- set_ecov(input = input, ecov = Ecov)

    } else if(EMsetup[iem, "miss_q"] == "Ecov"){ # Ecov impact on q, correctly specified option when seasonal impact correctly specified
      print("Ecov")
      Ecov <- list(label = "Ecov",
                   mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
                   logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
                   years = input$years, 
                   use_obs = matrix(rep(TRUE, n_years), ncol = 1),
                   lag = 0, 
                   where = "q", # Where/how/indices settings need to change if we do sensitivity runs
                   how = 1,
                   indices = index_list, # Ecov impact based on sesonal misspecification setting
                   process_model = "ar1", # End generic inputs for ecov
                   process_mean_vals = rep(0,n_years), # Mean = 0
                   process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
                   process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                   beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                           rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

      inputEM <- set_ecov(input = input, ecov = Ecov)

    } else if(EMsetup[iem, "miss_q"] == "qRand"){ # q random effect
      print("qRand")
      Ecov <- list(label = "qRand",
                   mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
                   logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
                   years = input$years, 
                   use_obs = matrix(rep(TRUE, n_years), ncol = 1),
                   lag = 0, 
                   where = "none", # Where/how/indices settings need to change if we do sensitivity runs
                   how = 1,
                   indices = index_list, # Ecov impact based on sesonal misspecification setting
                   process_model = "ar1", # End generic inputs for ecov
                   process_mean_vals = rep(0,n_years), # Mean = 0
                   process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
                   process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                   beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                           rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

      inputEM <- set_ecov(input = input, ecov = Ecov)

      # Catchability random effects
      inputEM <- set_q(input = inputEM, q_opts = qRand)

    } else{ # qRandEcov both q random effect and ecov impact on q
      print("qRandEcov")
      Ecov <- list(label = "qRandEcov",
                   mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
                   logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
                   years = input$years, 
                   use_obs = matrix(rep(TRUE, n_years), ncol = 1),
                   lag = 0, 
                   where = "q", # Where/how/indices settings need to change if we do sensitivity runs
                   how = 1,
                   indices = index_list, # Ecov impact based on sesonal misspecification setting
                   process_model = "ar1", # End generic inputs for ecov
                   process_mean_vals = rep(0,n_years), # Mean = 0
                   process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
                   process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                   beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                           rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

      inputEM <- set_ecov(input = input, ecov = Ecov)

      # Catchability random effects
      inputEM <- set_q(input = inputEM, q_opts = qRand)
    }

    # Save EM specification
    mkdir(here::here("Results", paste0("OM_", iom), paste0("EM_missSeason_", EMsetup[iem, "miss_season"], "_missQ_", EMsetup[iem, "miss_q"]))) # set up storage

    saveRDS(inputEM, file = here::here("Results", paste0("OM_", iom), paste0("EM_missSeason_", EMsetup[iem, "miss_season"], "_missQ_", EMsetup[iem, "miss_q"]), "EMinput.Rds")) # save EM input

  } # End loop over EM specifications

} # End loop over OMs

```

## Use common_code set_ecov function to adjust generic OM based on factorial OM settings above

```{r}
# # source(file.path(here::here(),"common_code", "set_ecov.R"))
# source(here::here("set_ecov.R"))
# source(here::here("get_FMSY.R")) # Contains set_F_scenario function
# 
# 
# # Loop over rows in OMsetup to initialize OM and EMs
# for(iom in 1:nrow(OMsetup)){
#   # Pull generic groundfish settings
#   input <- NULL # Clear info from prior iteration of loop
#   input <- genericInput
#   
#   # Ecov setup
#     Ecov <- list(label = "Ecov",
#                  mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
#                  logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
#                  years = input$years, 
#                  use_obs = matrix(rep(TRUE, n_years), ncol = 1),
#                  lag = 0, 
#                  where = "q", # Where/how/indices settings need to change if we do sensitivity runs
#                  how = 1,
#                  indices = list(1,2),
#                  process_model = "ar1", # End generic inputs for ecov
#                  process_mean_vals = rep(0,n_years), # Mean = 0
#                  process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
#                  process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
#                  beta_vals = list(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
#                                  rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2))) # impact on q by index (here 2 indices)
#     
#   input <- set_ecov(input = input, ecov = Ecov)
#   
#   
#   # Fishing history
#     # Set F40
#     temp <- fit_wham(input, do.fit = FALSE, MakeADFun.silent = TRUE)
#     F40 <- exp(temp$rep$log_FXSPR[1]) %>% round(.,digits=3)
#     Fmsy <- exp(temp$rep$log_FMSY[1]) %>% round(.,digits=3)
#     # Get steepness so that Fmsy = F40
#     steepness <- get_steepness(steepness_ini = 0.75, input = input, NAA_re = NAA_re, F40 = F40, brp_year = 1) # check brp_year
#     NAA_re$recruit_pars[1] <- steepness$par
#     input = set_NAA(input, NAA_re)
#     temp <- fit_wham(input, do.fit = FALSE, MakeADFun.silent = TRUE)
#     Fmsy <- exp(temp$rep$log_FMSY[1]) %>% round(.,digits=3)
#     # Initial NAA according to equilibium conditions
#     eq_F_N1 <- exp(temp$rep$log_FMSY[1]) # = F40
#     
#   
#   
#   input <- set_F_scenario(input = input, 
#                           Fhist = OMsetup[iom, "F_hist"], 
#                           Fmsy = Fmsy, 
#                           max_mult = max_mult_Fmsy, 
#                           min_mult= min_mult_Fmsy,
#                           change_time = F_change_time) # Function pulled from get_FMSY.R
# 
# F_hist <- c("H-MSY", "MSY") # High-then MSY vs. MSY for entire history Fmsy is an option
# 
# # Observation error - set L-N SD parameters for catch and index age comps # pulled from project 0 q_om_setup.R
# input$par$catch_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
# input$par$index_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
# input$data$agg_catch_sigma[] <- OMsetup[iom,"log_catch_sig"]
# input$data$agg_index_sigma[] <- OMsetup[iom,"log_index_sig"]
# input$data$catch_Neff[] <- 1 # Change Neff so scalar doesn't affect L-N SD
# input$data$index_Neff[] <- 1
# 
# 
# }
# 
# 
#

```