### 2023-09-22 Ecov_study/catchability/Catchability_ecov_sims.Rmd
Update simulations.

---
title: "Catchability_ecov_sims"
output: html_document
date: "2022-11-14"
editor_options: 
  markdown: 
    wrap: 72
---

Load packages & source functions used in simulation testing
```{r}
## Packages
library(tidyverse)
library(wham)

## OM/EM setup function
# source(here::here("Project_0", "code", "make_om.R"))
source(here::here("make_om.R")) # From Project_0/code/make_om.R
# source(file.path(here::here(),"common_code", "make_basic_info.R"))
source(here::here("make_basic_info.R"))

## Simulation testing functions
# source(here::here("Ecov_study","catchability","simOM.R"))
source(here::here("Other_Research", "simOM.R"))
# source(here::here("Ecov_study","catchability","simTestWHAM.R"))
source(here::here("Other_Research", "simTestWHAM.R"))

```

### Visual describing results structure

WHAM results objects only displayed for OM simulation 1, similar
structure for other simulations and for estimation models.

```{r}
library(DiagrammeR)

grViz("

digraph boxes_and_circles{
## Add node statements
# Define shape
node[shape = box]
OM[label = 'OM \n results$OM']
sim1[label = 'Simulation 1 results \n results$OM[1][[1]]']
sim2[label = 'Simulation 2 results \n results$OM[2][[1]]']
sim3[label = 'Simulation 3 results \n results$OM[3][[1]]']

EM1[label = 'EM1 \n results$EM1']
sim1_1[label = 'Simulation 1 results \n results[EMs[1]][[1]][1][[1]]']
sim1_2[label = 'Simulation 2 results \n results[EMs[1]][[1]][2][[1]]']
sim1_3[label = 'Simulation 3 results \n results[EMs[1]][[1]][2][[1]]']
EM2[label = 'EM2 \n results$EM2']
sim2_1[label = 'Simulation 1 results \n results[EMs[2]][[1]][1][[1]]']
sim2_2[label = 'Simulation 2 results \n results[EMs[2]][[1]][2][[1]]']
sim2_3[label = 'Simulation 3 results \n results[EMs[2]][[1]][3][[1]]']
ageslab[label = 'ages.lab']

# Add edge statements to connect nodes
edge[] # optional arrowhead options
# Connect nodes (may be labeled - appears on line)
results -> OM; OM -> sim1; OM -> sim2; OM -> sim3;
results -> EM1; EM1 -> sim1_1; EM1 -> sim1_2; EM1 -> sim1_3;
results -> EM2; EM2 -> sim2_1; EM2 -> sim2_2; EM2 -> sim2_3;
sim1 -> par;  
sim1 -> fn;
sim1 -> gr;
sim1 -> he;
sim1 -> hessian;
sim1 -> method;
sim1 -> retape;
sim1 -> env;
sim1 -> report;
sim1 -> simulate;
sim1 -> years;
sim1 -> years_full;
sim1 -> ageslab;
sim1 -> model_name;
sim1 -> input;
sim1 -> wham_version;
sim1 -> opt;
sim1 -> date;
sim1 -> dir;
sim1 -> rep;
sim1 -> TMB_version;
sim1 -> parList;
sim1 -> final_gradient;
sim1 -> is_sdrep;
sim1 -> na_sdrep;
sim1 -> runtime;
sim1 -> sdrep;

# Add graph statement
graph[nodesep = 0.1, rankdir = LR] # set distance between nodes # for left-right orientation use , rankdir = LR

}   
")

```

# Specifics for environmentally-linked catchability WHAM runs

## Set up operating and estimation models
This need only be done once by running Setup_OMs.Rmd !!! Need to move this code here so OMsetup and EMsetup available below for processing

## Simulation testing
```{r}
simTestWHAM(nsim = 5, OM = OM1, inputEMlist = list(EMq_env, EMq_rand_env, EMq_rand, EMq)) # Update once OM/EM specifics finalized!!!!

```

Run tests only for EMs with no seasonal misspecifications (only catchability misspecification misspecified)
```{r}
subsetEM <- EMsetup %>% as.data.frame() %>% filter(miss_season == "NONE")

for(iom in 1:192){
  testOM <- readRDS(here::here("Results", paste0("OM_", iom, ".Rds")))

  for(iem in 1:4){
    testEM <- readRDS(here::here("Results", paste0("OM_", iom), paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"]), "EMinput.Rds"))
  }
}
```

Run OMs with no ecov effect (NoEcov) and with ecov effect (Ecov) are equivalent here and the qRand and qRandEcov are also equivalent so only run the NoEcov and qRandEcov options
```{r}
subsetOM <- OMsetup %>% filter(Ecov_effect == 0)
# EMs 
subsetEM <- Msetup %>% filter(miss_season == "NONE") %>% filter(miss_q == "qRand" | miss_q == "NoEcov")
```

## Post-process simulation results

```{r}
postprocess_simTestWHAM(filenames = c("filename1.Rds", "filename2.Rds", "ect.Rds")) # !!! Update once completed runs
```

!!!! post-process based on storage & how sims run - could run sims based
varying isim - could also loop over OM but this could be computationally
expensive - I recommend labeling runs with OM and EM names (and setting
these up with informative) so it is easy to pull out results & compare
based on model setup - pull from mod\$model_name to help aggregate data

proposed function with: - fitted OM model object - list of EM inputs

### Result formatted as list of lists, save as rdata object in specified directory with timestamp in name

run1_11_14_22.Rda\<-result - EM1 - EM2 - EM3 - EM4 - OM result

# Other things to consider

-   Other EMs can be developed (e.g. with alternative ecov process
    assumptions) but should be given DIFFERENT model_names to help with
    post-processing

# Short script for function testing purposes only!!!

```{r}
library(wham)
# For testing function
testdata <- read_asap3_dat(paste("/Users/amandahart/Research/Plaice_Methods_Development", "data", "PlaiceWHAM-2019_revised_NEFSC-LW-WAA_splitNEFSC-BigUnit_fixSeason.DAT", sep="/")) # this can be swapped out for any ASAP file you have to set up a WHAM model for debugging
testinput <- prepare_wham_input(testdata)
testWHAM <- fit_wham(testinput, MakeADFun.silent = TRUE, do.osa = FALSE, do.retro=FALSE)

OM <- testWHAM
testinput1 <- prepare_wham_input(testdata, model_name = "test1")
testinput2 <- prepare_wham_input(testdata, model_name="test2")
testinput3 <- prepare_wham_input(testdata, model_name="test3")
inputEMlist <- list(testinput1, testinput2, testinput3)

test <- simTestWHAM(nsim = 2, OM = OM, inputEMlist = inputEMlist)
```

# To do next

- Finish OM setup
  - change how ecov passed in
  - set up range of uncertainty values to explore
  - tim sets equilibrium F (eq_F_init)
    - setup OM without fit & use model to find F40 proxy
  - Tim also sets intitial numbers at age based on equilibrium (overfished F=eqFN1*max_mult_Fmsy or not eqFN1=0)
- Tim lines 71-79 pf make_om use internal processing for prepare_wham_input to plug in setup from M_Ecov_om_setup.R
- Update what is saved as output
  - calculate performance metrics
- Run simulations, hpcc?
- Separate sims and function scripts

# 9/18/23
1) do OM & EM setup (EMs specified relative to each OM)
```{r}
Setup_OMs()
```
2) simulation test EMs for each OM
```{r}
simTestWHAM()
```
3) Post-process results
```{r}
postprocess_simTestWHAM()
```