### 2023-09-26 Ecov_study/catchability/Catchability_ecov_sims.Rmd
Add model names and update sim test code.

---
title: "Catchability_ecov_sims"
output: html_document
date: "2022-11-14"
editor_options: 
  markdown: 
    wrap: 72
---

Load packages & source functions used in simulation testing
```{r}
## Packages
library(tidyverse)
library(wham)
library(TAF)

## OM/EM setup functions
source(here::here("Ecov_study", "catchability", "make_om.R")) # Use revised copy that has option not to include a S-R relationship
source(file.path(here::here(),"common_code", "make_basic_info.R")) 
source(here::here("common_code", "set_ecov.R")) 
source(here::here("common_code", "set_NAA.R")) 
source(here::here("common_code", "set_q.R"))
source(here::here("common_code", "set_selectivity.R")) 
source(here::here("common_code", "set_M.R")) 
source(here::here("common_code", "get_FMSY.R"))

## Simulation testing functions
source(here::here("Ecov_study","catchability","simOM.R"))
#source(here::here("Other_Research", "simOM.R"))
source(here::here("Ecov_study","catchability","simTestWHAM.R"))
#source(here::here("Other_Research", "simTestWHAM.R"))
```

# Visual describing results structure

WHAM results objects only displayed for OM simulation 1, similar
structure for other simulations and for estimation models.

```{r}
library(DiagrammeR)

grViz("

digraph boxes_and_circles{
## Add node statements
# Define shape
node[shape = box]
OM[label = 'OM \n results$OM']
sim1[label = 'Simulation 1 results \n results$OM[1][[1]]']
sim2[label = 'Simulation 2 results \n results$OM[2][[1]]']
sim3[label = 'Simulation 3 results \n results$OM[3][[1]]']

EM1[label = 'EM1 \n results$EM1']
sim1_1[label = 'Simulation 1 results \n results[EMs[1]][[1]][1][[1]]']
sim1_2[label = 'Simulation 2 results \n results[EMs[1]][[1]][2][[1]]']
sim1_3[label = 'Simulation 3 results \n results[EMs[1]][[1]][2][[1]]']
EM2[label = 'EM2 \n results$EM2']
sim2_1[label = 'Simulation 1 results \n results[EMs[2]][[1]][1][[1]]']
sim2_2[label = 'Simulation 2 results \n results[EMs[2]][[1]][2][[1]]']
sim2_3[label = 'Simulation 3 results \n results[EMs[2]][[1]][3][[1]]']
ageslab[label = 'ages.lab']

# Add edge statements to connect nodes
edge[] # optional arrowhead options
# Connect nodes (may be labeled - appears on line)
results -> OM; OM -> sim1; OM -> sim2; OM -> sim3;
results -> EM1; EM1 -> sim1_1; EM1 -> sim1_2; EM1 -> sim1_3;
results -> EM2; EM2 -> sim2_1; EM2 -> sim2_2; EM2 -> sim2_3;
sim1 -> par;  
sim1 -> fn;
sim1 -> gr;
sim1 -> he;
sim1 -> hessian;
sim1 -> method;
sim1 -> retape;
sim1 -> env;
sim1 -> report;
sim1 -> simulate;
sim1 -> years;
sim1 -> years_full;
sim1 -> ageslab;
sim1 -> model_name;
sim1 -> input;
sim1 -> wham_version;
sim1 -> opt;
sim1 -> date;
sim1 -> dir;
sim1 -> rep;
sim1 -> TMB_version;
sim1 -> parList;
sim1 -> final_gradient;
sim1 -> is_sdrep;
sim1 -> na_sdrep;
sim1 -> runtime;
sim1 -> sdrep;

# Add graph statement
graph[nodesep = 0.1, rankdir = LR] # set distance between nodes # for left-right orientation use , rankdir = LR

}   
")

```

# Specifics for environmentally-linked catchability WHAM runs

## Set up operating and estimation models
### Factorial OM settings
Vary ecov process magnitude, fishing history and observation error
```{r}
##### OM
# Ecov process & effect magnitude set up
Ecov_process_sig <- c(0.1, 0.5)
Ecov_process_cor <- c(0, 0.5)
Ecov_process_obs_sig <- c(0.1, 0.5)
Ecov_effect <- c(0, 0.25, 0.5) # beta

# Fishing history
F_hist <- c("H-L", "Fmsy") # High-then FMSY vs. FMSY for entire history

# Observation error
ageComp_sig <- c(0.3, 1.5) # Fleet and index age comp treated with same error
log_index_sig <- c(0.1, 0.4) # agg_index_sigma
log_catch_sig <- 0.1 # agg_catch_sigma

# Factorial combinations
OMsetup <- expand.grid(Ecov_process_sig = Ecov_process_sig, Ecov_process_cor = Ecov_process_cor, Ecov_process_obs_sig = Ecov_process_obs_sig, Ecov_effect = Ecov_effect, F_hist = F_hist, ageComp_sig = ageComp_sig, log_index_sig = log_index_sig, log_catch_sig = log_catch_sig)

OMname <- 1:nrow(OMsetup)

OMsetup <- cbind(OMname, OMsetup)

OMsetup[,"F_hist"] <- as.character(OMsetup[,"F_hist"])

OMsetup
```
### Factorial EM settings
```{r}
##### EM
# EM factorial settings
miss_season <- c("BOTH", "ONE", "NONE") # Number of seasons with correctly specified effect, assumes 2 survey indices (spring & fall)
miss_q <- c("NoEcov", "Ecov", "qRand", "qRandEcov") # Catchability setup for EM
EMsetup <- expand.grid(miss_season = miss_season, miss_q = miss_q)
EMsetup 
```

### Set up OM and corresponding EMs based on above factorial variations (need only be run once)
- No stock-recruit relationship is implemented (as in project 0) so a variation of make_om() is used here to:
  - 1) Set Fmsy equal to the F40% proxy
  - 2) Estimate equilibirum initial numbers at age by setting NAA_re$N1_model = 1 rather than using the S-R
- By default, the environment only impacts catchability for the fall (2nd) index. EM misspecifications are based off of this assumption.
- Each OM is stored in a separate directory with input objects for EMs that are misspecified based on this OM in a subdirectory of the OM
```{r}
n_fleets <- 1
n_indices <- 2
n_ages <- 10
n_years <- 40

# Assume logistic selectivity for the fleet and all indices
sel_list <- list(model = c(rep("logistic", n_fleets),rep("logistic", n_indices)),
                 initial_pars = lapply(1:(n_fleets + n_indices), function(x) c(5,1)),
                 n_selblocks = (n_fleets + n_indices))

# Starting mean M estimate at 0.2, estimate constant M for all ages/years
M_list <-  list(initial_means = rep(0.2, n_ages))

# NAA_re = list(recruit_model = 3) # This required for Beverton-Holt S-R

for(iom in 1:nrow(OMsetup)){
  print(iom)
  ##### Set up generic input #####
  # Set up input without ecov
  input <- make_om(Fhist = OMsetup[iom,"F_hist"],
                   N1_state = "Fmsy", # Default, could also pick "overfished" or "unfished"
                   selectivity = sel_list,
                   M = M_list, 
                   catchability = NULL,
                   NAA_re = NULL) 
  # age_comp, # Default = "logistic-normal-miss0"
  # brp_year, # Default = 1
  # eq_F_init, # Default = 0.3
  # om_input = TRUE, # Don't fit model, only structure generic input
  # max_mult_Fmsy, # Default = 2.5
  # min_mult_Fmsy, # Default = 1
  # F_change_time # Default = 0.5, sets when F changes (in H-L or L-H F_hist scenarios)

  # Observation error - set L-N SD parameters for catch and index age comps # pulled from project 0 q_om_setup.R
  input$par$catch_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
  input$par$index_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
  input$data$agg_catch_sigma[] <- OMsetup[iom,"log_catch_sig"]
  input$data$agg_index_sigma[] <- OMsetup[iom,"log_index_sig"]
  input$data$catch_Neff[] <- 1 # Change Neff so scalar doesn't affect L-N SD
  input$data$index_Neff[] <- 1

  ##### Set up OM ecov #####
  # Ecov setup
  Ecov <- list(label = "Ecov",
               mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
               logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
               years = input$years, 
               use_obs = matrix(rep(TRUE, n_years), ncol = 1),
               lag = 0, 
               where = "q", # Where/how/indices settings need to change if we do sensitivity runs
               how = 1,
               indices = list(2), # Ecov only impact fall index catchability
               process_model = "ar1", # End generic inputs for ecov
               process_mean_vals = rep(0,n_years), # Mean = 0
               process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
               process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
               beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                       rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

  inputOM <- set_ecov(input = input, ecov = Ecov) #!!! something wrong with ecov$beta_vals[[j]][[n]] indexing

  # Finish setup & save OM
  inputOM$model_name <- paste0("OM_", paste(OMsetup[iom, ], collapse = "_")) # Set up OM name based on OMsetup 
  OM <- NULL # Clear prior OM object just in case
  OM <- fit_wham(input = inputOM, do.fit = FALSE, MakeADFun.silent = TRUE)
  mkdir(here::here("Ecov_study", "catchability", "Results", paste0("OM_", OMsetup$OMname[iom]))) # set up storage
  saveRDS(OM, file = here::here("Ecov_study", "catchability", "Results", paste0("OM_", OMsetup$OMname[iom]), paste0("OM_", OMsetup$OMname[iom], ".Rds"))) # save OM
  saveRDS(OMsetup[iom,], file = here::here("Ecov_study", "catchability", "Results", paste0("OM_", OMsetup$OMname[iom]), "OMsettings.Rds")) # save OM settings in a separate file since naming is based on arbitraty order in OMsetup table

  ##### Iterate over EM misspecifications for this OM
  for(iem in 1:nrow(EMsetup)){
    inputEM <- NULL # Reset at start of loop just in case

    # Seasonal misspecification (q random effect implemented with same misspecification as ecov impact)
    if(EMsetup[iem, "miss_season"] == "BOTH"){
      index_list <- list(1) # Both seasons misspecified
      qRand <- c("ar1", "none")
    } else if(EMsetup[iem, "miss_season"] == "ONE"){
      index_list <- list(1,2) # Spring misspecified, fall correct
      qRand <- c("ar1", "ar1")
    } else{
      index_list <- list(2) # Both seasons correctly specified
      qRand <- c("none", "ar1")
    }

    # Catchability model assumptions
    if(EMsetup[iem, "miss_q"] == "NoEcov"){ # No ecov impact on q or q random effect
      print("NoEcov")
      Ecov <- list(label = "NoEcov",
                   mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
                   logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
                   years = input$years, 
                   use_obs = matrix(rep(TRUE, n_years), ncol = 1),
                   lag = 0, 
                   where = "none", # Where/how/indices settings need to change if we do sensitivity runs
                   how = 1,
                   indices = index_list, # Ecov impact based on sesonal misspecification setting
                   process_model = "ar1", # End generic inputs for ecov
                   process_mean_vals = rep(0,n_years), # Mean = 0
                   process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
                   process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                   beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                           rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

      inputEM <- set_ecov(input = input, ecov = Ecov)

      # Set up EM name based on EMsetup
      input$model_name <- paste0("EM_", paste(unfactor(EMsetup[iem, ]), collapse = "_")) # Set up OM name based on OMsetup

    } else if(EMsetup[iem, "miss_q"] == "Ecov"){ # Ecov impact on q, correctly specified option when seasonal impact correctly specified
      print("Ecov")
      Ecov <- list(label = "Ecov",
                   mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
                   logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
                   years = input$years, 
                   use_obs = matrix(rep(TRUE, n_years), ncol = 1),
                   lag = 0, 
                   where = "q", # Where/how/indices settings need to change if we do sensitivity runs
                   how = 1,
                   indices = index_list, # Ecov impact based on sesonal misspecification setting
                   process_model = "ar1", # End generic inputs for ecov
                   process_mean_vals = rep(0,n_years), # Mean = 0
                   process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
                   process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                   beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                           rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

      inputEM <- set_ecov(input = input, ecov = Ecov)

      # Set up EM name based on EMsetup
      input$model_name <- paste0("EM_", paste(unfactor(EMsetup[iem, ]), collapse = "_")) # Set up OM name based on OMsetup

    } else if(EMsetup[iem, "miss_q"] == "qRand"){ # q random effect
      print("qRand")
      Ecov <- list(label = "qRand",
                   mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
                   logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
                   years = input$years, 
                   use_obs = matrix(rep(TRUE, n_years), ncol = 1),
                   lag = 0, 
                   where = "none", # Where/how/indices settings need to change if we do sensitivity runs
                   how = 1,
                   indices = index_list, # Ecov impact based on sesonal misspecification setting
                   process_model = "ar1", # End generic inputs for ecov
                   process_mean_vals = rep(0,n_years), # Mean = 0
                   process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
                   process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                   beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                           rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

      inputEM <- set_ecov(input = input, ecov = Ecov)

      # Catchability random effects
      inputEM <- set_q(input = inputEM, catchability = list(re = qRand))

      # Set up EM name based on EMsetup
      input$model_name <- paste0("EM_", paste(unfactor(EMsetup[iem, ]), collapse = "_")) # Set up OM name based on OMsetup

    } else{ # qRandEcov both q random effect and ecov impact on q
      print("qRandEcov")
      Ecov <- list(label = "qRandEcov",
                   mean = matrix(rep(0,n_years), ncol =1), # Mean = 0
                   logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), n_years, ncol=1),
                   years = input$years, 
                   use_obs = matrix(rep(TRUE, n_years), ncol = 1),
                   lag = 0, 
                   where = "q", # Where/how/indices settings need to change if we do sensitivity runs
                   how = 1,
                   indices = index_list, # Ecov impact based on sesonal misspecification setting
                   process_model = "ar1", # End generic inputs for ecov
                   process_mean_vals = rep(0,n_years), # Mean = 0
                   process_sig_vals = rep(OMsetup$Ecov_process_sig[iom]),
                   process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                   beta_vals = list(append(rep(list(matrix(rep(0, n_ages), nrow = 1)), 2), # No impact on R,M
                                           rep(list(matrix(rep(OMsetup$Ecov_effect[iom], n_ages), nrow = 1)), 2)))) # impact on q by index (here 2 indices)

      inputEM <- set_ecov(input = input, ecov = Ecov)

      # Catchability random effects
      inputEM <- set_q(input = inputEM, catchability = list(re = qRand))

      # Set up EM name based on EMsetup
      input$model_name <- paste0("EM_", paste(unfactor(EMsetup[iem, ]), collapse = "_")) # Set up OM name based on OMsetup 
    }

    # Save EM specification
    mkdir(here::here("Ecov_study", "catchability", "Results", paste0("OM_", iom), paste0("EM_missSeason_", EMsetup[iem, "miss_season"], "_missQ_", EMsetup[iem, "miss_q"]))) # set up storage

    saveRDS(inputEM, file = here::here("Ecov_study", "catchability", "Results", paste0("OM_", iom), paste0("EM_missSeason_", EMsetup[iem, "miss_season"], "_missQ_", EMsetup[iem, "miss_q"]), "EMinput.Rds")) # save EM input

  } # End loop over EM specifications

} # End loop over OMs
```

## Simulation testing
If OMsetup or EMsetup objects are found then source the "Factorial OM settings" and "Factorial EM settings" sections above. DO NOT rerun the actual setup (the names are used to ID what subset of the OMs/EMs are being tested).

### Test
```{r}
#simTestWHAM(nsim = 5, OM = OM1, inputEMlist = list(EMq_env, EMq_rand_env, EMq_rand, EMq)) # Update once OM/EM specifics finalized!!!!

# Set number of simulations to run
nsim <- 2

# Subset of OMs to run
subsetOM <- OMsetup %>% filter(F_hist == "H-L") %>% head(n=2)

# Subset of EMs to run
subsetEM <- EMsetup %>% as.data.frame() %>% filter(miss_season == "NONE")

# Run simulation tests
for(iom in 1:nrow(subsetOM)){
  omdir <- here::here("Ecov_study", "catchability", "Results", paste0("OM_", OMsetup[iom, "OMname"]))

  # Pull OM
  testOM <- readRDS(here::here(omdir, paste0("OM_", OMsetup[iom, "OMname"], ".Rds")))

  for(iem in 1:nrow(subsetEM)){
    # Pull EM
    testEM <- readRDS(here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"]), "EMinput.Rds"))

    # Run simulation test
    simTestWHAM(nsim = nsim,
                OM = testOM,
                inputEMlist = list(testEM), # Run one EM at a time
                outdir = here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"])))
  }
}
```

### Run tests only for EMs with no seasonal misspecifications (only catchability misspecification misspecified)
Goal = run 100 simulations for each OM
- postprocessing script will be set up to process multiple result files if these simulations need to be run in batches (i.e. reduce # of sims, # of OMs run, or # of EMs run at once)

#### Amanda run Fhist = H-L simulations
```{r}
# Set number of simulations to run
nsim <- 2

# Subset of OMs to run
subsetOM <- OMsetup %>% filter(F_hist == "H-L")

# Subset of EMs to run
subsetEM <- EMsetup %>% as.data.frame() %>% filter(miss_season == "NONE")

# Run simulation tests
for(iom in 1:nrow(subsetOM)){
  omdir <- here::here("Ecov_study", "catchability", "Results", paste0("OM_", OMsetup[iom, "OMname"]))

  # Pull OM
  testOM <- readRDS(here::here(omdir, paste0("OM_", OMsetup[iom, "OMname"], ".Rds")))

  for(iem in 1:nrow(subsetEM)){
    # Pull EM
    testEM <- readRDS(here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"]), "EMinput.Rds"))

    # Run simulation test
    simTestWHAM(nsim = nsim,
                OM = testOM,
                inputEMlist = list(testEM), # Run one EM at a time
                outdir = here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"])))
  }
}

```
#### Alex run F_hist = Fmsy simulations
This code loops over the rows in subsetOM and subsetEM. The easiest way to run smaller chunks is to change the number of rows in subsetOM (i.e. run fewer OMs at once) or run fewer simulations. Results are all time-stamped and postprocess_simTestWHAM() is set up to combine multiple files so this won't cause processing issues later on.
```{r}
# Set number of simulations to run
nsim <- 100

# Subset of OMs to run
subsetOM <- OMsetup %>% filter(F_hist == "Fmsy") %>% filter(Ecov_effect != 0)

# Subset of EMs to run
subsetEM <- EMsetup %>% as.data.frame() %>% filter(miss_season == "NONE")

# Run simulation tests
for(iom in 1:nrow(subsetOM)){
  omdir <- here::here("Ecov_study", "catchability", "Results", paste0("OM_", OMsetup[iom, "OMname"]))

  # Pull OM
  testOM <- readRDS(here::here(omdir, paste0("OM_", OMsetup[iom, "OMname"], ".Rds")))

  for(iem in 1:nrow(subsetEM)){
    # Pull EM
    testEM <- readRDS(here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"]), "EMinput.Rds"))

    # Run simulation test
    simTestWHAM(nsim = nsim,
                OM = testOM,
                inputEMlist = list(testEM), # Run one EM at a time
                outdir = here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"])))
  }
}

```

### Run OMs with no ecov effect 
EMs with no ecov effect (NoEcov) and with ecov effect (Ecov) are equivalent when the OM has no ecov effect and the qRand and qRandEcov are also equivalent so only run the NoEcov and qRandEcov EMs for these OMs
#### Amanda runs H-L Fhist
```{r}
# Set number of simulations to run
nsim <- 100

# Subset of OMs to run
subsetOM <- OMsetup %>% filter(F_hist == "H-L") %>% filter(Ecov_effect != 0)
# EMs 
subsetEM <- EMsetup %>% filter(miss_season == "NONE") %>% filter(miss_q == "qRand" | miss_q == "NoEcov")

# Run simulation tests
for(iom in 1:nrow(subsetOM)){
  omdir <- here::here("Ecov_study", "catchability", "Results", paste0("OM_", OMsetup[iom, "OMname"]))

  # Pull OM
  testOM <- readRDS(here::here(omdir, paste0("OM_", OMsetup[iom, "OMname"], ".Rds")))

  for(iem in 1:nrow(subsetEM)){
    # Pull EM
    testEM <- readRDS(here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"]), "EMinput.Rds"))

    # Run simulation test
    simTestWHAM(nsim = nsim,
                OM = testOM,
                inputEMlist = list(testEM), # Run one EM at a time
                outdir = here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"])))
  }
}
```
#### Alex runs Fmsy Fhist
```{r}
# Set number of simulations to run
nsim <- 100

# Subset of OMs to run
subsetOM <- OMsetup %>% filter(F_hist == "Fmsy") %>% filter(Ecov_effect != 0)
# EMs 
subsetEM <- EMsetup %>% filter(miss_season == "NONE") %>% filter(miss_q == "qRand" | miss_q == "NoEcov")

# Run simulation tests
for(iom in 1:nrow(subsetOM)){
  omdir <- here::here("Ecov_study", "catchability", "Results", paste0("OM_", OMsetup[iom, "OMname"]))

  # Pull OM
  testOM <- readRDS(here::here(omdir, paste0("OM_", OMsetup[iom, "OMname"], ".Rds")))

  for(iem in 1:nrow(subsetEM)){
    # Pull EM
    testEM <- readRDS(here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"]), "EMinput.Rds"))

    # Run simulation test
    simTestWHAM(nsim = nsim,
                OM = testOM,
                inputEMlist = list(testEM), # Run one EM at a time
                outdir = here::here(omdir, paste0("EM_missSeason_", subsetEM[iem, "miss_season"], "_missQ_", subsetEM[iem, "miss_q"])))
  }
}
```

### Run tests for EMs with seasonal misspecifications
Only run these for a subset of OMs to be selected following the outcomes of tests outlined above

## Post-process simulation results

```{r}
# ID all simulation result files 
files <- list.files(path = here::here(), pattern = "simWHAM_", recursive = TRUE)

postprocess_simTestWHAM(filenames = files) 
```

!!!! post-process based on storage & how sims run - could run sims based
varying isim - could also loop over OM but this could be computationally
expensive - I recommend labeling runs with OM and EM names (and setting
these up with informative) so it is easy to pull out results & compare
based on model setup - pull from mod\$model_name to help aggregate data

proposed function with: - fitted OM model object - list of EM inputs

If results processed in batches/on multiple computers use combinePerfMet to combine final results
```{r}

```