### 2023-09-29 Ecov_study/mortality/paper/paper.Rmd
save edits to ecov m paper

---
title: Factors affecting inferences on natural mortality and associated environmental effects
author: 
  - Timothy J. Miller^1^
  - Andrew Applegate 
  - Greg Britten
  - Elizabeth N. Brooks
  - Gavin Fay
  - Alex Hansell
  - Christopher M. Legault
  - Brandon Muffley
  - Brian C. Stock
  - John Wiedenmann
date: ''
fontsize: 12pt
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    includes:
      in_header:
      - options.sty
      - journalshortcuts.tex
    keep_tex: yes
    number_sections: yes
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{xspace}
- \usepackage{bm}
- \usepackage{caption,graphics}
- \usepackage{graphicx}
- \usepackage{makecell}
- \renewcommand\figurename{Fig.}
- \captionsetup{labelsep=period, singlelinecheck=false}
- \newcommand{\changesize}[1]{\fontsize{#1pt}{#1pt}\selectfont}
- \renewcommand{\arraystretch}{1.5}
- \renewcommand\theadfont{}
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
biblio-style: cjfas2.bst
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE, fig.align = "center")
knitr::opts_knit$set(root.dir = '../' )
library(knitr)
library(tidyverse)
library(pander)
library(kableExtra)
library(png)
library(forcats)
```

$^1$corresponding author: timothy.j.miller@noaa.gov, Northeast Fisheries Science Center, Woods Hole Laboratory, 166 Water Street, Woods Hole, MA 02543 USA\

\pagebreak

## Summary/Abstract {-}

We completed a large scale simulation study with 288 operating models with 12 estimated models fit to each. The factors defining operating model configuration inlcuded the source of process error on the population (recruitment, recruitment and survival, recruitment and natural mortality), the degree of temporal variation and autocorrelation of the environmental covariate, the uncertainty in the observation of the covariate, the uncertainty in indices and age composition observations, fishing history, and the magnitude of the effect of the covariate on natural mortality.

We found convergence of all estimation models was generally best when operating models assumed process errors in recruitment and survival, constant fishing rate, greater contrast in the true environmental covariate, and lower uncertainty in corresponding observations. Reliable convergence of all estimating models also occurred with the same process errors in the operating model and a step-change in fishing, but this also required lower uncertainty in index and age composition observations. Estimating models with process errors on recruitment and survival were unlikely to converge when the process errors in the operating model did not match whereas estimating models with process errors in recruitment and natural mortality converged for operating models without this match in certain cases. Probability of convergence generally decreased when the mean natural mortality rate parameter was estimated.

AIC paragraph

Bias paragraph

These results suggest that models with covariate effects on M will be most reliable when R+S is true, large contrast in the 
Recommendations
### Keywords {-}

\pagebreak

# Introduction {-}
<!--
Application of state-space models in fisheries stock assessment and management has expanded dramatically within ICES, Canada and the Northeast US \citep{nielsenberg14,cadigan16,stockmiller21}. State-space approaches that use random effects to parameterize process errors is considered best practice and a requirement for the next generation of stock assessment models.

Much is known about the reliability of state-space models that are linear or Gaussian \citep{aeberhardetal18}, but applications in fisheries management are nonlinear and typically include multiple types of observations with varying distributional assumptions. Furthermore, there is a wide range of potential random effects structures and model parameters that can be treated as random effects in assessment models. We know relatively little about the statistical reliability of such models or the ability of information criteria to distinguish among such alternative structures.

 But those studies focus primarily on Gaussian linear models.
review literature on reliability of hidden/latent process models. Primarily in other fields.

Convergence results can be useful for understanding how bad convergence in applications to real data might direct the practitioner to which alternative random effects configurations is be more appropriate. The type of convergence that we might use as a diagnostic is important. And what rule of thumb we use for certain criteria e.g. 1e-6? large SEs/CVs?
-->

Here we conduct a simulation study with operating models varying by degree of observation error uncertainty, sources of process error (M or survival), fishing history, temporal variation in environmental covariates, and magnitude of the effect of the covariate on natural mortality. The simulations from these operating models are fitted with estimating models that make alternative assumptions for sources of process error, and whether (mean) M is estimated. We evaluate effects of these factors on convergence of fitted models, whether AIC can correctly determine the correct source of process error and correct assumption about covariate effects on natural mortality, and the degree of bias in relevant parameters and outputs of the assessment model.

# Methods {-}

We used the WHAM package \citep[][version 1.0.7, commit 77bbd94]{millerstock20,stockmiller21} to configure all operating models and estimation models. This packages has also been used to configure operating and estimating models for closed loop simulations evaluating index-based assessment methods \citep{legaultetal23} and is used for management of GB haddock, butterfish, plaice, bluefish, and Atlantic cod stocks in the Northeast US.

We completed a simulation study with a number of operating models. Many of the characteristics of the operating models are the same as those used by \citet{milleretal_project0_wp}. The factors defining the configuration of each operating model which are described in detail in subsequent sections include 
\begin{itemize}
\item two configurations of fishing histories, 
\item two configurations of index and age composition observation error, 
\item four configurations of latent environmental covariate process errors, 
\item two configurations of environmental covariate observation uncertainty, 
\item three levels of covariate effects on natural mortality, and 
\item three configurations of sources of process errors for temporal changes in abundances at age. 
\end{itemize} 
The full factorial of the levels of each factor results in 288 different operating models. We simulated 100 data sets for each operating model that included simulations of process errors.

For each simulated data set we fit a set of 12 estimating models. The factors defining the configuration or each estimating model which are also described in detail below include
\begin{itemize}
\item three configurations of sources of process errors for temporal changes in abundances at age, 
\item whether the (mean) natural mortality rate was estimated or assumed known at the true value, and
\item whether the effect of the environmental covariate on natural mortality was estimated.
\end{itemize}

We did not use the bias-correction feature for process errors or observations described by \citet{stockmiller21} for operating and estimating models. Simulations were all carried out on the University of Massachusetts Green High-Performance Computing Cluster. Code for completing the simulations and summarization of results can be found at github.com/timjmiller/SSRTWG/ecov_study/mortality.

## Operating models {-}

### Population {-}

The population model tracks 10 age classes: ages 1 to 10+ and we assume spawning occurs 1/4 of the way through the year. The maturity at age was a logistic curve with a50 = 2.89 and slope = 0.88 and assumed known in all estimation models (Figure \ref{om_maturity}).

\begin{figure}
\caption{The proportion mature at age assumed for the population in all operating and estimating models.}\label{om_maturity}
\begin{center}
\includegraphics[width = \textwidth]{om_maturity.png}
\end{center}
\end{figure}

Weight at age was generated with a LVB growth function
$$
L_a = L_{\infty}\left(1 - e^{-k(a - t_0)}\right)
$$
where $t_0 = 0$, $L_\infty = 85$, and $k = 0.3$, and a L-W relationship such that 
$$
W_a = \theta_1 L_a^{\theta_2}
$$
where $\theta_1 = e^{-12.1}$ and $\theta_2 = 3.2$ (Figure \ref{om_waa}).

\begin{figure}
\caption{The weight at age assumed for the population in all operating and estimating models.}\label{om_waa}
\begin{center}
\includegraphics[width = \textwidth]{om_waa.png}
\end{center}
\end{figure}

We assumed a Beverton-Holt stock recruit function with constant pre-recruit mortality parameters (i.e., $\alpha$ and $\beta$) for all operating models. All post-recruit productivity components are constant in the NAA and survey selectivity process error operating models. Therefore steepness and unfished recruitment are also constant over the time period for those operating models \citep{millerbrooks21}. We specified unfished recruitment = $R_0 = e^{10}$ and $\Fmsy = \Fspr[40] = 0.348$ equated to a steepness of 0.69 and $\alpha=0.60$ and $\beta = 2.4 \times 10^{-5}$ for the
\[
N_{1,y} = \frac{\alpha \text{SSB}_{y-1}}{1 + \beta \text{SSB}_{y-1}} 
\]
Beverton-Holt parameterization (Figure \ref{om_sr}). For OMs without process errors on natural mortality we assumed the rate was assumed 0.2. For OMs with process errors on natural mortality the mean rate was 0.2.

\begin{figure}
\caption{The Beverton-Holt stock recruit relationship assume for all operating models.}\label{om_sr}
\begin{center}
\includegraphics[width = \textwidth]{om_sr.png}
\end{center}
\end{figure}

Two alternative fishing histories were used for operating models. In the first scenario, the stock experiences overfishing for the first 20 years and fishing at \Fmsy for the last 20 years. In the second scenario, the stock is fished at \Fmsy for the entire time period. The magnitude of the overfishing assumptions is based on average estimates of overfishing for NE groundfish stocks from \citet{wiedenmannetal19}. \Citet{legaultetal23} also used similar approaches to defining fishing mortality histories for operating models.

Initial population was configured at the equilibrium distribution fishing at either $F = 2.5\times \Fmsy$ or $F = \Fmsy$ for the two alternative fishing histories. That is for a deterministic model, the age composition would not change over time when the fishing mortality was constant at the respective level.

For operating models with time-varying random effects or covariate effects on M, steepness is not constant, but we used the same alpha and beta parameters as other operating models this equates to a steepness and $R_0$ at the mean of the time series process for M. For operating models with time-varying random effects for fishery selectivity, \Fmsy is also not constant however we use the same F history as other operating models which corresponds to \Fmsy at the mean selectivity parameters.

### Environmental covariate {-}

In the WHAM model, environmental covariates are assumed to be described as state-space processes with annual observations of the true latent covariate. The latent covariate is assumed to be Gaussian AR1
$$
E_y|E_{y-1} \sim \text{N}\left(\mu_\text{Ecov}\left(1-\rho_\text{Ecov}\right) + \rho_\text{Ecov} E_{y-1}, \left(1-\rho_\text{Ecov}^2\right)\sigma^2_\text{Ecov}\right)
$$
with marginal mean $\mu_\text{Ecov}=0$ and variance $\sigma^2_\text{Ecov}$. The configuration of the latent environmental covariate in the operating models have one of two alternate assumptions about the marginal standard deviation and autocorrelation parameter:
\begin{itemize}
\item marginal standard deviation $\sigma_\text{Ecov} \in \{0.1, 0.5\}$
\item AR1 correlation $\rho_\text{Ecov} \in \{0, 0.5\}$
\end{itemize}

The observations of the latent environmental covariate are assumed to be unbiased and Gaussian
$$
\text{ecov}_y|E_y \sim \text{N}\left(E_y,\sigma^2_\text{ecov}\right)
$$
The standard deviation of the environmental observations in the operating models is one of two values:
\begin{itemize}
\item standard deviation $\in \{0.1, 0.5\}$
\end{itemize}
Figure \ref{om_ecov_example} provides example simulations of the latent process and observations under the alternative configurations.

\begin{figure}
\caption{Example simulations of environmental covariate latent processes and observations with different levels of observation error, and different assumptions about variability of the latent process.}\label{om_ecov_example}
\begin{center}
\includegraphics[width = \textwidth]{Ecov_true_obs_example.png}
\end{center}
\end{figure}

### Random effects on recruitment, survival, and natural mortality {-}

Internally, WHAM treats natural mortality as a log-transformed parameter 
$$
\log M_y = \beta_M + \beta_{\text{Ecov}} E_y + \varepsilon_{M,y}
$$
that is a linear combination of a mean $\beta_M$ and any annual covariate $\beta_{\text{Ecov}}$ and random effects marginally distributed as $\epsilon_{M,y} \sim \text{N}\left(0,\sigma_M^2\right)$ \citep{stockmiller21}. For operating models with natural mortality random effects, we assume the same standard deviation $\sigma_M = 0.3$.

The covariate effect is one of 3 alternative values in the operating models:
\begin{itemize}
\item $\beta_\text{Ecov} \in \{0,0.25,0.5\}$
\end{itemize}

As described by \citet{stockmiller21}, the model for process errors on abundance at age in WHAM is
\begin{equation}
  \text{log}N_{a,y}=\left\{
    \begin{array}{@{}lll@{}}
      \text{log} \left( f(SSB_{y-1}) \right) + \varepsilon_{1,y}, & \text{if}\ a = 1 \\
      \text{log} \left( N_{a-1,y-1} \right) - Z_{a-1,y-1} + \varepsilon_{a,y}, & \text{if}\ 1 < a < A \\
      \text{log} \left( N_{A-1,y-1} e^{-Z_{A-1,y-1}} + N_{A,y-1} e^{-Z_{A,y-1}} \right) + \varepsilon_{A,y}, & \text{if}\ a = A
    \end{array}\right.
\end{equation}
We assume the the standard deviation of recruitment deviations is $\text{SD}\left(\varepsilon_{1,y}\right) = \sigma_R = 0.5$ for all operating models. For operating models with survival devations for older ages, we assume the same standard deviations for all older ages $\text{SD}\left(\varepsilon_{a,y}\right) = \sigma_{2+} = 0.3$.

The alternative configurations of random effects on recruitment, survival, and natural mortality are
\begin{itemize}
\item uncorrelated random effects on recruitment only ($\sigma_R = 0.5$) (R),
\item uncorrelated random effects on recruitment ($\sigma_R = 0.5$) and survival of older age classes ($\sigma_{2+} = 0.3$) (R+S), and
\item uncorrelated random effects on R ($\sigma_R = 0.5$) and log-natural mortality ($\sigma_M = 0.3$) (R+M).
\end{itemize}

### Fleets {-}

We assumed a single fleet operating year round for catch observations with logistic selectivity for the fleet with $a_{50} = 5$ and slope = 1 (Figure \ref{om_mean_selectivity}). This selectivity was used to define \Fmsy, \Fspr[40], and steepness for the Beverton-Holt stock recruitment parameters above. We assumed a logistic-normal distribution for the age-composition observations associated with the fleet where errors in the multivariate normal transformation are independent. The standard deviation parameter is also constant across ages.

\begin{figure}
\caption{The selectivity at age assumed for the fishing fleet in operating models without selectivity random effects. The mean selectivity at age across time for the fishing fleet in operating models with selectivity random effects.}\label{om_mean_selectivity}
\begin{center}
\includegraphics[width = \textwidth]{om_mean_selectivity.png}
\end{center}
\end{figure}

### Indices {-}

Two time series of surveys are assumed and observed in numbers rather than biomass for the entire 40 year period with one occurring in the spring (0.25 way through the year) and one in the fall (0.75 way through the year). Catchability of both surveys are assumed to be 0.1.  We assumed the same selectivity and age composition distribution as the fleet for both indices.

### Observation Uncertainty {-}

Standard deviation for log-aggregate catch was 0.1. There were two levels of observation error variance for indices and age composition for both indices and fleet catch. A low uncertainty specification assumed standard deviation of both series of log-aggregate index observations was 0.1 and the standard deviation of the logistic-normal for age composition observations was 0.3. In the high uncertainty specification the standard deviation for log-aggregate indices was 0.4 and that for the age composition observations was 1.5. For all estimating models, standard deviation for log-aggregate observations was assumed known whereas that for the logistic-normal age composition observations was estimated.

## Estimating models {-}

For each data set simulated from an operating model 12 estimating models were fit. There were three factors defining the configuration of each estimating model

\begin{itemize}
\item whether the mean natural mortality $\beta_M$ was estimated or assumed known ($\log 0.2$)
\item whether an environmental effect $\beta_\text{Ecov}$ was estimated or not (fixed at 0)
\item whether the process errors were just on recruitment (R), on recruitment and survival (R+S) or on recruitment and natural mortality (R+M)
\end{itemize}
The configuration of the process errors in the estimating models matched the corresponding options in the operating models. For example, if uncorrelated R+S was assumed for both the estimating and operating model, the assumptions about process errors matched. The environmental covariate observations were included in all estimation models to ensure comparability of information criteria. All fixed effects parameters for selectivity, catchability, fully-selected fishing mortality, mean recruitment, initial abundance at age, and variances for logistic-normal age composition distributions were estimated. Any process error variance parameters for recruitment, survival, and natural mortality were also estimated. The observation error variance of the environmental observations and aggregate catch and indices were all assumed known at the true values.

## Measures of reliability {-}

### Convergence {-}

The first measure of reliability we investigated was frequency of convergence when fitting each estimating model to the simulated data sets. There are various ways to assess convergence of the fit. We summarized 5 alternative categories of convergence:
\begin{enumerate}
\item Did the optimization routine (stats::nlminb) complete without error?
\item Did the stats::nlminb convergence flag = 0 indicate successful convergence?
\item Was the maximum absolute value of the gradient of the log-likelihood < $1\times10^{-6}$?
\item Did TMB::sdreport provide non-NA values for all fixed effects standard errors?
\item Did TMB::sdreport provide all standard errors < 100?
\end{enumerate}
The first convergence criterion assesses whether the model crashes. The third is a measure that assesses how flat the likelihood is at the optimized point of the likelihood surface. The fourth and fifth criteria are specific to the hessian-based standard error reporting provided by TMB. the TMB::sdreport function will sometimes return standard error estimates even when the calculated hessian is not invertible (fourth criterion). It will also provide large standard errors for some parameters that are not estimated well or are near bounds on the transformed scale that is of primary interest. For example, variance parameters for random effects can often be estimated near 0 when the fit suggests no variation in the random effects or, equivalently, that the model without random effects is better.

### AIC for model selection {-}

We measured the frequency of correct model selection using marginal AIC. For a given operating model the set of models that were considered all made the same assumptions on whether or not to estimate (mean) natural mortality rate or it is assumed at the true value $\beta_M = log(0.2)$. For model $m$, the marginal AIC is a function of the marginal log-likelihood maximized with respect to the fixed effects in the model $\boldsymbol{\theta}$ and the number of fixed effects $n\left(\boldsymbol{\theta}\right)$ estimated,
$$
\text{AIC}_m = -2\left[{\text{argmax}}_{\boldsymbol{\theta}} \log L_m\left({\boldsymbol{\theta}}\right) - n\left({\boldsymbol{\theta}}\right)\right].
$$

### Bias {-}

We calculated median relative errors of
\begin{itemize}
\item $\beta_\text{Ecov}$, the effect of environmental covariates on natural mortality, 
\item $\beta_M$, the mean log-natural mortality rate,
\item annual natural mortality rate,
\item annual spawning stock biomass, and
\item annual fully-selected fishing mortality rate.
\end{itemize}
For the $i$th simulated data the relative error for a parameter $\theta$ provided from the fitted estimation model is
$$
{\text{RE}_i}\left(\theta\right) = \frac{\widehat \theta_i - \theta_i}{\theta_i}
$$
and measured bias as the median relative error and constructed 95\% confidence intervals using the binomial distribution approach as in \citet{stockmiller21}. If the confidence interval contains zero we do not have evidence of bias from the simulation study. We used estimates from all simulations that satisfied the first type of convergence described above. We were not concerned about more restrictive types of convergence because some estimation models would not be expected to satisfy these criteria because of mis-specified structures. For example, an estimation model that includes survival random effects fitted to a set of observations simulated without survival random effects should have trouble estimating variance of the survival random effects, but the estimates of the parameters and derived output described above, might be reliable.

# Results {-}

## Convergence performance {-}

We focused on the fourth convergence criteria, that the hessian of the negative-marginal log-likelihood at the optimized parameters was invertible. When the (mean) natural mortality rate was assumed at the true value, corresponding to usual practice in application of assessment models, good convergence for all EMs was observed for R+S operating models with \Fmsy fishing history, more variation in the latent environmental covariate ($\sigma_\text{Ecov} = 0.5$), and lower error in the associated observations $\sigma_\text{ecov} = 0.1$ (Figure \ref{convergence_M_fixed}). Good convergence for all EMs was also observed for R+S OMs with the step change in fishing history, but with low error in indices and age comp observations.

EMs that assumed no survival or natural mortality random effects (R OMS), nearly always converged across all operating models that assumed more variation in the latent environmental covariate ($\sigma_\text{Ecov} = 0.5$) and lower error in the associated observations $\sigma_\text{ecov} = 0.1$. EMs that assumed recruitment and survival random effects (R+S OMS), had poor converge probability when the OMs had alternative process errors (R or R+M). The R+S EMs showed best convergence for the R+S OMs where there was lower error in the environmental observations $\sigma_\text{ecov} = 0.1$ or higher error in environmental observations of latent true environmental covariates with greater temporal variation. EMs that assumed recruitment and natural mortality random effects (R+M OMs), had poor convergence probability for all OMs that had a step change in fishing history and higher uncertainty in indices and age composition.

When the OMs and EMs both assume process errors only on recruitment (R) or on recruitment and survival (R+S), convergence was worst with less variation in the true environmental covariate and larger uncertainty in associated observations. When the OMs and EMs both assume process errors on recruitment and natural mortality (R+M) convergence was problematic for all OMs with the step-change in fishing history. The best convergence was observed with this match between OMs and EMs was when the fishing history was constant and there was low uncertainty in environmental observations.

\begin{landscape}
\begin{figure}
\caption{Probability of convergence of estimating models with alternative process error assumptions fit to simulated populations and data sets from operating models. The convergence criterion is an invertible hessian providing standard error estimates for all fixed effects parameters. Vertical lines represent 95\% confidence intervals. All estimating models estimate environmental covariate effects on natural mortality but assume the mean mortality parameter $\beta_\text{M} = \log 0.2$.}\label{convergence_M_fixed}
\begin{center}
\includegraphics[height = \textheight]{proportion_good_hessian_ecov_effect_est_M_fixed.png}
\end{center}
\end{figure}
\end{landscape}

As might be expected, there was an overall drop in the probability of convergence when the mean natural mortality rate was estimated rather than assumed at the true value (Figure \ref{convergence_M_estimated}). Otherwise, general trends described above with mean natural mortality fixed, apply when estimated.

\begin{landscape}
\begin{figure}
\caption{Probability of convergence of estimating models with alternative process error assumptions fit to simulated populations and data sets from operating models. The convergence criterion is an invertible hessian providing standard error estimates for all fixed effects parameters. Vertical lines represent 95\% confidence intervals. All estimating models estimate environmental covariate effects on natural mortality and the mean mortality parameter $\beta_\text{M}$.}\label{convergence_M_estimated}
\begin{center}
\includegraphics[height = \textheight]{proportion_good_hessian_ecov_effect_est_M_est.png}
\end{center}
\end{figure}
\end{landscape}

## AIC performance {-}

When estimating models assumed the mean natural mortality rate was known, the best accuracy of AIC for model selection occurred for models with R+S process errors. R+S estimating models ranked best with very low frequency for R or R+M operating models and with very high frequency for R+S operating models (Figure \ref{aic_rank_M_fixed}). R estimating models were determined best with high frequency for R operating models, but also for R+M operating models. R+M estimating models were rarely determined best for any operating models including those where the process errors matched.

AIC was conservative for determining whether the environmental covariate affected natural mortality. AIC was highly accurate in determining no effect when there was no effect in the operating model, but AIC ranked the null model best with high frequency even when there was an effect in the operating model in many cases. However the accuracy of AIC improved in certain operating models. Increased effect size, increased temporal contrast in the covariate, and lower uncertainty in all observations types lead to increased accuracy of determining covariate effects.

Relative to the assumption that the mean natural mortality rate was known, estimating the mean natural mortality rate had small effects on the accuracy of AIC in selecting the appropriate process error and whether the covariate affect natural mortality (Figure \ref{aic_rank_M_estimated}). Where there were differences there were small decreases in accuracy for determining the appropriate process error and determining a covariate effect when there was one.

<!--
\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets for each operating model where the estimation model with the lowest AIC had correct assumptions of process errors (R,R+S,R+M) and/or about whether the environmental covariate affected natural mortality. All estimating models assumed the mean mortality parameter $\beta_\text{M} = \log 0.2$.}\label{correct_assumptions_aic_M_fixed}
\begin{center}
\includegraphics[height = \textheight]{proportion_correct_PE_effect_M_fixed.png}
\end{center}
\end{figure}
\end{landscape}
-->

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets for each operating model where the estimation model type (process error and covariate effect assumptions) had the lowest AIC. All estimating models assumed the mean mortality parameter $\beta_\text{M} = \log 0.2$.}\label{aic_rank_M_fixed}
\begin{center}
\includegraphics[height = \textheight]{proportion_best_AIC_M_fixed.png}
\end{center}
\end{figure}
\end{landscape}

<!--
\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets for each operating model where the estimation model with the lowest AIC had correct assumptions of process errors (R,R+S,R+M) and/or about whether the environmental covariate affected natural mortality. All estimating models estimate the mean mortality parameter $\beta_\text{M}$.}\label{correct_assumptions_aic_M_estimated}
\begin{center}
\includegraphics[height = \textheight]{proportion_correct_PE_effect_M_estimated.png}
\end{center}
\end{figure}
\end{landscape}
-->

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets for each operating model where the estimation model type (process error and covariate effect assumptions) had the lowest AIC. All estimating models assumed the mean mortality parameter $\beta_\text{M} = \log 0.2$.}\label{aic_rank_M_estimated}
\begin{center}
\includegraphics[height = \textheight]{proportion_best_AIC_M_estimated.png}
\end{center}
\end{figure}
\end{landscape}

## Bias {-}

### Environmental effect {-}

When the EMs assumed the mean natural mortality rate was known, we observed generally accurate estimation of environmental effects across all EM and OM process error assumptions and all true covariate effect sizes, when ther was low uncertainty in environmental observations and larger temporal contrast in the simulated true environmental covariate (Figure \ref{Ecov_beta_bias_M_fixed}). We observed a negative trend in bias of the environmental effect with increased effect size when temporal variation in the covariate was lower and/or uncertainty in the covariate observations was higher. When the OMs had R+S process errors with low temporal variation in the true environmental covariate and lower uncertainty in the indices age age composition, estimated covariate effects were highly variable.  In most cases the relative error of $\beta_\text{Ecov}$ did not depend on the source of process error assumed in the EM. When there was an effect of the EM process error assumption it was when OMs had R+S process errors. The worst bias was observed when OMs assumed R+S process errors, high uncertainty in covariate observations, low variability in the covariate, and low uncertainty in index and age composition observations.

When the mean natural mortality rate was estimated in the EM, results were similar except estimated effects were even more variable for data simulated with R+S process errors (Figure \ref{Ecov_beta_bias_M_estimated}). There was also more separation of reliability of the estimation of the effect among EMs with different process error assumptions. The separation was most apparent when OMs simulated R+S process errors and larger variability in the environmental covariate wehere EMs with process errors other than R+S showing more bias than when the mean natural mortality rate was assumed known.

\begin{landscape}
\begin{figure}
\caption{Median relative error of the estimated environmental effect on natural mortality $\beta_\text{Ecov}$ from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model. All estimating models assumed the mean mortality parameter $\beta_\text{M} = \log 0.2$.}\label{Ecov_beta_bias_M_fixed}
\begin{center}
\includegraphics[height = \textheight]{Ecov_beta_bias_all_PE_effect_M_fixed.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative error of the estimated environmental effect on natural mortality $\beta_\text{Ecov}$ from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model. All estimating models assumed the mean mortality parameter $\beta_\text{M}$ is also estimated.}\label{Ecov_beta_bias_M_estimated}
\begin{center}
\includegraphics[height = \textheight]{Ecov_beta_bias_all_PE_effect_M_estimated.png}
\end{center}
\end{figure}
\end{landscape}

### Mean natural mortality rate {-}

\begin{landscape}
\begin{figure}
\caption{Median relative error of the estimated mean natural mortality parameter $\beta_\text{M}$ from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model.}\label{mean_M_bias}
\begin{center}
\includegraphics[height = \textheight]{mean_M_bias_all_PE_effect.png}
\end{center}
\end{figure}
\end{landscape}

### Annual natural morality rate {-}

\begin{landscape}
\begin{figure}
\caption{Median relative error of annual natural mortality rate in years 1 (Start), 21 (Middle), and 40 (End) from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model. Estimating models also assume the mean natural mortality parameter $\beta_\text{M} = \log 0.2$ and no environmental covariate effect $\beta_\text{Ecov} = 0$.}\label{annual_M_bias_M_fixed_beta_fixed}
\begin{center}
\includegraphics[height = \textheight]{annual_M_bias_all_PE_effect_M_fixed_beta_fixed.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative error of annual natural mortality rate in years 1 (Start), 21 (Middle), and 40 (End) from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model. Estimating models also assume the mean natural mortality parameter $\beta_\text{M} = \log 0.2$ and the environmental covariate effect $\beta_\text{Ecov}$ is estimated.}\label{annual_M_bias_M_fixed_beta_estimated}
\begin{center}
\includegraphics[height = \textheight]{annual_M_bias_all_PE_effect_M_fixed_beta_estimated.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative error of annual natural mortality rate in years 1 (Start), 21 (Middle), and 40 (End) from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model. Estimating models also assume both the mean natural mortality parameter $\beta_\text{M}$ and the environmental covariate effect $\beta_\text{Ecov}$ are estimated.}\label{annual_M_bias_M_estimated_beta_estimated}
\begin{center}
\includegraphics[height = \textheight]{annual_M_bias_all_PE_effect_M_estimated_beta_estimated.png}
\end{center}
\end{figure}
\end{landscape}

### Spawning stock biomass {-}

\begin{landscape}
\begin{figure}
\caption{Median relative error of annual SSB in years 1 (Start), 21 (Middle), and 40 (End) from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model. Estimating models also assume the mean natural mortality parameter $\beta_\text{M} = \log 0.2$ and no environmental covariate effect $\beta_\text{Ecov} = 0$.}\label{SSB_bias_M_fixed_beta_fixed}
\begin{center}
\includegraphics[height = \textheight]{SSB_bias_all_PE_effect_M_fixed_beta_fixed.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative error of annual SSB in years 1 (Start), 21 (Middle), and 40 (End) from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model. Estimating models also assume the mean natural mortality parameter $\beta_\text{M} = \log 0.2$ and the environmental covariate effect $\beta_\text{Ecov}$ is estimated.}\label{SSB_bias_M_fixed_beta_estimated}
\begin{center}
\includegraphics[height = \textheight]{SSB_bias_all_PE_effect_M_fixed_beta_estimated.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative error of annual SSB in years 1 (Start), 21 (Middle), and 40 (End) from fitting simulated observations from each operating model with alternative process error assumptions in the estimating model. Estimating models also assume both the mean natural mortality parameter $\beta_\text{M}$ and the environmental covariate effect $\beta_\text{Ecov}$ are estimated.}\label{SSB_bias_M_estimated_beta_estimated}
\begin{center}
\includegraphics[height = \textheight]{SSB_bias_all_PE_effect_M_estimated_beta_estimated.png}
\end{center}
\end{figure}
\end{landscape}

# Discussion {-}

The estimating models assumed variances of aggregate catch and index observations was known. This approximation may be appropriate for indices where we have a reliable estimate of uncertainty based on the survey design (), but there may be better approaches for the aggregate catch such as an informed prior on the standard errors with realistic bounds.

We found EMs with R+M process errors were rarely determined as the appropriate model when OMs simulated R+M process errors. This was unexpected but is likely due to the size of the variance assumed for those process errors ($\sigma_M = 0.3$) relative to the variances assumed for index and age composition observations. This is related to the difficulty in separately estimating observation and process error variances when the ratio of process to observation uncertainty is low.

Note that the results for bias of covariate effect all assume ecov beta is estimated. The lack of bias in certain situations might suggest including the effect even if AIC says it isn't better at least when contrast in Ecov is high.

The bias in natural mortality estimation resulted in biased estimation of stock size and likely harvest rates, but it will also propagate into biological reference points and possibly stock status.

# Conclusions {-}

As we would expect, reliable detection of covariate effects requires informative data. 
AIC preferred simpler models than the true model when information content in data and contrast in covariates and abundance were low. Null model for environmental covariate effect (no covariate effect) was selected when contrast in the time series was low  and/or uncertainty in observations was high.

When there was process error in recruitment and M  (R+M), models with process error only in recruitment were preferred. This could be because the variation in the M random effects was small relative to observation variability. The marginal variance for log M random effects was 0.3.

Null selection likely decreases with strength of the effect on M. We only examined two non-zero effect sizes: 0.25, and 0.5, but our results suggest larger effect sizes, with the same observation error and contrast in time series, would allow better AIC performance for determining covariate effects.

Covariate effect estimation can be robust to process error assumptions (For R+M OMs) with contrast in covariate and low observation error.

Reliable estimation of environmental effect on M regardless of the process error assumed by the EM as long as the contrast in the covariate is sufficient and the uncertainty in the observations is low.

# Acknowledgements {-}

This work was funded by NOAA Fisheries Northeast Fisheries Science Center.

<!--
# CRediT authorship contribution statement {-}

**Timothy J. Miller:** Conceptualization, Methodology, Writing - original draft, Formal analysis, Visualization.
 -->

\pagebreak

\bibliography{paper}
<!-- \bibliography{project-0-paper}
\bibliographystyle{/home/tmiller2/work/bibliographies/cjfas2}  -->

<div id="refs"></div>

\pagebreak