### 2024-02-06 Ecov_study/recruitment_functions/presentation/SRR_presentation.Rmd
presentation updates

---
title: "Identification and performance of stock-recruitment functions in state
space assessment models"
subtitle: Working Paper 1
author: Greg Britten, Liz Brooks, Tim Miller
output: beamer_presentation
fig_caption: FALSE
theme: "Singapore"
slide-level: 3
# titlegraphic: WHOI_NOAA_logo_smaller.png
header-includes: 
- \addtobeamertemplate{title page}{\centering \includegraphics[width=9.5cm]{WHOI_NOAA_logo_smaller.png}\par }{}
classoption: "aspectratio=169"
date: "`r Sys.Date()`"
---
# Background

## Introduction

- State Space Research Track Working Group (SSRTWG) is investigating performance of the Woods Hole Assessment Model (WHAM)
- Simulation studies with data generated from operating models (OMs), then fit with a series of estimation models (EMs)
- This framework allows us to evaluate how the EM fits compare to the known OM "true values", evaluate model selection, bias, precision, etc.

## Terms of Reference (TORs) Addressed

- TOR 2: Investigate the efficacy of estimating stock-recruit functions within state-space models and their utility in generating scientific advice.

- TOR 3: Develop guidelines for including ecosystem and environmental effects in assessment models and how to treat them for generating biological reference points and scientific advice.

## Outline

- 'Stock' parameters, fishery and index parameters
- OM factors and simulated data examples
- EM models
- Beta standardization
- Analysis & Results
- Conclusions
- Future Work

<!-- ==================================================================================== -->

# Inputs

## Stock parameters for generic gadid

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'presentation', 'plots', 'bio.plot.png') )
```

- Natural mortality = 0.2 at all ages
- Maturity, weight, natural mortality are time-invariant

## Fishery and index parameters

<!-- -- example chunk to include figure from file -->
```{r,  echo=FALSE, message=FALSE, warnings=FALSE, results='hide', out.width='60%', fig.align='center'}
library(tidyverse)
library(wham)
naa_om_inputs <- readRDS(file.path(here::here(),"Project_0","inputs", "NAA_om_inputs.RDS"))
om.data <- naa_om_inputs[[1]]$data

gf_info = make_basic_info()

#selectivity is not changing
gf_selectivity <- list(model       =c(rep("logistic", gf_info$n_fleets),rep("logistic", gf_info$n_indices)),
                       initial_pars=rep(list(c(5,1)), gf_info$n_fleets + gf_info$n_indices)) #fleet, index

#fishery
a50 = gf_selectivity$initial_pars[[1]][1]
k= gf_selectivity$initial_pars[[1]][2]
sel.tmp <- 1.0/(1.0 + exp(-(seq(1, om.data$n_ages) - a50)/k))
sel <- sel.tmp/sel.tmp[om.data$n_ages]
fish.sel <- as_tibble(cbind(Age=seq(1, om.data$n_ages), Sel=sel))
ggplot(fish.sel, aes(x=Age, y=Sel)) +
  geom_line(color='green4', linewidth=1) +
  geom_point(color='green4', size=2.5) +
  ylab('Fishery Selectivity at Age') +
  theme(axis.text.x = element_text(size = 15))   +
  theme(axis.text.y = element_text(size = 15))  +
  theme(axis.title.y = element_text(size = 16))  +
  theme(axis.title.x = element_text(size = 16))

#index-1 and index-2 have same selectivity as fleet (why?!)
# a50.i1 = gf_selectivity$initial_pars[[2]][1]
# k.i1 = gf_selectivity$initial_pars[[2]][2]
# 
# a50.i2 = gf_selectivity$initial_pars[[2]][1]
# k.i2 = gf_selectivity$initial_pars[[2]][2]

ind.time <- naa_om_inputs[[1]]$data$fracyr_indices[1,]
q.hi <- (naa_om_inputs[[1]]$data$q_upper)
q.lo <- (naa_om_inputs[[1]]$data$q_lower)
ind.q <-  unique((q.lo + q.hi/(1 + exp(-1*naa_om_inputs[[1]]$par$logit_q)) ))

```
- two fishery independent indices were also generated, taking place at `r ind.time` yr
- catchability for both indices was `r ind.q `; selectivity was same as fishery

<!-- ==================================================================================== -->

# OM

## OM factors and simulated data examples

<!-- -- example chunk to include figure from file -->
```{r, echo=FALSE, message=FALSE, warnings=FALSE, out.width='50%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'model_probability_beta_fix.png') )
```

- make some plots to illustrate OM inputs and resulting simulated data generated
- looks like i can have multiple bullets below a figure

## Couple of slides for Beta standardization

- greg's plots

<!-- ==================================================================================== -->

# EM

## EM models

- simple bullets above table
- we only have 6 EMs that were considered

```{r,  echo=FALSE, message=FALSE, warnings=FALSE}

df.ems    <- readRDS(file.path(here::here(),"Ecov_study", "recruitment_functions", "inputs", "df.ems.RDS"))

em_tib <- as_tibble(df.ems) %>%
  mutate(SR=ifelse(r_mod==2, 'Mean', 'BH')) %>%
  mutate(EM_mod = paste0(SR, "_", ecov_how), EM=seq(1,6))

   knitr::kable(em_tib)
  # print(em_tib)

```

<!-- ==================================================================================== -->

# Analysis & Results

## Analyses

\begin{enumerate}
  \item Convergence of the estimating models
  \item Model identifiability of an underlying stock recruitment model and/or an underlying relationship
   between environmental covariate
  \item $\Delta$AIC and model probability
  \item Assessment error (recruitment, spawning stock biomass, and Fbar)
  \item Bias of estimated parameters
  \item Mohn's $\rho$
  \item Projection performance relative to assumptions about the environmental covariate
\end{enumerate}

## Convergence

## Model Identifiability

##  $\Delta$AIC and model probability

## Assessment error (recruitment, spawning stock biomass, and Fbar)

## Bias of estimated parameters

## Mohn's $\rho$

## Projection performance relative to assumptions about the environmental covariate

<!-- ==================================================================================== -->

# Conclusions

## Take-aways
- These are the take-aways (copy from WP)

# Future Work

## List of what's next

- This is what we suggest for follow-up
- Note that WP1-Appendix looked at $\sigma_{R}$=0.5 and found no difference from results in WP1

<!-- ==================================================================================== -->

---

Acknowledgements

- This work could not have been completed without the use of Azure computing (NOAA) and MIT (... greg to fill in)
- We thank other members of the SSRTWG for thoughtful comments during earlier discussions and presentations of this work

```{r, echo=FALSE, message=FALSE, warnings=FALSE, out.width='35%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'presentation', 'abacus_crop.jpg') )
```

<!-- ==================================================================================== -->

<!-- ## R Markdown -->

<!-- If you have trouble knitting this file, check out this page: https://stackoverflow.com/questions/67696286/error-generating-pdf-using-knitr-to-pdf-in-rstudio -->

<!-- ```{r cars, echo = TRUE} -->
<!-- summary(cars) -->
<!-- ``` -->

<!-- ## Slide with Plot -->

<!-- ```{r pressure} -->
<!-- plot(pressure) -->
<!-- ``` -->