### 2024-02-07 Ecov_study/recruitment_functions/presentation/SRR_presentation.Rmd
nearly final presentation

---
title: "Identification and performance of stock-recruitment functions in state
space assessment models"
subtitle: Working Paper 1
author: Greg Britten, Liz Brooks, Tim Miller
output: beamer_presentation
fig_caption: FALSE
theme: "Singapore"
slide-level: 3
# titlegraphic: WHOI_NOAA_logo_smaller.png
header-includes: 
- \addtobeamertemplate{title page}{\centering \includegraphics[width=9.5cm]{WHOI_NOAA_logo_smaller.png}\par }{}
classoption: "aspectratio=169"
date: "`r Sys.Date()`"
---

# Background

## Introduction

- State Space Research Track Working Group (SSRTWG) is investigating performance of the Woods Hole Assessment Model (WHAM)
- Simulation studies with data generated from operating models (OMs), then fit with a series of estimation models (EMs)
- This framework allows us to evaluate how the EM fits compare to the known OM "true values", evaluate model selection, bias, precision, etc.

## Terms of Reference (TORs) Addressed

- TOR 2: Investigate the efficacy of estimating stock-recruit functions within state-space models and their utility in generating scientific advice.

- TOR 3: Develop guidelines for including ecosystem and environmental effects in assessment models and how to treat them for generating biological reference points and scientific advice.

## Outline

- 'Stock' parameters, fishery and index parameters
- OM factors and simulated data examples
- EM models
- Beta standardization
- Analysis & Results
- Conclusions
- Future Work

<!-- ==================================================================================== -->

# Inputs

## Stock parameters for generic gadid

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'presentation', 'plots', 'bio.plot.png') )
```

- Natural mortality = 0.2 at all ages
- Maturity, weight, natural mortality are time-invariant

## Fishery and index parameters

<!-- -- example chunk to include figure from file -->
```{r,  echo=FALSE, message=FALSE, warnings=FALSE, results='hide', out.width='60%', fig.align='center'}
library(tidyverse)
library(wham)
naa_om_inputs <- readRDS(file.path(here::here(),"Project_0","inputs", "NAA_om_inputs.RDS"))
om.data <- naa_om_inputs[[1]]$data

gf_info = make_basic_info()

#selectivity is not changing
gf_selectivity <- list(model       =c(rep("logistic", gf_info$n_fleets),rep("logistic", gf_info$n_indices)),
                       initial_pars=rep(list(c(5,1)), gf_info$n_fleets + gf_info$n_indices)) #fleet, index

#fishery
a50 = gf_selectivity$initial_pars[[1]][1]
k= gf_selectivity$initial_pars[[1]][2]
sel.tmp <- 1.0/(1.0 + exp(-(seq(1, om.data$n_ages) - a50)/k))
sel <- sel.tmp/sel.tmp[om.data$n_ages]
fish.sel <- as_tibble(cbind(Age=seq(1, om.data$n_ages), Sel=sel))
ggplot(fish.sel, aes(x=Age, y=Sel)) +
  geom_line(color='green4', linewidth=1) +
  geom_point(color='green4', size=2.5) +
  ylab('Fishery Selectivity at Age') +
  theme(axis.text.x = element_text(size = 15))   +
  theme(axis.text.y = element_text(size = 15))  +
  theme(axis.title.y = element_text(size = 16))  +
  theme(axis.title.x = element_text(size = 16)) +
  scale_x_continuous(breaks=fish.sel$Age, labels=as.character(fish.sel$Age))

#index-1 and index-2 have same selectivity as fleet (why?!)
# a50.i1 = gf_selectivity$initial_pars[[2]][1]
# k.i1 = gf_selectivity$initial_pars[[2]][2]
# 
# a50.i2 = gf_selectivity$initial_pars[[2]][1]
# k.i2 = gf_selectivity$initial_pars[[2]][2]

ind.time <- naa_om_inputs[[1]]$data$fracyr_indices[1,]
q.hi <- (naa_om_inputs[[1]]$data$q_upper)
q.lo <- (naa_om_inputs[[1]]$data$q_lower)
ind.q <-  unique((q.lo + q.hi/(1 + exp(-1*naa_om_inputs[[1]]$par$logit_q)) ))

```
- two fishery independent indices were also generated, taking place at `r ind.time` yr
- catchability for both indices was `r ind.q `; selectivity was same as fishery

<!-- ==================================================================================== -->

# OM

## OM factors 
```{r,  echo=FALSE, message=FALSE, warnings=FALSE}

df.oms    <- readRDS(file.path(here::here(),"Ecov_study", "recruitment_functions", "inputs", "df.oms.RDS"))

om.table <- cbind( unique(df.oms$Fhist), unique(df.oms$R_sig), unique(df.oms$NAA_cor), unique(df.oms$Ecov_effect), unique(df.oms$Ecov_re_cor), unique(df.oms$obs_error)  )
colnames(om.table) <- c('Fhistory', 'R_sigma', 'R_cor', 'Ecov_effect', 'Ecov_cor', 'Obs_error')
   knitr::kable(om.table)

```

- These levels were combined factorially with 4 stock recruit models (all Beverton-Holt) for 256 different OM
- Factors that did not vary were the observation and process error of the Ecov (bot fixed at 0.1)

## OM factor for Recruitment variability (R_sig) relative to recent assessments
```{r,  echo=FALSE, message=FALSE, warnings=FALSE}

rsig.assmt <- read.csv(file=file.path(here::here(),"Ecov_study", "recruitment_functions", "tables", "sigmaR_assessment_estimates.csv"), header=TRUE)

   knitr::kable(rsig.assmt )

```

## OM factor: Beverton-Holt functional (1 with no Ecov effect, 3 with effect)

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'paper', 'dRdX_1RdRdx.pdf') )
```

## OM factor: Beta standardization

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'paper', 'dRdX.pdf') )
```

## Some Example OM Simulated Data (100 data sets for each OM)
- still need to make this

<!-- ==================================================================================== -->

# EM

## EM models

- All OMs used Beverton-Holt stock recruit relationship (SRR)
- For the EMs, we also fit a mean SRR with random effects
- There were 6 total EMs fit to each simulated data set (100) for all 256 OMs 
- 153,600 fitted models X 2 for $\beta$ unstandardized and $\beta$ standardized cases

```{r,  echo=FALSE, message=FALSE, warnings=FALSE}
library(tidyverse)

df.ems    <- readRDS(file.path(here::here(),"Ecov_study", "recruitment_functions", "inputs", "df.ems.RDS"))
df.ems2 <- cbind(EM=seq(1,6), df.ems)

em_tib <- as_tibble(df.ems) %>%
  mutate(SR=ifelse(r_mod==2, 'Mean', 'BH')) %>%
  mutate(EM_mod = paste0(SR, "_", ecov_how), EM=seq(1,6)) %>%
  rename(Ecov_how=ecov_how, SRR=SR) %>%
  mutate(Ecov=c('None', 'Controlling', 'None', 'Controlling', 'Limiting', 'Masking' )) %>%
  relocate(EM, SRR, Ecov_how, Ecov, EM_mod) %>%
  select(-r_mod)

   knitr::kable(em_tib)
  # print(em_tib)

```

<!-- ==================================================================================== -->

# Analysis & Results

## Analyses

\begin{enumerate}
  \item Convergence of the estimating models
  \item Model identifiability of an underlying stock recruitment model and/or an underlying relationship
   between environmental covariate
  \item $\Delta$AIC and model probability
  \item Assessment error (recruitment, spawning stock biomass, and Fbar)
  \item Bias of estimated parameters
  \item Mohn's $\rho$
  \item Projection performance relative to assumptions about the environmental covariate
\end{enumerate}

- Results will be shown for $\beta$ unstandardized case, because $\beta$ standardization had no appreciable effect
- Most results summarized by relative error (RE): (Est-True)/True

## Crashes

- ~3.3-3.5% crashes, typically for Ecov=Masking ("BH_4") and $\sigma_{R}$=0.1
- most problematic for Masking (BH_4) and exacerbated by $\sigma_{R}$=0.1
```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'crashed.models.plot_beta_fix.png') )
```

## Convergence Criteria

\begin{enumerate}
 \item the optimization routine (stats::nlminb) completed without error
 \item the stats::nlminb convergence flag = 0 indicated successful convergence
 \item the maximum absolute value of the gradient of the log-likelihood is < 1e-6
 \item TMB::sdreport provided non-NA values for all fixed effects standard errors
 \item TMB::sdreport provided all standard errors < 100
\end{enumerate}

- Models not satisfying all of these conditions were dropped from Results summaries

## Unconverged Models (from EM having'best' EM per OM-Simulation)
- most problematic for $\sigma_{R}$=0.1; BH with no Ecov and high fishing contrast (H-MSY) and mean SRR with no contrast (MSY)
```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'bad.mods_lowestAIC.plot_grad_6_SE_100_beta_fix.png') )
```

## Unconverged Models (considering all EMs per OM-Simulation)
- most problematic for Beverton-Holt SRR, worse with no contrast in fishing history (MSY)

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'bad.mods_all_grad_6_SE_100_beta_fix.plot.png') )
```

## Unconverged Model - The Offending Parameters 
- the SRR recruitment parameters led to most convergence failures

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'bad_par_N_barplot_grad_6_SE_100_beta_fix.png') )
```

## Unconverged Model - Second Most Offending Parameters

- the Ecov parameters led to the second most convergence failures

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'bad_par_2nd_N_barplot_grad_6_SE_100_beta_fix.png') )
```

## Model Identifiability - Marginal barplots for SRR
 - Marginal univariate model selection results for whether an SRR relationship was correctly
identified (indicated by dark shaded regions)

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'raw_boxplots_SR_beta_fix.pdf') )
```

## Model Identifiability - Marginal barplots for Ecov_how
 - Marginal univariate model selection results for whether an Ecov was correctly identified (indicated by dark shaded regions)

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'raw_boxplots_Ecov_beta_fix.pdf') )
```

## Model Identifiability - Marginal barplots for both SRR & Ecov_how
 - Marginal univariate model selection results for whether the SRR AND Ecov were both correctly identified (indicated by dark shaded regions)

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'raw_boxplots_form_beta_fix.pdf') )
```

## Model Identifiability - Effect size
 - Effect sizes for proportion correctly identified using a binomial generalized linear model and a
binomial generalized linear mixed model with simulation random number seed as a random intercept

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'effect_size_glm_glmm_beta_fix.pdf') )
```

## Model Identifiability - Classification tree for identifying EM = OM
 - Classification tree analysis of model selection results, with default settings in R package 'rpart'. The complexity parameter is set at cp = 0.01, which represents the minimum classification improvement required for any split.

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'reg_tree_AIC_best_beta_fix.pdf') )
  ```

##  $\Delta$AIC and model probability - Classification tree
 - Classification tree analysis of for the difference in AIC among EMs, the rank of EMs, and the model probability of EMs (calculated from Akaike weights).

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'reg_tree_AIC_prob_beta_fix.pdf') )
  ```

## Results layout (example)
- Plots show median relative error (filled circle) with 2.5th and 97.5th percentiles
- OM factors as facets (top and right side)
- EM models on the x-axis
- Blue indicates EM=OM (diagonal of first 4 columns w/in facet blocks), incorrect EMs in red

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='65%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.all.yrs.plot_beta_fix.png') )
  ```

##  $\Delta$AIC  - Median and 95% probability interval
- Smaller $\Delta$AIC when $\sigma_{R}$ is larger

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'model_dAIC.ylim_beta_fix.png') )
  ```

##  Model probability  - Median and 95% probability interval
- True OMs receive more probability with $\sigma_{R}$=0.1 and high fishing contrast; otherwise, mean SRR received more probability

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'model_probability_beta_fix.png') )
  ```

## Assessment error - Recruitment (all years)
- Median unbiased, $\sigma_{R}$ determined quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.all.yrs.plot_beta_fix.png') )
  ```

## Assessment error - Recruitment (last 10 years)
- Median unbiased, $\sigma_{R}$ determined quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.last.10.plot_beta_fix.png') )
  ```

## Assessment error - Recruitment (terminal year)
- Median unbiased, $\sigma_{R}$ determined quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.last.yr.plot_beta_fix.png') )
  ```

## Assessment error - Spawning Stock Biomass (all years)
- Median unbiased, $\sigma_{R}$ determined quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.ssb.all.yrs.plot_beta_fix.png') )
  ```

## Assessment error - Spawning Stock Biomass (terminal year)
- Median unbiased, $\sigma_{R}$ determined quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.ssb.last.yr.plot_beta_fix.png') )
  ```

## Assessment error - Fbar (all years)
- Median unbiased, $\sigma_{R}$ determined quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.fbar.all.yrs.plot_beta_fix.png') )
  ```

## Assessment error - Fbar (terminal year)
- Median unbiased, $\sigma_{R}$ determined quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.fbar.last.yr.plot_beta_fix.png') )
  ```

## Effect size for Relative error in Recruitment

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'effect_size_glm_glmm_recr_re_beta_fix.pdf') )
```

## Effect size for Relative error in SSB

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'effect_size_glm_glmm_ssb_re_beta_fix.pdf') )
```

 ## Effect size for Relative error in Fbar

```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'effect_size_glm_glmm_fbar_re_beta_fix.pdf') )
```

# Alert
## Is everybody still awake?
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'presentation', 'asleep_keyboard.jpg') )
  ```

# Results

## Bias of estimated parameters - Ecov $\beta$

- Limiting and Masking had worst performance (BH_3 and BH_4)

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov_beta.plot_beta_fix.png') )
  ```

## Bias of estimated parameters - Ecov mean
- Larger quantile range when Ecov correlation is high (0.8)
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov.mean.plot_beta_fix.png') )
  ```

## Bias of estimated parameters - $\rho_{Ecov}$
- Well estimated when Ecov correlation was high (0.8)
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov.rho.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\sigma_{Ecov}$
- Positively biased when Ecov correlation is high (0.8)
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.Ecov.sigma.plot_beta_fix.png') )
  ```

## Bias of estimated parameters - SRR a
- Unbiased when $\sigma_{R}$ =0.1 and high fishing contrast; otherwise biased +/- 0.2 (BH_4 could be much worse)
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.mean_rec1.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - SRR b
- Unbiased when $\sigma_{R}$ =0.1 and high fishing contrast; otherwise biased +/- 0.2 (BH_4 could be much worse)

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.mean_rec2.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\sigma_{R}$
- Bias pattern mirroring between $\sigma_{R}$ and correlation
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.R_sigma.plot.ylim_beta_fix.png') )
  ```

## Bias of estimated parameters - $\rho_{R}$
- Bias pattern mirroring between $\sigma_{R}$ and correlation
- For most informative case (H-MSY, $\sigma_{R}$=0.1, low recruitment correlation), mean SRR has large positive bias
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'RE.R_cor.plot_beta_fix.png') )
  ```

## Mohn's $\rho$ - Recruitment (7 year peel)  
- Missing symbol indicates failed peels for one or more EM fits
- $\sigma_{R}$ and observation error were influential factors
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'recr.rho_beta_fix.png') )
  ```

## Mohn's $\rho$ - SSB  (7 year peel)   
- observation error was influential factor

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'ssb.rho_beta_fix.png') )
  ```

## Mohn's $\rho$ - Fbar  (7 year peel) 
- observation error was influential factor

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'fbar.rho_beta_fix.png') )
  ```

## Mohn's $\rho$ - Classification Tree for Recruitment, SSB, Fbar
-

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'reg_tree_Rho_beta_fix.pdf') )
  ```

## Mohn's $\rho$ - Random Effects on Recruitment (7 year peel)
- Negative bias when $\sigma_{R}$ is low
- Expected worse $\rho$ from incorrect EM, but was not the case

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'recr.dev.rho_beta_fix.png') )
  ```

## Mohn's $\rho$ - Classification Tree for Random Effects on Recruitment

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'reg_tree_Rdev_retro_beta_fix.pdf') )
  ```

## Projections: Specifications and Ecov assumptions 
- Projections were made from the 10th peel of each fitted model (Terminal year - 10)
- A 5-year average of maturity, weight at age, and selectivity specified (time-invariant anyway)
- F in the projection was set at the true OM F value
- The environmental index was treated 3 ways

    \begin{enumerate}
      \item A 5-year average of the most recent index value was held constant in the 10 year projection ('Avg5yr')
      \item The estimated process continued in the 10 year projection ('Continue')
      \item The observed environmental index was specified in the 10 year projection ('ObsEcov') 
    \end{enumerate}

- Projected Recruitment, SSB, and Catch were compared to the true OM values in the projection

## Projection performance - Relative Error Recruitment (first 3 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH
 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.proj.plot_beta_fix.png') )
    ```

## Projection performance - Relative Error Recruitment (all 10 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.rec.proj.plot_beta_fix.png') )
  ```

## Projection performance - Recruitment Uncertainty (first 3 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.rec.proj.plot3yr_beta_fix.png') )
    ```

## Projection performance - Recruitment Uncertainty (all 10 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.rec.proj.plot_beta_fix.png') )
  ```

## Projection performance - Relative Error SSB (first 3 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.ssb.proj.plot_beta_fix.png') )
    ```

## Projection performance - Relative Error SSB (all 10 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.ssb.proj.plot_beta_fix.png') )
  ```

## Projection performance - SSB Uncertainty (first 3 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.ssb.proj.plot3yr_beta_fix.png') )
    ```

## Projection performance - SSB Uncertainty (all 10 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.ssb.proj.plot_beta_fix.png') )
  ```

## Projection performance - Relative Error Catch (first 3 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.catch.proj.plot_beta_fix.png') )
    ```

## Projection performance - Relative Error Catch (all 10 years)
- Minimal bias (median at most -0.1 for mean SRR)
- Quantile range much larger for BH

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 're.catch.proj.plot_beta_fix.png') )
  ```

## Projection performance - Catch Uncertainty (first 3 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.catch.proj.plot3yr_beta_fix.png') )
    ```

## Projection performance - Catch Uncertainty (all 10 years)
- CV slightly lower for mean SRR, usually narrower quantile range

 ```{r,  echo=FALSE, message=FALSE, warnings=FALSE,  out.height='75%', fig.align='center'}
  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'sd.catch.proj.plot_beta_fix.png') )
  ```

## Projection Summary
- Minimal difference in median bias
- Recruitment is highly uncertain in projections, exacerbated by length of projection
- SSB and Catch also show increased uncertainty in longer projections
- Long projections are expected to be very imprecise, recommend shorter projections
- Tighter quantile range and slightly lower CV for mean SRR

<!-- ==================================================================================== -->

# Conclusions

## Take-aways

- Stock-recruitment model identification was poor, requiring unrealistic OM scenarios

- EMs with mean SRR and recruitment random effects generally performed similarly well as OM=EM, including for cases with an SR relationship and/or an ecov effect

- In conclusion, we recommend recruitment random effects from a mean SRR as the default EM in WHAM

# Future Work

## Recommendations for further analyses

- In WP-1, we recommended exploring additional levels of $\sigma_{R}$=0.5; however, we note that WP1-Appendix looked at $\sigma_{R}$=0.5 and found no difference

<!-- ==================================================================================== -->

---

Acknowledgements

- This work could not have been completed without support from the Microsoft Cooperative Research and Development Agreement (CRADA) and NOAA's National Cloud Program Office (OCIO), as well as MIT (... greg to fill in)
- We thank other members of the SSRTWG for thoughtful comments during earlier discussions and presentations of this work

```{r, echo=FALSE, message=FALSE, warnings=FALSE, out.width='35%', fig.align='center'}
 knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'presentation', 'abacus_crop.jpg') )
```

<!-- ==================================================================================== -->

<!-- ## R Markdown -->

<!-- If you have trouble knitting this file, check out this page: https://stackoverflow.com/questions/67696286/error-generating-pdf-using-knitr-to-pdf-in-rstudio -->

<!-- <!-- -- example chunk to include figure from file --> -->
<!-- ```{r, echo=FALSE, message=FALSE, warnings=FALSE, out.width='50%', fig.align='center'} -->
<!--  knitr::include_graphics(file.path(here::here(), 'Ecov_study','recruitment_functions', 'plots_beta_fix', 'model_probability_beta_fix.png') ) -->
<!-- ``` -->