### 2025-10-01 Project_0/manuscript/bib_text_marked.tex
saving work on project 0 revisons

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\protect\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-aeberhardetal18}
Aeberhard, W.H., Flemming, J.M., and Nielsen, A. 2018. Review of
{State}-{Space Models} for {Fisheries Science}. Annual Review of
Statistics and Its Application \textbf{5}(1): 215--235.
doi:\href{https://doi.org/10.1146/annurev-statistics-031017-100427}{10.1146/annurev-statistics-031017-100427}.

\bibitem[\citeproctext]{ref-augeretal16}
Auger-Méthé, M., Field, C., Albertsen, C.M., Derocher, A.E., Lewis,
M.A., Jonsen, I.D., and Mills Flemming, J. 2016. State-space models'
dirty little secrets: Even simple linear {G}aussian models can have
estimation problems. Scientific reports \textbf{6}(1): 26677.
doi:\href{https://doi.org/10.1038/srep26677}{10.1038/srep26677}.

\bibitem[\citeproctext]{ref-augeretal21}
{Auger-Méthé, M., Newman, K., Cole, D., Empacher, F., Gryba, R., King,
A.A., Leos-Barajas, V., Mills Flemming, J., Nielsen, A., Petris, G., and
others}. 2021. A guide to state--space modeling of ecological time
series. Ecological Monographs \textbf{91}(4): e01470.
doi:\href{https://doi.org/10.1002/ecm.1470}{10.1002/ecm.1470}.

\bibitem[\citeproctext]{ref-breimanetal84}
Breiman, L., Friedman, J.H., Olshen, R.A., and Stone, C.J. 1984.
Classification and regression trees. Chapman; Hall/CRC, New York, NY
USA.
doi:\href{https://doi.org/10.1201/9781315139470}{10.1201/9781315139470}.

\bibitem[\citeproctext]{ref-breiviketal23}
Breivik, O.N., Aldrin, M., Fuglebakk, E., and Nielsen, A. 2023.
Detecting significant retrospective patterns in state space fish stock
assessment. Canadian Journal of Fisheries and Aquatic Sciences
\textbf{80}(9): 1509--1518.
doi:\href{https://doi.org/10.1139/cjfas-2022-0250}{10.1139/cjfas-2022-0250}.

\bibitem[\citeproctext]{ref-brittenetalinreview}
Britten, G., Brooks, E.N., and Miller, T.J. In review. Identification
and performance of environmentally-driven stock-recruitment
relationships in state space assessment models. Canadian Journal of
Fisheries and Aquatic Sciences.

\bibitem[\citeproctext]{ref-brookslegault16}
Brooks, E.N., and Legault, C.M. 2016. Retrospective forecasting --
evaluating performance of stock projections for {N}ew {E}ngland
groundfish stocks. Canadian Journal of Fisheries and Aquatic Sciences
\textbf{73}(6): 935--950.
doi:\href{https://doi.org/10.1139/cjfas-2015-0163}{10.1139/cjfas-2015-0163}.

\bibitem[\citeproctext]{ref-cadigan16}
Cadigan, N.G. 2016. A state-space stock assessment model for northern
cod, including under-reported catches and variable natural mortality
rates. Canadian Journal of Fisheries and Aquatic Sciences
\textbf{73}(2): 296--308.
doi:\href{https://doi.org/10.1139/cjfas-2015-0047}{10.1139/cjfas-2015-0047}.

\bibitem[\citeproctext]{ref-carvalhoetal21}
Carvalho, F., Winker, H., Courtney, D., Kapur, M., Kell, L., Cardinale,
M., Schirripa, M., Kitakado, T., Yemane, D., Piner, K.R., Maunder, M.N.,
Taylor, I., Wetzel, C.R., Doering, K., Johnson, K.F., and Methot, R.D.
2021. A cookbook for using model diagnostics in integrated stock
assessments. Fisheries Research \textbf{240}: 105959.
doi:\url{https://doi.org/10.1016/j.fishres.2021.105959}.

\bibitem[\citeproctext]{ref-clopperpearson34}
Clopper, C.J., and Pearson, E.S. 1934. The use of confidence or fiducial
limits illustrated in the case of the binomial. Biometrika
\textbf{26}(4): 404--413.
doi:\href{https://doi.org/10.1093/biomet/26.4.404}{10.1093/biomet/26.4.404}.

\bibitem[\citeproctext]{ref-collieretal22}
Collier, Z.K., Zhang, H., and Soyoye, O. 2022. Alternative methods for
interpreting {M}onte {C}arlo experiments. Communications in Statistics -
Simulation and Computation: 1--16.
doi:\href{https://doi.org/10.1080/03610918.2022.2082474}{10.1080/03610918.2022.2082474}.

\bibitem[\citeproctext]{ref-connetal10}
Conn, P.B., Williams, E.H., and Shertzer, K.W. 2010. When can we
reliably estimate the productivity of fish stocks? Canadian Journal of
Fisheries and Aquatic Sciences \textbf{67}(3): 511--523.
doi:\href{https://doi.org/10.1139/F09-194}{10.1139/F09-194}.

\bibitem[\citeproctext]{ref-croninpunt21}
Cronin-Fine, L., and Punt, A.E. 2021. Modeling time-varying selectivity
in size-structured assessment models. Fisheries Research \textbf{239}:
105927. Elsevier.

\bibitem[\citeproctext]{ref-devalpinehastings02}
{de Valpine, P., and Hastings, A.} 2002. Fitting population models
incorporating process noise and observation error. Ecological Monographs
\textbf{72}(1): 57--76.

\bibitem[\citeproctext]{ref-fischetal23}
Fisch, N., Shertzer, K., Camp, E., Maunder, M., and Ahrens, R. 2023.
Process and sampling variance within fisheries stock assessment models:
Estimability, likelihood choice, and the consequences of incorrect
specification. ICES Journal of Marine Science \textbf{80}(8):
2125--2149.
doi:\href{https://doi.org/10.1093/icesjms/fsad138}{10.1093/icesjms/fsad138}.

\bibitem[\citeproctext]{ref-fleischmanetal13}
Fleischman, S.J., Catalano, M.J., Clark, R.A., and Bernard, D.R. 2013.
An age-structured state-space stock--recruit model for {P}acific salmon
(\emph{{O}ncorhynchus spp}.). Canadian Journal of Fisheries and Aquatic
Sciences \textbf{70}(3): 401--414.
doi:\href{https://doi.org/10.1139/cjfas-2012-0112}{10.1139/cjfas-2012-0112}.

\bibitem[\citeproctext]{ref-gonzalezetal18}
Gonzalez, O., O'Rourke, H.P., Wurpts, I.C., and Grimm, K.J. 2018.
Analyzing {M}onte {C}arlo simulation studies with classification and
regression trees. Structural Equation Modeling: A Multidisciplinary
Journal \textbf{25}(3): 403--413.
doi:\href{https://doi.org/10.1080/10705511.2017.1369353}{10.1080/10705511.2017.1369353}.

\bibitem[\citeproctext]{ref-harwelletal18}
Harwell, M., Kohli, N., and Peralta-Torres, Y. 2018. A survey of
reporting practices of computer simulation studies in statistical
research. The American Statistician \textbf{72}(4): 321--327.
doi:\href{https://doi.org/10.1080/00031305.2017.1342692}{10.1080/00031305.2017.1342692}.

\bibitem[\citeproctext]{ref-hoenigetal25}
Hoenig, J.M., Hearn, W.S., Leigh, G.M., and Latour, R.J. 2025.
Principles for estimating natural mortality rate. Fisheries Research
\textbf{281}: 107195.
doi:\href{https://doi.org/10.1016/j.fishres.2024.107195}{10.1016/j.fishres.2024.107195}.

\bibitem[\citeproctext]{ref-hoyleetal22}
Hoyle, S.D., Maunder, M.N., Punt, A.E., Mace, P.M., Devine, J.A., and
A???mar, Z.T. 2022. Preface: Developing the next generation of stock
assessment software. Fisheries Research \textbf{246}: 106176.
doi:\href{https://doi.org/10.1016/j.fishres.2021.106176}{10.1016/j.fishres.2021.106176}.

\bibitem[\citeproctext]{ref-hurtadoferroetal15}
Hurtado-Ferro, F., Szuwalski, C.S., Valero, J.L., Anderson, S.C.,
Cunningham, C.J., Johnson, K.F., Licandeo, R., McGilliard, C.R.,
Monnahan, C.C., Muradian, M.L., Ono, K., Vert-Pre, K.A., Whitten, A.R.,
and Punt, A.E. 2014. Looking in the rear-view mirror: Bias and
retrospective patterns in integrated, age-structured stock assessment
models. ICES Journal of Marine Science \textbf{72}(1): 99--110.
doi:\href{https://doi.org/10.1093/icesjms/fsu198}{10.1093/icesjms/fsu198}.

\bibitem[\citeproctext]{ref-huynhetal22}
Huynh, Q.C., Legault, C.M., Hordyk, A.R., and Carruthers, T.R. 2022. A
closed-loop simulation framework and indicator approach for evaluating
impacts of retrospective patterns in stock assessments. ICES Journal of
Marine Science \textbf{79}(7): 2003--2016.
doi:\href{https://doi.org/10.1093/icesjms/fsac066}{10.1093/icesjms/fsac066}.

\bibitem[\citeproctext]{ref-johnsonetal16}
Johnson, K.F., Councill, E., Thorson, J.T., Brooks, E., Methot, R.D.,
and Punt, A.E. 2016. Can autocorrelated recruitment be estimated using
integrated assessment models and how does it affect population
forecasts? Fisheries Research \textbf{183}: 222--232.
doi:\href{https://doi.org/10.1016/j.fishres.2016.06.004}{10.1016/j.fishres.2016.06.004}.

\bibitem[\citeproctext]{ref-kapuretal24}
Kapur, M.S., Ducharme-Barth, N., Oshima, M., and Carvalho, F. 2025. Good
practices, trade-offs, and precautions for model diagnostics in
integrated stock assessments. Fisheries Research \textbf{281}: 107206.
doi:\href{https://doi.org/10.1016/j.fishres.2024.107206}{10.1016/j.fishres.2024.107206}.

\bibitem[\citeproctext]{ref-kassraftery95}
Kass, R.E., and Raftery, A.E. 1995. Bayes factors. Journal of the
American Statistical Association \textbf{90}(430): 773--795.
doi:\href{https://doi.org/10.1080/01621459.1995.10476572}{10.1080/01621459.1995.10476572}.

\bibitem[\citeproctext]{ref-katz81}
Katz, R.W. 1981. On some criteria for estimating the order of a {M}arkov
chain. Technometrics \textbf{23}(3): 243--249.
doi:\href{https://doi.org/10.1080/00401706.1981.10486293}{10.1080/00401706.1981.10486293}.

\bibitem[\citeproctext]{ref-knape08}
Knape, J. 2008. Estimability of density dependence in models of time
series data. Ecology \textbf{89}(11): 2994--3000.
doi:\href{https://doi.org/10.1890/08-0071.1}{10.1890/08-0071.1}.

\bibitem[\citeproctext]{ref-kristensenetal16}
Kristensen, K., Nielsen, A., Berg, C.W., Skaug, H., and Bell, B.M. 2016.
{TMB}: Automatic differentiation and {L}aplace approximation. Journal of
Statistical Software \textbf{70}(5): 1--21.
doi:\href{https://doi.org/10.18637/jss.v070.i05}{10.18637/jss.v070.i05}.

\bibitem[\citeproctext]{ref-leeetal11}
Lee, H.-H., Maunder, M.N., Piner, K.R., and Methot, R.D. 2011.
Estimating natural mortality within a fisheries stock assessment model:
An evaluation using simulation analysis based on twelve stock
assessments. Fisheries Research \textbf{109}(1): 89--94.
doi:\href{https://doi.org/10.1016/j.fishres.2011.01.021}{10.1016/j.fishres.2011.01.021}.

\bibitem[\citeproctext]{ref-legault09}
Legault, C.M. 2009. Report of the retrospective working group, 14-16
january 2008. US Department of Commerce Northeast Fisheries Science
Center Reference Document 09-01. US Department of Commerce Northeast
Fisheries Science Center. Woods Hole, MA.

\bibitem[\citeproctext]{ref-legaultpalmer16}
Legault, C.M., and Palmer, M.C. 2016. In what direction should the
fishing mortality target change when natural mortality increases within
an assessment? Canadian Journal of Fisheries and Aquatic Sciences
\textbf{73}(3): 349--357.
doi:\href{https://doi.org/10.1139/cjfas-2015-0232}{10.1139/cjfas-2015-0232}.

\bibitem[\citeproctext]{ref-legaultrestrepo99}
Legault, C.M., and Restrepo, V.R. 1999. A flexible forward
age-structured assessment program. Col. Vol. Sci. Pap. ICCAT
\textbf{49}(2): 246--253.

\bibitem[\citeproctext]{ref-legaultetal23}
Legault, C.M., Wiedenmann, J., Deroba, J.J., Fay, G., Miller, T.J.,
Brooks, E.N., Bell, R.J., Langan, J.A., Cournane, J.M., Jones, A.W., and
Muffley, B. 2023. Data-rich but model-resistant: An evaluation of
data-limited methods to manage fisheries with failed age-based stock
assessments. Canadian Journal of Fisheries and Aquatic Sciences
\textbf{80}(1): 27--42.
doi:\href{https://doi.org/10.1139/cjfas-2022-0045}{10.1139/cjfas-2022-0045}.

\bibitem[\citeproctext]{ref-lietalinreview_a}
Li, C., Deroba, J.J., Berger, A.M., Goethel, D.R., Langseth, B.J.,
Schueller, A.M., and Miller, T.J. In review. Random effects on
numbers-at-age transitions implictly account for movement dynamics and
improve performance within a state-space stock assessment. Canadian
Journal of Fisheries and Aquatic Sciences.

\bibitem[\citeproctext]{ref-lietalinreview}
Li, C., Deroba, J.J., Miller, T.J., Legault, C.M., and Perretti, C. In
review. Guidance on bias-correction of log-normal random effects and
observations in state-space assessment models. Canadian Journal of
Fisheries and Aquatic Sciences.

\bibitem[\citeproctext]{ref-lietal24}
Li, C., Deroba, J.J., Miller, T.J., Legault, C.M., and Perretti, C.T.
2024. An evaluation of common stock assessment diagnostic tools for
choosing among state-space models with multiple random effects
processes. Fisheries Research \textbf{273}: 106968.
doi:\href{https://doi.org/10.1016/j.fishres.2024.106968}{10.1016/j.fishres.2024.106968}.

\bibitem[\citeproctext]{ref-liljestrandetal24}
Liljestrand, E.M., Bence, J.R., and Deroba, J.J. 2024. The effect of
process variability and data quality on performance of a state-space
stock assessment model. Fisheries Research \textbf{275}: 107023.
doi:\href{https://doi.org/10.1016/j.fishres.2024.107023}{10.1016/j.fishres.2024.107023}.

\bibitem[\citeproctext]{ref-mackinnonetal95}
MacKinnon, D.P., Warsi, G., and Dwyer, J.H. 1995. A simulation study of
mediated effect measures. Multivariate Behavioral Research
\textbf{30}(1): 41--62.
doi:\href{https://doi.org/10.1207/s15327906mbr3001/_3}{10.1207/s15327906mbr3001\textbackslash\_3}.

\bibitem[\citeproctext]{ref-magnussonhilborn07}
Magnusson, A., and Hilborn, R. 2007. What makes fisheries data
informative? Fish and Fisheries \textbf{8}(4): 337--358.
doi:\href{https://doi.org/10.1111/j.1467-2979.2007.00258.x}{10.1111/j.1467-2979.2007.00258.x}.

\bibitem[\citeproctext]{ref-methotwetzel13}
Methot, R.D., and Wetzel, C.R. 2013. Stock synthesis: A biological and
statistical framework for fish stock assessment and fishery management.
Fisheries Research \textbf{142}: 86--99.
doi:\href{https://doi.org/10.1016/j.fishres.2012.10.012}{10.1016/j.fishres.2012.10.012}.

\bibitem[\citeproctext]{ref-millerbrooks21}
Miller, T.J., and Brooks, E.N. 2021. Steepness is a slippery slope. Fish
and Fisheries \textbf{22}(3): 634--645.
doi:\href{https://doi.org/10.1111/faf.12534}{10.1111/faf.12534}.

\bibitem[\citeproctext]{ref-milleretal16}
Miller, T.J., Hare, J.A., and Alade, L. 2016. A state-space approach to
incorporating environmental effects on recruitment in an age-structured
assessment model with an application to {S}outhern {N}ew {E}ngland
yellowtail flounder. Canadian Journal of Fisheries and Aquatic Sciences
\textbf{73}(8): 1261--1270.
doi:\href{https://doi.org/10.1139/cjfas-2015-0339}{10.1139/cjfas-2015-0339}.

\bibitem[\citeproctext]{ref-millerhyun18}
Miller, T.J., and Hyun, S.-Y. 2018. Evaluating evidence for alternative
natural mortality and process error assumptions using a state-space,
age-structured assessment model. Canadian Journal of Fisheries and
Aquatic Sciences \textbf{75}(5): 691--703.
doi:\href{https://doi.org/10.1139/cjfas-2017-0035}{10.1139/cjfas-2017-0035}.

\bibitem[\citeproctext]{ref-millerlegault17}
Miller, T.J., and Legault, C.M. 2017. Statistical behavior of
retrospective patterns and their effects on estimation of stock and
harvest status. Fisheries Research \textbf{186}: 109--120.
doi:\href{https://doi.org/10.1016/j.fishres.2016.08.002}{10.1016/j.fishres.2016.08.002}.

\bibitem[\citeproctext]{ref-milleretal18}
Miller, T.J., O'Brien, L., and Fratantoni, P.S. 2018. Temporal and
environmental variation in growth and maturity and effects on management
reference points of {G}eorges {B}ank {A}tlantic cod. Canadian Journal of
Fisheries and Aquatic Sciences \textbf{75}(12): 2159--2171.
doi:\href{https://doi.org/10.1139/cjfas-2017-0124}{10.1139/cjfas-2017-0124}.

\bibitem[\citeproctext]{ref-millerstock20}
Miller, T.J., and Stock, B.C. 2020. The {Woods Hole Assessment Model}
({WHAM}). Available from \url{https://timjmiller.github.io/wham/}.

\bibitem[\citeproctext]{ref-mohn99}
Mohn, R. 1999. The retrospective problem in sequential population
analysis: An investigation using cod fishery and simulated data. ICES
Journal of Marine Science \textbf{56}(4): 473--488.
doi:\href{https://doi.org/10.1006/jmsc.1999.0481}{10.1006/jmsc.1999.0481}.

\bibitem[\citeproctext]{ref-nefsc22}
NEFSC. 2022a. Final report of the haddock research track assessment
working group. {Available} at
https://s3.us-east-1.amazonaws.com/nefmc.org/14b\_EGB\_Research\_Track\_Haddock\_WG\_Report\_DRAFT.pdf.

\bibitem[\citeproctext]{ref-nefsc22a}
NEFSC. 2022b. Report of the {A}merican plaice research track working
group. {Available} at
https://s3.us-east-1.amazonaws.com/nefmc.org/2\_American-Plaice-WG-Report.pdf.

\bibitem[\citeproctext]{ref-nefsc24}
NEFSC. 2024. Butterfish research track assessment report. US Dept Commer
Northeast Fish Sci Cent Ref Doc. 24-03; 191 p.

\bibitem[\citeproctext]{ref-nefsc25}
NEFSC. 2025. Yellowttail flounder research track working group report.
{Available} at
https://d23h0vhsm26o6d.cloudfront.net/10c.-Yellowtail-Flounder-RT-WG-Report.pdf.

\bibitem[\citeproctext]{ref-nielsenberg14}
Nielsen, A., and Berg, C.W. 2014. Estimation of time-varying selectivity
in stock assessments using state-space models. Fisheries Research
\textbf{158}: 96--101.
doi:\href{https://doi.org/10.1016/j.fishres.2014.01.014}{10.1016/j.fishres.2014.01.014}.

\bibitem[\citeproctext]{ref-pedersenberg17}
Pedersen, M.W., and Berg, C.W. 2017. A stochastic surplus production
model in continuous time. Fish and Fisheries \textbf{18}(2): 226--243.
doi:\href{https://doi.org/10.1111/faf.12174}{10.1111/faf.12174}.

\bibitem[\citeproctext]{ref-perrettietal20}
Perretti, C.T., Deroba, J.J., and Legault, C.M. 2020. Simulation testing
methods for estimating misreported catch in a state-space stock
assessment model. ICES Journal of Marine Science \textbf{77}(3):
911--920.
doi:\href{https://doi.org/10.1093/icesjms/fsaa034}{10.1093/icesjms/fsaa034}.

\bibitem[\citeproctext]{ref-polanskyetal09}
Polansky, L., De Valpine, P., Lloyd-Smith, J.O., and Getz, W.M. 2009.
Likelihood ridges and multimodality in population growth rate models.
Ecology \textbf{90}(8): 2313--2320.
doi:\href{https://doi.org/10.1890/08-1461.1}{10.1890/08-1461.1}.

\bibitem[\citeproctext]{ref-dupontaviceetal22}
Pontavice, H. du, Miller, T.J., Stock, B.C., Chen, Z., and Saba, V.S.
2022. Ocean model-based covariates improve a marine fish stock
assessment when observations are limited. ICES Journal of Marine Science
\textbf{79}(4): 1259--1273.
doi:\href{https://doi.org/10.1093/icesjms/fsac050}{10.1093/icesjms/fsac050}.

\bibitem[\citeproctext]{ref-punt23}
Punt, A.E. 2023. Those who fail to learn from history are condemned to
repeat it: A perspective on current stock assessment good practices and
the consequences of not following them. Fisheries Research \textbf{261}:
106642.
doi:\href{https://doi.org/10.1016/j.fishres.2023.106642}{10.1016/j.fishres.2023.106642}.

\bibitem[\citeproctext]{ref-puntetal14}
Punt, A.E., Hurtado-Ferro, F., and Whitten, A.R. 2014. Model selection
for selectivity in fisheries stock assessments. Fisheries Research
\textbf{158}: 124--134.
doi:\href{https://doi.org/10.1016/j.fishres.2013.06.003}{10.1016/j.fishres.2013.06.003}.

\bibitem[\citeproctext]{ref-ruddthorson18}
Rudd, M.B., and Thorson, J.T. 2018. Accounting for variable recruitment
and fishing mortality in length-based stock assessments for data-limited
fisheries. Canadian Journal of Fisheries and Aquatic Sciences
\textbf{75}(7): 1019--1035.
doi:\href{https://doi.org/10.1139/cjfas-2017-0143}{10.1139/cjfas-2017-0143}.

\bibitem[\citeproctext]{ref-shibata76}
Shibata, R. 1976. Selection of the order of an autoregressive model by
{A}kaike's information criterion. Biometrika \textbf{63}(1): 117--126.
doi:\href{https://doi.org/10.1093/biomet/63.1.117}{10.1093/biomet/63.1.117}.

\bibitem[\citeproctext]{ref-stewartmonnahan17}
Stewart, I.J., and Monnahan, C.C. 2017. Implications of process error in
selectivity for approaches to weighting compositional data in fisheries
stock assessments. Fisheries Research \textbf{192}: 126--134.
doi:\href{https://doi.org/10.1016/j.fishres.2016.06.018}{10.1016/j.fishres.2016.06.018}.

\bibitem[\citeproctext]{ref-stockmiller21}
Stock, B.C., and Miller, T.J. 2021. The {W}oods {H}ole {A}ssessment
{M}odel ({WHAM}): {A} general state-space assessment framework that
incorporates time- and age-varying processes via random effects and
links to environmental covariates. Fisheries Research \textbf{240}:
105967.
doi:\href{https://doi.org/10.1016/j.fishres.2021.105967}{10.1016/j.fishres.2021.105967}.

\bibitem[\citeproctext]{ref-stocketal21}
Stock, B.C., Xu, H., Miller, T.J., Thorson, J.T., and Nye, J.A. 2021.
{Implementing two-dimensional autocorrelation in either survival or
natural mortality improves a state-space assessment model for Southern
{N}ew {E}ngland-Mid {A}tlantic yellowtail flounder}. Fisheries Research
\textbf{237}: 105873.
doi:\href{https://doi.org/10.1016/j.fishres.2021.105873}{10.1016/j.fishres.2021.105873}.

\bibitem[\citeproctext]{ref-szuwalskietal18}
Szuwalski, C.S., Ianelli, J.N., and Punt, A.E. 2018. Reducing
retrospective patterns in stock assessment and impacts on management
performance. ICES Journal of Marine Science \textbf{75}(2): 596--609.
doi:\href{https://doi.org/10.1093/icesjms/fsx159}{10.1093/icesjms/fsx159}.

\bibitem[\citeproctext]{ref-thorsonminto15}
Thorson, J.T., and Minto, C. 2015. Mixed effects: A unifying framework
for statistical modelling in fisheries biology. ICES Journal of Marine
Science \textbf{72}(5): 1245--1256.
doi:\href{https://doi.org/10.1093/icesjms/fsu213}{10.1093/icesjms/fsu213}.

\bibitem[\citeproctext]{ref-thulin14}
Thulin, M. 2014. {The cost of using exact confidence intervals for a
binomial proportion}. Electronic Journal of Statistics \textbf{8}(1):
817--840.
doi:\href{https://doi.org/10.1214/14-EJS909}{10.1214/14-EJS909}.

\bibitem[\citeproctext]{ref-trijouletetal20}
Trijoulet, V., Fay, G., and Miller, T.J. 2020. Performance of a
state-space multispecies model: What are the consequences of ignoring
predation and process errors in stock assessments? Journal of Applied
Ecology \textbf{57}(1): 121--135.
doi:\href{https://doi.org/10.1111/1365-2664.13515}{10.1111/1365-2664.13515}.

\bibitem[\citeproctext]{ref-wangetal17}
Wang, S., Cadigan, N.G., and Benoît, H.P. 2017. Inference about
regression parameters using highly stratified survey count data with
over-dispersion and repeated measurements. Journal of Applied Statistics
\textbf{44}(6): 1013--1030.
doi:\href{https://doi.org/10.1080/02664763.2016.1191622}{10.1080/02664763.2016.1191622}.

\bibitem[\citeproctext]{ref-wiedenmannetal19}
Wiedenmann, J., Free, C.M., and Jensen, O.P. 2019. Evaluating the
performance of data-limited methods for setting catch targets through
application to data-rich stocks: A case study using northeast {U.S.}
Fish stocks. Fisheries Research \textbf{209}(1): 129--142.
doi:\href{https://doi.org/10.1016/j.fishres.2018.09.018}{10.1016/j.fishres.2018.09.018}.

\bibitem[\citeproctext]{ref-xuetal19}
Xu, H., Thorson, J.T., Methot, R.D., and Taylor, I.G. 2019. A new
semi-parametric method for autocorrelated age- and time-varying
selectivity in age-structured assessment models. Canadian Journal of
Fisheries and Aquatic Sciences \textbf{76}(2): 268--285.
doi:\href{https://doi.org/10.1139/cjfas-2017-0446}{10.1139/cjfas-2017-0446}.

\bibitem[\citeproctext]{ref-yee08}
Yee, T.W. 2008. The {VGAM} package. R News \textbf{8}(2): 28--39.
Available from
\url{https://journal.r-project.org/articles/RN-2008-014/}.

\bibitem[\citeproctext]{ref-yee15}
Yee, T.W. 2015. Vector generalized linear and additive models: With an
implementation in {R}. Springer, New York, NY USA.
doi:\href{https://doi.org/10.1007/978-1-4939-2818-7}{10.1007/978-1-4939-2818-7}.

\end{CSLReferences}