### 2025-10-01 Project_0/manuscript/manuscript_R1.Rmd

---
title: An investigation of factors affecting inferences from and reliability of state-space age-structured assessment models
author: 
  - Timothy J. Miller^1,2^
  - Greg Britten^3^
  - Elizabeth N. Brooks^2^
  - Gavin Fay^4^
  - Alex Hansell^2^
  - Christopher M. Legault^2^
  - Chengxue Li^2^
  - Brandon Muffley^5^
  - Brian C. Stock^6^
  - John Wiedenmann^7^
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontsize: 12pt
output:
  pdf_document:
#    citation_package: natbib
    fig_caption: yes
    includes:
      in_header:
      - options.sty
      - journalshortcuts.tex
    keep_tex: yes
    number_sections: yes
  html_document:
    df_print: paged
#  word_document: default
header-includes:
#- \usepackage{makecell}
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
bibliography: manuscript.bib
#biblio-style: cjfas2.bst
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE, fig.align = "center")
#knitr::opts_knit$set(root.dir = '../' )
library(knitr)
library(tidyverse)
library(pander)
library(kableExtra)
library(png)
library(forcats)
type <- "latex"
if(knitr::is_latex_output()) type <- "latex"
if(knitr::is_html_output()) type <- "html"
```

$^1$corresponding author: timothy.j.miller@noaa.gov\
$^2$Northeast Fisheries Science Center, Woods Hole Laboratory, 166 Water Street, Woods Hole, MA 02543 USA\
$^3$Biology Department, Woods Hole Oceanographic Institution, 266 Woods Hole Rd. Woods Hole, MA, USA\
$^4$Department of Fisheries Oceanography, School for Marine Science and Technology, University of Massachusetts Dartmouth, 836 S Rodney French Boulevard, New Bedford, MA 02740, USA\
$^5$Mid-Atlantic Fishery Management Council, 800 North State Street, Suite 201, Dover, DE 19901 USA\
$^6$Institute of Marine Research, Nye Flødevigveien 20, 4817 His, Norway\
$^7$Department of Ecology, Evolution, and Natural Resources. Rutgers University\

\pagebreak

**keywords:** stock assessment, state-space, model selection, bias, convergence, retrospective patterns

## Abstract {-}

State-space models are increasingly used for stock assessment, and evaluations of statistical inferences made from them is needed. We simulated 72 operating models that varied fishing pressure and observation error across process errors in recruitment, apparent survival, selectivity, catchability, and/or natural mortality. We fit estimating models with different assumptions on the process error source and whether natural mortality and stock-recruit relationship were estimated. We measured reliability of estimating models by probability of convergence, accuracy of AIC for correct process error configuration and stock-recruit function, estimation bias, and magnitude of retrospective patterns. Across all measures reliability was generally better with lower observation error, contrast in fishing pressure over time, and median natural mortality rate known, but using marginal AIC to distinguishing the stock recruit relationship further required large contrast in spawning biomass and low residual variation in recruitment. Marginal AIC accurately distinguished process errors on recruitment, apparent survival, and selectivity.  These results improve our understanding of when results from state space assessment models, the next-generation of fisheries stock assessment, can be expected to be reliable.

\pagebreak

# Introduction {-}

Application of state-space models in fisheries stock assessment and management has expanded dramatically within International Council for the Exploration of the Sea (ICES), Canada, and the Northeast US [@nielsenberg14; @cadigan16; @pedersenberg17; @stockmiller21]. State-space models latent population characteristics as statistical time series with periodic observations that also may have error due to sampling or other sources of measurement error. Traditional assessment models may use state-space approaches to account for temporal variability in population characteristics [@methotwetzel13;@legaultrestrepo99], but these models treat the annual parameters as penalized fixed effects parameters where the variance parameters controlling the penalties are assumed known [@thorsonminto15]. Modern state-space models can estimate the annually varying parameters as random effects with variance parameters estimated using maximum marginal likelihood or corresponding Bayesian approaches. These latter approaches are considered best practice and a recommended for the next generation of stock assessment models [@hoyleetal22; @punt23].

State-space stock assessment models, with nonlinear functions of latent parameters and multiple types of observations with varying distributional assumptions, are one of the most complex examples of this analytical approach. Statistical aspects of state-space models and their application within fisheries have been studied extensively, but previous work has focused primarily on linear and Gaussian state-space models [@aeberhardetal18; @augeretal21]. Therefore, current understanding of the reliability of state-space models does not extend to usage for stock assessment.

As state-space models provide greater flexibility by allowing multiple processes to vary as random effects [@nielsenberg14; @aeberhardetal18; @stocketal21], one of the most immediate questions regards the implications of mis-specification among alternative sources of process error. Incorrect treatment of population attributes as temporally varying [@trijouletetal20; @liljestrandetal24] could lead to misidentification of stock status and biased population estimates, ultimately impacting fisheries management decisions [@legaultpalmer16; @szuwalskietal18; @croninpunt21]. Furthermore, biological, fishery, and observational processes are often confounded in catch-at-age data, which may adversely affect ability to distinguish between true process variability and observational error [@puntetal14; @stewartmonnahan17; @croninpunt21; @fischetal23; @lietalinreview_a].

@lietal24 conducted a full-factorial simulation-estimation study to assess model reliability when confounding random-effects processes (numbers-at-age, fishery selectivity, and natural mortality) were included. Their results suggest that while state-space models can generally identify sources of process error, overly complex models, even when misspecified (i.e., incorporating process error that did not exist in reality), often performed similarly to correctly specified models, with little to no bias in key management quantities. Similarly, @liljestrandetal24 found little downside in assuming process error in recruitment or selectivity, even when it was absent.

Despite mounting efforts, several limitations remain. First, confounding processes that can be treated as random effects in the model were not thoroughly examined or tested within a simulation-estimation framework. Second, previous studies relied on operating models conditioned on specific fisheries, limiting their generalizability [@lietalinreview_a; @liljestrandetal24]. In particular, the effects of observation error and underlying fishing history have not been fully isolated in simulation study designs, making it challenging to disentangle the interplay between process and observation error magnitudes, as demonstrated in @fischetal23. Third, explicitly modeling stock-recruit relationships (SRRs) as mechanistic drivers of population dynamics is promising [@fleischmanetal13; @dupontaviceetal22], but reliability of inferences within integrated state-space age-structured models has not been evaluated. Evidence from other studies suggests that when both process and observation errors are unknown, estimating density dependence parameters becomes highly uncertain [@knape08; @polanskyetal09]. In particular, @knape08 demonstrated that stronger density dependence becomes increasingly difficult to estimate in the presence of observation error. Therefore, it is crucial to assess whether density dependence mechanisms can be estimated with sufficient precision for use in fisheries management [@augeretal16]. Finally, although the importance of autocorrelation in process errors is recognized, investigations of the ability to distinguish state-space assessment models with and without autocorrelation and whether such misspecification is detrimental to estimation of important population metrics are lacking [@johnsonetal16;@xuetal19].

In the present study, we conduct a simulation study with operating models (OMs) varying by degree of observation error, source and variability of process error, and fishing history. The simulations from these OMs are fitted with estimation models (EMs) that make alternative assumptions for sources of process error, whether a SRR was estimated, and whether natural mortality is estimated. Given the confounding nature of process errors, developing diagnostic tools to detect model misspecification is of great scientific interest and could aid the next generation of stock assessments [@augeretal21]. We evaluate whether convergence and Akaike Information Criterion (AIC) can correctly determine the source of process error and the existence of a SRR. We also evaluate when retrospective patterns occur and the degree of bias in outputs of the assessment model that are important for management.

# Methods {-}

We used the Woods Hole Assessment Model (WHAM) to configure OMs and EMs in our simulation study [@millerstock20;@stockmiller21]. WHAM is an R package freely available via a github repository and is built on the Template Model Builder package [@kristensenetal16]. For this study we used [version 1.0.6.9000, commit 77bbd94](https://github.com/timjmiller/wham/tree/77bbd946e4881216a439933473d1c58b21c270c3). WHAM has also been used to configure OMs and EMs for closed loop simulations evaluating index-based assessment methods [@legaultetal23] and is currently used or accepted for use in management of numerous NEUS fish stocks [e.g., @nefsc22;@nefsc22a;@nefsc24].

We completed a simulation study with a number of OMs that can be categorized based on where process error random effects were assumed: recruitment (R, assumed present in all models), apparent survival (denoted R+S), natural mortality (R+M), fleet selectivity (R+Sel), or index catchability (R+q). We refer to the (R+S) OMs as modeling apparent survival because on logscale the random effects ($\epsilon_{a,y}$) are additive to the total mortality (F+M) between numbers at age, thus they modify the survival term. However, as @stockmiller21 note, these random effects can be due to events other than mortality, such as immigration, emigration, missreported catch, and other sources of misspecification.  For each OM, assumptions about the magnitude of the variance of process errors and observations are required and the values we used were based on a review of the range of estimates from Northeast Unite States (NEUS) assessments using WHAM.

In total, we configured 72 OMs with alternative assumptions about the source and magnitude of process errors, magnitude of observation error in indices and age composition data, and contrast in fishing pressure over time. We fitted 20 EMs to observations generated from each of 100 simulations where process errors were also simulated. Each EM differed in assumptions about the source of process errors, whether natural mortality (or the median for models with process error in natural mortality) was estimated, and whether a Beverton-Holt SRR was estimated within the EM. Details of each of the OMs and EMs are described below.

We did not use the log-normal bias-correction feature for process errors or observations described by [@stockmiller21] for OMs and EMs to simplify interpretation of the study results [@lietalinreview]. All code we used to perform the simulation study and summarize results can be found at https://github.com/timjmiller/SSRTWG/tree/main/Project_0/code.

## Operating models {-}

### Population {-}

The population consists of 10 age classes, ages 1 to 10+, with the last being a plus group that accumulates ages 10 and older. We assume spawning occurs annually 1/4 of the way through the year. The maturity at age was a logistic curve with $a_{50}$ = 2.89 and slope = 0.88 (Figure \ref{om_inputs_fig}, top left).

Weight at age was generated with a von Bertalanffy growth function
$$
L_a = L_{\infty}\left(1 - e^{-k(a - t_0)}\right)
$$
where $t_0 = 0$, $L_\infty = 85$, and $k = 0.3$, and a L-W relationship such that 
$$
W_a = \theta_1 L_a^{\theta_2}
$$
where $\theta_1 = e^{-12.1}$ and $\theta_2 = 3.2$ (Figure \ref{om_inputs_fig}, top right).

We assumed a Beverton-Holt SRR with constant pre-recruit mortality parameters for all OMs. All biological inputs to calculations of spawning biomass per recruit (i.e., weight, maturity, and natural mortality at age) are constant in the apparent survival (R+S) selectivity (R+Sel), and survey catchability (R+q) process error OMs. Therefore, steepness and unfished recruitment are also constant over the time period for those OMs [@millerbrooks21]. We specified unfished recruitment equal to $e^{10}$ and $\Fmsy = F_{40\%} = 0.348$, which equates to a steepness of 0.69 and $a=0.60$ and $b = 2.4 \times 10^{-5}$ for the Beverton-Holt parameterization
\[
N_{1,y} = \frac{a \text{SSB}_{y-1}}{1 + b \text{SSB}_{y-1}} 
\]
(Figure \ref{om_inputs_fig}, bottom right). We assumed a value of 0.2 for the natural mortality rate in OMs without process errors on natural mortaliyt and for the median rate for OMs with process errors on natural mortality.

We used two fishing scenarios for OMs. In the first scenario, the stock experiences overfishing at 2.5\Fmsy for the first 20 years followed by fishing at \Fmsy for the last 20 years (denoted $2.5\Fmsy \rightarrow \Fmsy$). In the second scenario, the stock is fished at \Fmsy for the entire time period (40 years). The magnitude of the overfishing assumptions is based on average estimates of overfishing for NEUS groundfish stocks from @wiedenmannetal19 and similar to the approach in @legaultetal23.

We specified initial population abundance at age at the equilibrium distribution that corresponds to fishing at either $F = 2.5\times \Fmsy$ or $F = \Fmsy$. This implies that, for a deterministic model, the abundance at age would not change from the first year to the next.

For OMs with time-varying random effects for M , steepness is not constant. However, we used the same $a$  and $b$ parameters as other OMs, which equates to a steepness and R0 at the median of the time series process for M. For OMs with time-varying random effects for fishery selectivity, \Fmsy is also not constant, but since we use the same F history as other OMs, this corresponds to \Fmsy at the mean selectivity parameters.

### Fleets {-}

We assumed a single fleet operating year round for catch observations with logistic selectivity for the fleet ($a_{50} = 5$ and slope = 1; Figure \ref{om_inputs_fig}, bottom left). This selectivity was used to define \Fmsy for the Beverton-Holt SRR parameters above. We assumed a logistic-normal distribution with no correlation on the multivariate normal scale for the age-composition observations for the fleet.

### Indices {-}

Two time series of fishery-independent surveys in numbers are generated for the entire 40 year period with one occurring in the spring (0.25 of each year) and one in the fall (0.75 of each year).  Catchability of both surveys are assumed to be 0.1.  Like the fishing fleet, we assumed logistic selectivity for both indices ($a_{50} = 5$ and slope = 1) and a logistic-normal distribution with no correlation on the multivariate normal scale for the age-composition observations.

### Observation Uncertainty {-}

The standard deviation for log-aggregate catch was 0.1. Two levels of observation error variance (high and low) were specified for indices and all age composition observations (both indices and catch). The low uncertainty specification assumed a standard deviation of 0.1 for both series of log-aggregate index observations, and the standard deviation of the logistic-normal for age composition observations was 0.3. In the high uncertainty specification, the standard deviation for log-aggregate indices was 0.4 and that for the age composition observations was 1.5. For all EMs, the standard deviation for log-aggregate observations was assumed known whereas that for the logistic-normal age composition observations was estimated.

### Operating models with random effects on numbers at age {-}

For operating models with random effects on recruitment and(or) apparent survival (R, R+S), we assumed marginal standard deviations for recruitment of $\sigma_R \in \{0.5,1.5\}$ and marginal standard deviations for older age classes of $\sigma_{2+} \in \{0,0.25, 0.5\}$. The full factorial combination of these process error assumptions (2x3 levels) and scenarios for fishing history (2 levels) and observation error (2 levels) scenarios described above results in 24 different R ($\sigma_{2+} = 0$) and R+S operating models (Table \ref{naa_om_table}).

### Operating models with random effects on natural mortality {-}
All R+M OMs treat natural mortality as constant across age, but with annually varying random effects. WHAM treats natural mortality as a log-transformed parameter 
$$
\log M_{y,a} = \mu_{M} + \epsilon_{M,y}
$$
that is a linear combination of a mean log-natural mortality parameter that is constant across ages ($\mu_{M} = \log(0.2)$) and any annual random effects are marginally distributed as $\epsilon_{M,y} \sim \text{N}\left(0,\sigma_M^2\right)$. The marginal standard deviations we assumed for log natural mortality random effects were $\sigma_M \in \{0.1, 0.5\}$ and the random effects were either uncorrelated or first-order autoregressive (AR1, $\rho_M \in \{0,0.9\}$). Uncorrelated random effects were also included on recruitment with $\sigma_R = 0.5$ (hence, we denote these OMs as  R+M). The full factorial combination of these process error assumptions and fishing history (2 levels) and observation error (2 levels) scenarios described above results in 16 different R+M OMs (Table \ref{M_om_table}).

### Operating models with random effects on fleet selectivity {-}

WHAM treats selectivity parameter $s$ as a logit-transformed parameter 
$$
\log\left(\frac{p_{s,y}-l_{s}}{u_{s}-p_{s,y}}\right) = \mu_s + \epsilon_{s,y}
$$
that is a linear combination of a mean $\mu_s$ and any annual random effects marginally distributed as $\epsilon_{s,y} \sim \text{N}\left(0,\sigma_s^2\right)$, where the lower and upper bounds of the parameter ($l_s$ and $u_s$) can be specified by the user. All selectivity parameters ($a_{50}$ and slope parameters) were bounded by 0 and 10 for all OMs and EMs. The marginal standard deviations we assumed for logit scale random effects were $\sigma_s \in \{0.1, 0.5\}$ and AR1 autocorrelation parameters of $\rho_s \in \{0,0.9\}$.  Like R+M OMs, the full factorial combination of these process error assumptions (2x2 levels) and scenarios described above for fishing history (2 levels) and observation error (2 levels) results in 16 different R+Sel OMs (Table \ref{sel_om_table}).

### Operating models with random effects on index catchability {-}

Like selectivity parameters, WHAM treats catchability for an index $i$ as a logit-transformed parameter 
$$
\log\left(\frac{q_{i,y}-l_{i}}{u_{i}-q_{i,y}}\right) = \mu_i + \epsilon_{i,y}
$$
that is a linear combination of a mean $\mu_i$ and any annual random effects marginally distributed as $\epsilon_{i,y} \sim \text{N}\left(0,\sigma_i^2\right)$ where the lower and upper bounds of the catchability ($l_i$ and $u_i$) can be specified by the user. We assumed bounds of 0 and 1000 for all OMs and EMs. For all OMs and EMs with process errors on catchability, the temporal variation only applies to the first index, which could be interpreted as capturing some unmeasured seasonal process that affects availability to the survey. The marginal standard deviations we assumed for logit scale random effects were $\sigma_i \in \{0.1, 0.5\}$ and AR1 autocorrelation parameters of $\rho_i \in \{0,0.9\}$.  Like R+M and R+Sel OMs, the full factorial combination of these process error assumptions and fishing history (2 levels) and observation error (2 levels) scenarios described above results in 16 different R+q OMs (Table \ref{q_om_table}).

## Estimation models {-}

For each of the data sets simulated from an OM, 20 EMs were fit. A total of 32 different EMs were fit across OMs where the subset of 20 depended on the source of process error in the OM (Table \ref{em_table}). The EMs have different assumptions about the source of process error (R+S, R+M, R+Sel, R+q) and whether or not 1) there is temporal autocorrelation, 2) a Beverton-Holt SRR is estimated, and 3) the natural mortality rate ($\mu_M$, the constant or mean on log scale for R+M EMs) is estimated. For simplicity we refer to the derived estimate $e^{\mu_M}$ as the median natural mortality rate regardless of whether natural mortality random effects are estimated in the EM.

Subsets of 20 EMs in Table \ref{em_table} were fit to simulate data sets from each of the OM process error categories. For R and R+S OMs, fitted EMs had matching process error assumptions as well as R+Sel, R+M, and R+q assumptions without autocorrelation. Similarly, For other OM process error categories, we fit EMs with matching process error assumptions as well as other process error types without autocorrelation. The maturity at age, weight at age for catch and spawning stock biomass (SSB), and observation error variance of aggregate catch and indices were all assumed known at the true values. However, the variance parameters for the logistic-normal distributions for age composition observations were estimated in the EMs. As such, EMs would either be configured completely correctly for the OM, or there could be mis-specification in assumptions of process error autocorrelation, the type of process error, or the SRR (Beverton-Holt or none).

## Measures of reliability {-}

### Convergence {-}

The first measure of reliability we investigated was frequency of convergence when fitting each EM to the simulated data sets. There are various ways to assess convergence of the fit [e.g., @carvalhoetal21; @kapuretal24], but given the importance of estimates of uncertainty when using assessment models in management, we estimated probability of convergence as measured by occurrence of a positive-definite hessian matrix at the optimized negative log-likelihood that could be inverted (i.e., providing hessian-based standard error estimates). We also provide results in the Supplementary Materials for convergence defined by the maximum absolute gradient < $1^{-6}$ and the maximum of the absolute gradient values for all fits of a given EM to all simulated data sets from a given OM that produced hessian-based standard errors for all estimated fixed effects. This provides an indication of how poor the calculated gradients can be, but still presumably converged adequately enough for parameter inferences.

### AIC for model selection {-}

We investigated the reliability of AIC-based model selection for two purposes. First, we analyzed selection of each process error model structure (R, R+S, R+M, R+Sel, R+q) using marginal AIC. For a given OM simulated data set, we compared AIC for EMs with different process error assumption conditional on whether median natural mortality rate and the Beverton-Holt SRR were estimated. We tabulated the models providing the lowest AIC across simulated data sets. Second, we analyzed AIC-based selection between EMs with and without the Beverton-Holt SRR assumed. Contrast in fishing pressure and time series with recruitment at low stock size has been shown to improve estimation of SRR parameters [@magnussonhilborn07; @connetal10]. Our preliminary inspections indicated generally poor performance of AIC in determining the Beverton-Holt model for a given set of OM factors (including contrast in fishing pressure). Therefore, we also considered the effect of of the log-standard deviation of the true log(SSB) (similar to the log of the coefficient of variation for SSB) on model selection since simulations with realized SSB producing low and high recruitment would have larger variation in realized SSB.

All model selection results condition only on completion of the optimization process without failure for all of the compared EMs. We did not condition on convergence as defined above because optimization could correctly determine an inappropriate process error assumption by estimating variance parameters at the lower bound of zero. Such an optimization could indicate poor convergence but the likelihood would be equivalent to that without the mis-specified random effects and the AIC would be appropriately higher because more (variance) parameters were estimated. All other measures of reliability described below (bias and Mohn's $\rho$) use these same criteria for inclusion of EM fits in the summarized results.

### Bias {-}

We also investigated bias in estimation of various model attributes as a measure of reliability. For a given model attribute we calculated the relative error 
\begin{equation}\label{relerror}
\text{RE}\left(\theta_i\right) = \frac{\widehat \theta_i - \theta_i}{\theta_i}
\end{equation}
from fitting a given estimating model to simulated data set $i$ configured for a given OM where $\widehat \theta_i$ and $\theta_i$ are the estimated and true values for simulation $i$. We analyzed simulation results for estimates of terminal year estimates of SSB and recruitment, Beverton-Holt SRR parameters ($a$ and $b$), and median natural mortality rate.

### Mohn's $\rho$ {-}

Finally, we investigated presence of retrospective patterns in fitted models as a measure of reliability. We calculated Mohn's $\rho$ for SSB, fishing mortality (averaged over all age classes), and recruitment for each EM fit to each OM simulated data set [@mohn99]. We fit $P = 7$ peels to each simulated data set and calculated Mohn's $\rho$ for a given attribute $\theta$ as
\begin{equation}\label{mohns_rho}
\rho\left(\theta\right) = \frac{1}{P}\sum^P_{p=1} \frac{\widehat \theta_{Y-p,Y-p} - \widehat\theta_{Y-p,Y}}{\widehat\theta_{Y-p,Y}}
\end{equation}
where $\widehat \theta_{i,j}$ is the estimate for attribute $\theta$ in year $i$ from a model fit using data up to year $j$.

### Summarizing results across OM and EM attributes {-}

The measures of central tendency and variability of observed values for specific OM and EM attributes (e.g., low or high observation error) that we described above can indicate scenarios that provide better or poorer reliability. However, the OM and EM attributes that we investigated are numerous, so we used two methods to summarize the most important factors for differences in results.  The first method was fitting regression models with the response being each of the measures of reliability described above and predictor variables were defined based on OM and EM characteristics [e.g., @mackinnonetal95; @wangetal17; @harwelletal18]. For the binary indicators of convergence and AIC-based selection of a SRR, we performed logistic regressions. For indicators of AIC-based selection of EM process error type (multiple categories) we performed multinomial regressions. For other measures of reliability we fit linear regression models to transformed responses. Because relative errors (Eq. \ref{relerror}) and Mohn's $\rho$ for the various parameters are bounded below at -1, we used a transformation of these values
\begin{equation}\label{bias_regression_response}
y_i = \log\left[f\left(\widehat \theta_i,\theta_i\right)+1\right]
\end{equation}
where $f$ is either the relative error (Eq. \ref{relerror}) or Mohn's $\rho$ (Eq. \ref{mohns_rho}) for simulation $i$, so that values are unbounded. For relative errors, $y_i$ is the log-scale error. We omitted simulation estimates equal to zero (RE = -1). For all regressions we fit separate models with individual factors included, with all factors combined, with including all second order interactions, and including all third order interactions. For the multinomial regression, we used the \verb|vglm| function from the VGAM package [@yee08;@yee15]. We tabulated percent reduction in residual deviance for each of regression fits. We did not perform formal statistical analyses of effects of OM and EM attributes on results (e.g., ANOVA) because of the lack of independence of the "observations" that results from fitting multiple EMs to each simulated data set.

The second method involved fitting classification and regression trees [@breimanetal84] to show how the OM and EM attributes, and their interactions, split the values for each measure of reliability [e.g., @gonzalezetal18;@collieretal22]. We used classification trees for categorical measures (convergence and AIC) and regression trees for the other measures with continuous scales (relative error and Mohn's $\rho$). The response variables were the same as the regressions for the deviance reduction analyses. We used the \verb|rpart| function in the rpart package (https://cran.r-project.org/package=rpart) to fit trees. Full trees were determined using default settings except that we increased the number of cross-validations to 100. For clarity, we pruned the full trees to show just the primary branches.

We also provide detailed results for all measures of reliability at each combination of OM and EM attributes in the Supplementary Materials. For confidence intervals of probability of convergence, we used the Clopper-Pearson exact method [@clopperpearson34;@thulin14]. For AIC selection of process error configuration we provide estimates of the proportions of simulations where each EM type was selected. For AIC selection of the SRR, we provide predicted probabilities from logistic regressions as a function of SSB variability for each OM and EM type. We estimated bias as the median of the relative errors across all simulations for a given OM and EM combination. We constructed 95\% confidence intervals for the median relative bias, and Mohn's $\rho$ using the binomial distribution approach as in @millerhyun18 and @stockmiller21.  For each EM we calculated median and 95\% confidence intervals using the same methods as that for relative bias.

# Results {-}

## Convergence performance {-}

For probability of convergence, EM process error assumption was the single attribute that resulted in the largest percent reduction in deviance (14-28\%) for all OM process error types other than R+S OMs where the EM M assumption explained the most residual deviance (>11\%; Table \ref{convergence_PRD_table}). However, including interactions of OM and EM factors also provided large reductions in residual deviance (35-47\%), suggesting successful convergence depended on a combination OM and EM attributes.

Classification trees for each OM process error type, all had the primary branch defined using the same attribute that provided the largest reduction in deviance (Figure \ref{conv_class}). EMs that assumed R+S process errors converged poorly for all OMs that were simulated with the alternative process error assumptions (R, R+M, R+Sel, an R+q OMs). For all trees, branches based on the OM fishing mortality history showed better convergence when the OM included a change in fishing pressure. Branches based on whether the Beverton-Holt SRR was assumed or not, showed better convergence when it was not estimated and branches based on whether the EM estimated the median natural mortality rate showed better convergence when it was treated as known. For certain R+M and R+Sel OMs, better convergence was also observed when there was lower observation uncertainty.

When convergence is defined by a gradient threshold, the primary factor explaining deviance reduction is the same for all OM process error types, but there are some differences in deviance reduction for secondary factors (Table \ref{convergence_gradient_PRD_table}), and probability of convergence, overall, was lower (Figure \ref{conv_gradient_class}). We found a wide range of maximum absolute values of gradients for models that had invertible Hessians (Figure \ref{hess_grad}). The largest value observed for a given EM and OM combination was typically $<10^{-3}$, but many converged models had values greater than 1. For many OMs, EMs that assumed the correct process error type and did not estimate median natural mortality or the Beverton-Holt SRR produced the lowest gradient values.

## AIC performance {-}

### Process error structure {-}

For AIC selection of the correct process error configuration, the magnitude of observation and process error were the attributes that resulted in the largest percent reductions in deviance across OM process error types other than R OMs (Table \ref{AIC_PE_PRD_table}). Both variance of apparent survival random effects ($\sigma_{2+}$) and degree of observation error explained the largest reductions for R+S (17-22\%) and R+Sel (8-26\%) OMs, whereas variance of process errors provided the largest reductions in R+M (>9\%) and R+q (>13\%) OMs. Comparatively, none of the OM or EM attributes explained particularly large reductions in deviance for R OMs, but fishing history, whether a SRR was estimated, and whether median natural mortality was known or estimated provided similar and the largest reductions (approximately 5-6\%). Inclusion of second and third order interactions, did not provide large reductions in deviance for any of the OM process error types.

For all OM process error types other than R OMs, the attributes defining the primary branches of classification trees matched those that provided the largest reductions in deviance (Figure \ref{AIC_PE_class}). Across all OMs, AIC was more accurate for the process error type when process error variability was greater and when observation error was lower. No branches were estimated for classification trees fit to the R OMs, likely because accuracy was high across all simulations (0.94), although inspection of the fine-scale results shows there is some degradation in AIC selection when a SRR and median natural mortality rate are estimated for R OMs with constant fishing pressure and high observation error (Figure \ref{pe_aic}, top left).

### Stock-recruit relationship {-}

Logistic regressions for AIC selection of the Beverton-Holt SRR, showed OM fishing history and variation in SSB ($\log \text{SD}_{\text{SSB}}$) provided substantial reductions in deviance for R+M (>13\%), R+Sel (>26\%), and R+q (>24\%) OMs (Table \ref{AIC_SRR_PRD_table}). For R OMs, fishing history provided the largest reduction in deviance (>9\%) whereas none of the attributes individually provided large reductions in deviance for R+S OMs (all <5\%). However, inclusion of all attributes provided larger reductions in deviance than the sum of individual contributions for both R (>30\%) and R+S (>18\%) OMs. For all OM process error types, inclusion of interaction terms provided relatively little reduction in residual deviance.

Attributes defining the primary branches of classification trees for AIC selection of the SRR assumption were the same as those explaining the largest reductions in deviance for the logistic regression models (Figure \ref{AIC_SRR_class}). All branches based on variation in SSB showed better accuracy with larger variability in SSB and all branches based on fishing history showed better accuracy when there was contrast in fishing pressure. Branched based on OM observation error or recruitment variability showed better accuracy when they were lower. For R OMs, a combination of lower recruitment variability, contrast in fishing pressure, and higher SSB variability produced AIC accuracy over 0.8. For R+S OMs, lower recruitment variability and observation error and higher SSB variability produced AIC accuracy of 0.79. For R+M, R+Sel, and R+q OM,s accuracy of 0.87 to 0.94 was observed just with higher SSB variability.

## Bias {-}

### Terminal year spawning stock biomass, fishing mortality, and recruitment {-}

Regression models for log-scale errors in SSB that included the various OM and EM factors showed little reduction in deviance (<5\%) for any of the factors across all OM process error types (Table \ref{bias_SSB_PRD_table}). The attributes producing the largest reductions were the EM assumption for median natural mortality (known or estimated) for R, R+M, R+Sel, and R+q OMs (1-3\%), EM process error type for R+S OMs (4\%) and fishing history for all OM types (1-5\%). Including second interactions provided largest reductions in residual deviance (10- 26\%). Including third order interactions also provided large further reductions for R, R+S, and R+q OMs (between 5 and 11\%). In all regression trees, branches based on fishing history and level of observation error generally showed lower bias in SSB with contrast in fishing and lower observation error (Figure \ref{SSB_bias_regtree}). For branches based on treatment of median natural mortality rate, bias was generally lower when it was known rather than estimated. For some R+Sel and R+q OMs, less bias in SSB was shown when the EM process error configuration was correct.

Results for bias in fishing mortality and recruitment generally matched those for SSB, except that directions of bias for fishing mortality were opposite to those for SSB and recruitment. Effects of individual OM and EM factors on regression models were similarly small as measured by reduction in deviance (Tables \ref{bias_F_PRD_table} and \ref{bias_R_PRD_table}). Primary branches of regression trees were in most cases identical to those for SSB  (Figures \ref{F_bias_regtree} and \ref{R_bias_regtree}).

### Stock-recruit parameters {-}

Regression models for transformed relative errors of estimates of both the Beverton-Holt $a$ and $b$ parameters showed none of the factors explained large percent reductions in deviance (Table \ref{bias_SR_pars_PRD_table}). The OM fishing history provided the largest reduction for most OM process error types for both parameters, but percent reductions were less than 5.6\% except for R+Sel OMs where the percent reductions were 11.37\% and 7.97\% for the $a$ and $b$ parameters, respectively and for just the $b$ parameter for R+q OMs (10\%). The EM process error assumption provided similar reductions in deviance for both parameters for R OMs. Including interactions also did not produce important reductions in deviance (increase of .

For regression trees of log-scale error in Beverton-Holt $a$ and $b$ parameters, lower bias was indicated with contrast in OM fishing pressure for all branches in trees for each OM process error type (Figures \ref{SR_a_bias_regtree} and \ref{SR_b_bias_regtree}). For all branches based on recruitment variability in trees for R and R+S OMs, lower bias in both $a$ and $b$ was observed with less recruitment variability. For R OMs with contrast in fishing pressure and greater recruitment variability EMs that assumed the incorrect R+M process errors produced lower bias in both $a$ and $b$ than other process error assumptions.  Across all OMs, there was generally less bias and(or) lower variability in estimation of the $a$ parameter than the $b$ parameter (Figure \ref{SR_rel_error}).

### Median natural mortality rate {-}

Fitted regression models for log-scale errors in median natural mortality rate showed largest percent reductions in residual deviance for R+S and R_M models (Table \ref{bias_median_M_PRD_table}). The largest reductions for a single attribute was the EM process error assumption (>20\%) and fishing history (>15\%) for R+S OMs. Fishing history also provided >10\% reduction for R+M OMs, but reductions for all factors in R, R+Sel, and R+q OMs were relatively low (<6\%). Interactions of OM and EM factors also provided substantial further reductions for R+S and R+M OMs (between 8 and 15 \% for second order interactions).

Regression trees with branches based on fishing history showed lower bias in median natural mortality rate with contrast in fishing pressure and branches based on level of observation error showed lower bias with more precise observations (Figure \ref{med_M_bias_regtree}). For R OMs, branches based on EM process error assumption showed lower bias with EMs assuming the correct R and the incorrect R+S assumption. For R+S and R+M OMs, branches based on EM process error showed only the correct EM process error assumption with less bias.

## Mohn's $\rho$ {-}

Regression models for Mohn's $\rho$ of SSB showed little reduction in deviance for any of the OM an EM attributes (<2\%; Table \ref{mohns_rho_SSB_PRD_table}). The lack of explanatory power is also reflected in the regression trees where median Mohn's $\rho$ values are near zero unless a large combinations of OM and EM conditions occur (Figure \ref{SSB_mohns_rho_regtree}). For example, In R+S OMs, with constant fishing pressure, high observation error, and higher apparent survival process error, EMs that assume R+M process errors have a median Mohn's $\rho = -0.068$.

Similarly poor explanatory power of the OM and EM attributes occured when we fit regression models for Mohn's $\rho$ of fishing mortality and recruitment (Tables \ref{mohns_rho_F_PRD_table} and \ref{mohns_rho_R_PRD_table}). Regression trees for Mohn's $\rho$ of fishing mortality were similar to those for SSB in that median values of Mohn's $\rho$ were close to zero for most combinations of OM and EM attributes (Figure \ref{F_mohns_rho_regtree}). However, we observed median Mohn's $\rho$ for recruitment greater than 0.1 at  branches much closer to the base of the trees with fewer interactions of the OM and EM attributes (Figure \ref{R_mohns_rho_regtree}). These branches with consistently large retrospective patterns were typically defined by larger OM observation error, OM constant fishing pressure, or incorrect EM process error configuration. Comparing regression model and regression tree fits, attributes defining the primary branches for all regression trees of all Mohn's $\rho$ values (SSB, fishing mortality, and recruitment) generally matched those that explained the largest reductions in deviance.

# Discussion {-}

## Convergence {-}

Analyses of model convergence across simulations can be useful for understanding the utility of alternative convergence criteria used in applications to real data for directing the practitioner to more appropriate random effects configurations. It is common during the assessment model fitting process to check that the maximum absolute gradient component is less than some threshold prior to inspecting the Hessian of the optimized likelihood for invertibility [@carvalhoetal21]. However, there is no accepted standard for the gradient threshold [e.g., @leeetal11; @hurtadoferroetal15;@ruddthorson18] and some thresholds would exclude models that in fact have an invertible Hessian. We found the Hessian at the optimized log-likelihood can often be invertible when the maximum absolute gradient was much larger than what might be perceived to be a sensible threshold.

@lietal24 found that convergence rate could be a useful diagnostic especially for separating the correct model from overly complex models. However, the criteria for convergence used in their study may also lead to limited ability to distinguish the correct model from overly simplistic models, a pattern that was also noted by @liljestrandetal24 in which one process error may absorb all sources of process error when the magnitude of other process errors are low.

Often poor convergence result when parameter estimates are at their bounds [@carvalhoetal21], and this also applies to variance parameters for random effects with state-space assessment models. Even when the Hessian is invertible, parameters that are poorly informed will have extremely large variance estimates. This further inspection can lead to a more appropriate and often more parsimonious model configuration where the problematic parameters are not estimated. For example, process error variance parameters that are estimated close to 0 indicates that the random effects are estimated to have little or no variability and removing these process errors is warranted. Generally, our results suggest we can expect lower probability of convergence of state-space assessment models when estimating natural mortality or SRRs because of the difficulty distinguishing these parameters from others being estimated in assessment model with data that are typically available. Our experiments did not aim to emulate the practitioner decision process in developing model configurations (e.g. removing a source of process error and refitting the model when process error variance parameters were estimated close to 0). Evaluating the efficacy of such a decision process when applying EMs might be important in closed loop simulations (e.g. MSE) aimed at quantifying management performance.

A factor affecting the convergence criteria, particularly for maximum likelihood estimation of models with random effects, is numerical accuracy. All optimizations performed in these simulations are of the Laplace approximation of the marginal likelihood and, therefore, gradients and Hessians are also with respect to this approximation (see TMB::sdreport in the Template Model Builder package). Functionality within the Template Model Builder package exists (i.e., TMB::checkConsistency) to check the validity of the Laplace approximation and the utility of this as a diagnostic for state-space assessment models should be explored further. Furthermore, numerical methods are used to calculate and invert the Hessian for variance estimation for models with random effects. Our results, along with the potential lack of accuracy imposed by these approximations, suggest at least investigating whether the Hessian is positive definite when the calculated absolute gradients are not terribly large (e.g, $< 1$).

## AIC {-}

Of the OM process error configurations we considered, we found AIC to be accurate for selecting models with process errors on recruitment and apparent survival (R and R+S). Fitting models to other OMs rarely preferred R+S EMs, and R and R+S EMs were nearly always selected for the matching OMs; a similar result was reported by @liljestrandetal24. For other sources of process error, accuracy of AIC was improved when there was larger variability in the process errors and/or lower observation error.

Across all OM process error configurations, AIC performed poorly in identifying that the presence of the Beverton-Holt SRR in the OM unless there was contrast in fishing pressure possibly in combination with other factors such as lower variability in recruitment process errors (in R and R+S models) or greater variation in natural mortality process errors (for R+M OMs, Fig. \ref{sr_aic}). As such, properly accounting for process error in natural mortality could be important [@lietal24] when evaluating SRRs in state-space models. Curiously, we did not find a marked effect of the level of observation error on ability to detect the SRR, but it is possible that AIC would perform better if observations have even lower uncertainty than we considered.

Although we did not compare models with alternative SRRs (e.g., Ricker and Beverton-Holt), we do not expect AIC to perform any better distinguishing between relationships. Our finding that AIC tended to choose simpler recruitment models in most cases contrasts with the noted bias in AIC for more complex models [@shibata76;@katz81;@kassraftery95], but, whereas those findings apply to the much more common comparison of models that are fit to raw and independent observations, here we are comparing state-space models which account for observation error and estimate process errors in latent variables.

Our results comport with those of @devalpinehastings02 who found AIC could not distinguish among state-space SRRs that were fit just to SSB and recruitment observations (i.e., not an assessment model). Similarly,  @brittenetalinreview found AIC could not reliably distinguish alternative environmental effects on SRR parameters. However, @milleretal16 did find AIC to prefer a SRR with environmental effects when applied to data for the SNEMA yellowtail flounder stock and  AIC also selected an environmental covariate on a SRR for the most recent stock assessment of Georges Bank yellowtail flounder [@nefsc25]. Both of these yellowtail flounder stocks have large changes in stock size and the values of environmental covariates over time. Additionally, this species is well-observed by the bottom trawl survey that is used for an index in assessment models.

## Bias {-}

As expected, bias in all parameters and assessment output was generally improved with lower observation error. Estimation of SRR parameters was reliable in ideal scenarios of low observation error and contrast in fishing for some R+Sel and R+M OMs, but generally estimation was biased and(or) highly variable. We found substantial bias in estimated SRR parameters in R and R+S OMs particularly with high variability in recruitment and apparent survival process errors, suggesting that practitioners should be cautious when fitted assessment models have these properties.

On the other hand, estimation of median natural mortality was reliable in many OM scenarios with contrast in fishing pressure, consistent with @hoenigetal25. In some OMs, when EMs estimated the SRR parameters and median natural mortality, bias for those parameters was improved.  Conversely, for some R+Sel and R+q OMs where there was bias in natural mortality due to high observation error, estimating the SRR reduced the bias in median natural mortality rate.  However, estimating median natural mortality did cause poor accuracy in SSB estimation in many OMs without contrast in fishing pressure over time and with higher observation error. Thus, estimating median natural mortality should be approached with caution in state-space assessment models, particularly given its significant impact on determination of reference point and stock status [@lietal24].

## Retrospective patterns {-}

Incorrect EM process error assumptions did not produce strong retrospective patterns for SSB for any OMs regardless of whether median natural mortality or a SRR was estimated, but some weak retrospective patterns occur when observation error was high and there was contrast in fishing pressure. However, retrospective patterns tended to be more variable for recruitment and were sometimes large even when the EM was correct. Therefore, we recommend emphasis on inspection of retrospective patterns primarily for SSB and $F$, but further research on retrospective patterns in other assessment model parameters, management quantities such as biological reference points, and projections may be beneficial [@brookslegault16].

The general lack of retrospective patterns with mis-specified process errors is perhaps to be expected. Retrospective patterns are often induced in simulation studies by rapid changes in a quantity such as index catchability, natural mortality, or perceived catch during years toward the end of the time series [@legault09; @millerlegault17; @huynhetal22; @breiviketal23]. In our simulations, the process errors changing over time may have trends in particular simulations, particularly when strong autocorrelation is imposed, but the random effects have no trend on average across simulations. @szuwalskietal18 and @lietal24 also found relatively small retrospective patterns when the source of mis-specification was temporal variation in demography attributes. Indeed, it is common for the flexibility provided by temporal random effects to reduce retrospective patterns [@milleretal18; @stockmiller21; @stocketal21], though it does not necessarily indicate a more accurate assessment model [@perrettietal20; @lietal24; @liljestrandetal24]. Our results together with the existing literature seem to suggest that when a strong retrospective pattern is observed in an assessment it is more likely to be due to a mis-specification of a rapid shift in some model attribute rather than whether a particular process is assumed to be randomly varying temporally.

## Analytical approach {-}

We found the use of regression models and classification and regression trees extremely useful in understanding the most important OM and EM attributes explaining variation in the measures of reliability we examined across all simulations. The classification and regression trees are generally a good tool for determining the values of the OM and EM attributes that produce better or worse measures of reliability. However, determining the combination of attributes that produce the best or worst measures of reliability can be challenging using the trees. For example, in the regression tree for median natural mortality rate estimates in R OMs (Figure \ref{median_M_bias_regtree}),  the both of the first branches imply bias is low regardless of OM fishing history, but when OM fishing pressure is constant, results are much better when OM observation error is low (median RE about -6\%) than when OM observation error is high (median RE about 40\%). Therefore, inspection of results by all combinations of OM and EM factors (as provided in the supplementary materials) can be important.

## Conclusions {-}

Our simulation study examined the importance of several factors for reliable inferences from state-space age-structured assessment models. Contrast in fishing pressure and lower observation error consistently improved all measures of reliability we examined. AIC accurately distinguished models with process errors on recruitment only (R) or on recruitment and apparent survival (R+S). Therefore, we expect practitioners will find R+S configurations to provide satisfactory diagnostics across a range of life history and data quality scenarios.  Accurate AIC selection of the Beverton-Holt SRR generally required a combination of low recruitment variability, contrast in fishing pressure and large variation in SSB over time. However, when the Beverton-Holt was correctly assumed, some bias in estimation in at least one of the parameters existed regardless of any of the other OM and EM configurations. Because bias in terminal SSB and retrospective patterns were indifferent to whether or not the SRR was estimated, the prevalence of bias in SRR parameter estimation, and often better convergence without the SRR, a sensible default would be to not assume a SRR when fitting assessment models.

# Acknowledgements {-}

This work was funded by NOAA Fisheries Northeast Fisheries Science Center. We thank Jon Deroba, 2 anonymous reviewers, and the associate editor for helpful comments on earlier versions of this manuscript that markedly improved its clarity.

\pagebreak

# References {-}

<div id="refs"></div>

\pagebreak

\clearpage

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{convergence_classification_plots}
\end{center}
\caption{Classification trees indicating primary factors determining convergence as defined by providing hessian-based standard errors for R, R+S, R+M, R+Sel and R+q OMs. Lower or higher convergence rates are indicated by more red or green polygons, respectively}\label{conv_class}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{AIC_PE_classification_plots}
\end{center}
\caption{Classification trees indicating primary factors determining which EM process error assumption provides the lowest AIC for R+S, R+M, R+Sel and R+q OMs. Each node shows the proportion of EM process error models with lowest AIC (top) and number of observations (bottom) for the corresponding subset. Lower or higher accuracy of the process error assumption are indicated by more red or green polygons, respectively.}\label{AIC_PE_class}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{AIC_SRR_classification_plots}
\end{center}
\caption{Classification trees indicating primary factors determining which EM SRR assumption (none or Beverton-Holt) provides the lowest AIC for R, R+S, R+M, R+Sel and R+q OMs. Each node shows the proportion of EMs that assume the SRR with lowest AIC (top) and number of observations (bottom) for the corresponding subset. Lower or higher accuracy of the process error assumption are indicated by more red or green polygons, respectively.}\label{AIC_SRR_class}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{SSB_bias_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in estimation measured by Eq. \ref{bias_regression_response} for terminal year SSB for R+S, R+M, R+Sel and R+q OMs. Each node shows the median error (top) and number of observations (bottom) for the corresponding subset. Median errors closer to or further from zero are indicated by more green or red polygons, respectively.}\label{SSB_bias_regtree}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{SR_a_bias_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in estimation measured by Eq. \ref{bias_regression_response} for the Beverton-Holt SRR parameter $a$ for R+S, R+M, R+Sel and R+q OMs. Each node shows the median error (top) and number of observations (bottom) for the corresponding subset. Lower or higher median absolute errors of the process error assumption are indicated by more green or red polygons, respectively.}\label{SR_a_bias_regtree}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{SR_b_bias_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in estimation measured by Eq. \ref{bias_regression_response} for the Beverton-Holt SRR parameter $b$ for R+S, R+M, R+Sel and R+q OMs. Each node shows the median error (top) and number of observations (bottom) for the corresponding subset. Lower or higher median absolute errors of the process error assumption are indicated by more green or red polygons, respectively.}\label{SR_b_bias_regtree}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{med_M_bias_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in estimation measured by Eq. \ref{bias_regression_response} for the median natural mortality rate for R+S, R+M, R+Sel and R+q OMs. Each node shows the median error (top) and number of observations (bottom) for the corresponding subset. Lower or higher median absolute errors of the process error assumption are indicated by more green or red polygons, respectively.}\label{med_M_bias_regtree}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{SSB_mohns_rho_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in transformed Mohn's $\rho$ (Eq. \ref{bias_regression_response}) for SSB for R+S, R+M, R+Sel and R+q OMs. Each node shows the median Mohn's $\rho$ (top) and number of observations (bottom) for the corresponding subset. Median Mohn's $\rho$ closer to or further from zero are indicated by more green or red polygons, respectively.}\label{SSB_mohns_rho_regtree}
\end{figure}
\end{landscape}

\pagebreak

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for logistic regression models fit to indicators of convergence (providing hessian-based standard errors) with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{convergence_PRD_table}
{\input{convergence_PRD_table}}
\end{table}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for multinomial logistic regression models fit to indicators of EM process error assumption with lowest AIC with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{AIC_PE_PRD_table}
{\input{AIC_PE_PRD_table}}
\end{table}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for logistic regression models fit to indicators of EM SRR assumption (none or Beverton-Holt) with lowest AIC with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{AIC_SRR_PRD_table}
{\input{AIC_SRR_PRD_table}}
\end{table}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for linear regression models fit to errors in estimation measured by Eq. \ref{bias_regression_response} for the terminal year SSB with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{bias_SSB_PRD_table}
{\input{bias_SSB_PRD_table}}
\end{table}

\begin{landscape}
\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for linear regression models fit to errors in estimation measured by Eq. \ref{bias_regression_response} for the Beverton-Holt SRR parameters with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{bias_SR_pars_PRD_table}
{\input{bias_SR_pars_PRD_table}}
\end{table}
\end{landscape}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for linear regression models fit to errors in estimation measured by Eq. \ref{bias_regression_response} for the median natural mortality rate parameter with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{bias_median_M_PRD_table}
{\input{bias_median_M_PRD_table}}
\end{table}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for linear regression models fit to transformed Mohn's $\rho$ values for each simulation (Eq. \ref{bias_regression_response}) for SSB with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{mohns_rho_SSB_PRD_table}
{\input{mohns_rho_SSB_PRD_table}}
\end{table}

\begin{landscape}
\end{landscape}
\pagebreak

\setcounter{figure}{0}
\renewcommand\thefigure{S\arabic{figure}}

\setcounter{table}{0}
\renewcommand\thetable{S\arabic{table}}

\pagebreak

# Supplementary Materials {-}

\pagebreak

## Referenced Figures {-}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width = \textwidth]{om_input_plots_figure}
\end{center}
\caption{The proportion mature at age, weight at age, fleet and index selectivity at age, and Beverton-Holt SRR assumed for the population in all operating models. For operating models with random effects on fleet selectivity, this represents the selectivity at the mean of the random effects.}\label{om_inputs_fig}
\end{figure}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{convergence_gradient_classification_plots}
\end{center}
\caption{Classification trees indicating primary factors determining convergence as defined by a maximum absolute gradient < $10^{-6}$ for R, R+S, R+M, R+Sel and R+q OMs. Lower or higher convergence rates are indicated by more red or green polygons, respectively}\label{conv_gradient_class}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{hess_grad_convergence_plots}
\end{center}
\caption{The maximum of the absolute values of all gradient values for all fits that provided hessian-based standard errors across all simuated data sets of a given OM configuration (A: R and R+S, B: R+M, C: R+Sel, or D: R+q).  Results are conditional on EM fits with alternative process error type (colored points and lines), median natural mortality (estimated or known) and recruitment assumptions (Beverton-Holt SRR or not). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{hess_grad}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{pe_aic_plots}
\end{center}
\caption{Estimated probability of lowest AIC for EMs assuming alternative process error structures (colored bars) conditional on alternative assumptions for median natural mortality (estimated or known) and Beverton-Holt SRR (estimated or not; along x-axis) when fitted to operating models that have R and R+S (A), R+Sel (B), R+M (C), or R+q (D) process error structures. Striped bars indicate results where the EM process error structure matches that of the operating model.}\label{pe_aic}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{F_bias_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in estimation measured by Eq. \ref{bias_regression_response} for terminal year fully-selected fishing mortality for R+S, R+M, R+Sel and R+q OMs. Each node shows the median error (top) and number of observations (bottom) for the corresponding subset. Median errors closer to or further from zero are indicated by more green or red polygons, respectively.}\label{F_bias_regtree}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{R_bias_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in estimation measured by Eq. \ref{bias_regression_response} for terminal year recruitment for R+S, R+M, R+Sel and R+q OMs. Each node shows the median error (top) and number of observations (bottom) for the corresponding subset. Median errors closer to or further from zero are indicated by more green or red polygons, respectively.}\label{R_bias_regtree}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{sr_bias_plots}
\end{center}
\caption{Median relative error of Beverton-Holt SRR parameters ($a$ and $b$) for estimating models fitted to data sets simulated with alternative process error structures: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{SR_rel_error}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{F_mohns_rho_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in transformed Mohn's $\rho$ (Eq. \ref{bias_regression_response}) for fishing mortality averaged over all age classes for R+S, R+M, R+Sel and R+q OMs. Each node shows the median Mohn's $\rho$ (top) and number of observations (bottom) for the corresponding subset. Median Mohn's $\rho$ closer to or further from zero are indicated by more green or red polygons, respectively.}\label{F_mohns_rho_regtree}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{R_mohns_rho_regtree_plots}
\end{center}
\caption{Regression trees indicating primary factors determining reductions in sums of squares of errors in transformed Mohn's $\rho$ (Eq. \ref{bias_regression_response}) for recruitment for R+S, R+M, R+Sel and R+q OMs. Each node shows the median Mohn's $\rho$ (top) and number of observations (bottom) for the corresponding subset. Median Mohn's $\rho$ closer to or further from zero are indicated by more green or red polygons, respectively.}\label{R_mohns_rho_regtree}
\end{figure}
\end{landscape}

## Referenced Tables {-}

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on recruitment and apparent survival (R.R+S). Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant.}\label{naa_om_table}
{\footnotesize \input{naa_om_tab}}
\end{table}
\end{landscape}

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on recruitment and natural mortality (R+M). Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}\label{M_om_table}
{\input{M_om_tab}}
\end{table}
\end{landscape}

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on recruitment and selectivity (R+Sel). Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}\label{sel_om_table}
{\input{Sel_om_tab}}
\end{table}
\end{landscape}

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on recruitment and catchability (R+q). Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}\label{q_om_table}
{\input{q_om_tab}}
\end{table}
\end{landscape}

\begin{table}
\caption{Distinguishing characteristics of the estimating models and operating model process error categories (R, R+S, R+M, R+Sel, R+q) where used.}\label{em_table}
{\scriptsize \input{em_tab}}
\end{table}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for logistic regression models fit to indicators of convergence (maximum absolute gradient < $10^{-6}$) with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{convergence_gradient_PRD_table}
{\input{convergence_gradient_PRD_table}}
\end{table}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for linear regression models fit to errors in estimation measured by Eq. \ref{bias_regression_response} for the terminal year fully-selected fishing mortality with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{bias_F_PRD_table}
{\input{bias_F_PRD_table}}
\end{table}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for linear regression models fit to errors in estimation measured by Eq. \ref{bias_regression_response} for the terminal year recruitment with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{bias_R_PRD_table}
{\input{bias_R_PRD_table}}
\end{table}

\clearpage

## Further Detailed Results {-}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{term_SSB_bias_plots}
\end{center}
\caption{Median relative error of terminal year SSB for estimating models fitted to data sets simulated with alternative process error structures: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{SSB_rel_error}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{term_F_bias_plots}
\end{center}
\caption{Median relative error of terminal year fully-selected fishing mortality for estimating models fitted to data sets simulated with alternative process error structures: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{F_rel_error}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{term_R_bias_plots}
\end{center}
\caption{Median relative error of terminal year recruitment for estimating models fitted to data sets simulated with alternative process error structures: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{R_rel_error}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{M_bias_plots}
\end{center}
\caption{Median relative error of median natural mortality for estimating models fitted to data sets simulated with alternative process error structures: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{M_rel_error}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{mohns_rho_ssb_plots}
\end{center}
\caption{Median Mohn's $\rho$ for SSB for estimating models fitted to data sets simulated with alternative process error structures: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{mohns_rho_ssb}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{mohns_rho_F_plots}
\end{center}
\caption{Median Mohn's $\rho$ of fishing mortality averaged over all age classes for estimating models fitted to data sets simulated with alternative process error structures: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{mohns_rho_F}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{mohns_rho_R_plots}
\end{center}
\caption{Median Mohn's $\rho$ of recruitment for estimating models fitted to data sets simulated with alternative process error structures: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{mohns_rho_R}
\end{figure}
\end{landscape}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for linear regression models fit to transformed Mohn's $\rho$ values for each simulation (Eq. \ref{bias_regression_response}) for fishing mortality averaged over all age classes with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{mohns_rho_F_PRD_table}
{\input{mohns_rho_F_PRD_table}}
\end{table}

\begin{table}
\caption{For each OM process error type (columns), percent reduction in deviance for linear regression models fit to transformed Mohn's $\rho$ values for each simulation (Eq. \ref{bias_regression_response}) for recruitment with each OM and EM factor (rows) included individually, combined, and with second and third order interactions.}\label{mohns_rho_R_PRD_table}
{\input{mohns_rho_R_PRD_table}}
\end{table}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{type_4_convergence_plots}
\end{center}
\caption{Probability of EMs providing hessian-based standard errors with alternative process error (colored points and lines), and median natural mortality (estimated or known) and Beverton-Holt SRR (estimated or not; along x-axis) assumptions when fitted to OMs that have R and R+S (A), R+Sel (B), R+M (C), or R+q (D) process error structures. Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{hessian_SE_convergence}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{type_3_convergence_plots}
\end{center}
\caption{Probability of EMs providing maximum absolute values of gradients less than $10^{-6}$ with alternative process error (colored points and lines), and median natural mortality (estimated or known) and Beverton-Holt SRR (estimated or not; along x-axis) assumptions when fitted to OMs that have R and R+S (A), R+Sel (B), R+M (C), or R+q (D) process error structures. Circled values indicate results where the EM process error structure matches that of the operating model and vertical lines represent 95\% confidence intervals.}\label{gradient_convergence}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[height = 0.8\textheight]{sr_aic_plots_rev}
\end{center}
\caption{Probability of lowest AIC from logistic regression on the log-standard deviation of the true log(SSB) in each simulation for estimating model with Beverton-Holt SRRs, rather than the otherwise equivalent EM without the SRR. Results are conditional on median M is known in the EM and alternative assumptions EMs having the correct process error structure: R and R+S (A), R+Sel (B), R+M (C), or R+q (D), and median M is assumed known in the EM. Solid and dashed lines are for OMs with and without temporal contrast in fishing pressure, respectively, and polygons represent 95\% confidence intervals. Range of results indicates the range of log-standard deviation of log(SSB) for simulations of the particular OM.}\label{sr_aic}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\begin{center}
\includegraphics[width = 1.4\textwidth]{sr_aic_plots}
\end{center}
\caption{Estimated probability of lowest AIC from logistic regression on the log-standard deviation of the true log(SSB) in each simulation for estimating model with Beverton-Holt SRRs, rather than the otherwise equivalent EM without the SRR. Results are conditional on alternative assumptions for median natural mortality (estimated or known) and on EMs having the correct process error structure: R and R+S (A), R+Sel (B), R+M (C), or R+q (D). Rug along x-axis denotes $SD(\log(SSB))$ values for each simulation and polygons represent 95\% confidence intervals.}\label{sr_aic_supp}
\end{figure}
\end{landscape}